### 2014-10-17 inst/writeup.Rnw
add paper

\documentclass[11pt]{article}
\usepackage{xspace}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
% See the ``Article customise'' template for come common customisations

\title{A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}}
\author{Sam Tyner and Heike Hofmann}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle

\section{Introduction}
\par The three necessary elements of any network visualization are the vertices, the edges, and the layout.  But once those three items are visualized, we usually find our visualization to be lacking.  We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree 10 twice the size of vertices of degree 5.  Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by Gabor Csardi, \texttt{sna} by Carter T. Butts, and \texttt{network} by Carter T. Butts et al. but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network. For instance, the \texttt{igraph} package allows for coloring vertices by groups but the user must assign the colors to each vertex individually as opposed to assigning color by a grouping factor variable. \\

\par The \texttt{GGally} package by Barret Schloerke et al. contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these 3 necessary graph attributes.  The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.  Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.  
  \subsection{Data Structure}
  In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information. The vertex data frame should contain all the relevant vertex information. \emph{The only necessary variables are the vertex names/labels and the vertex degree (if undirected) or the vertex in degree and out degree (if directed).  The names and degrees are stored in columns and the rows are the vertices.  Other values of interest, such as grouping variables, for each vertex should be stored as columns in this dataframe, with an observation for each vertex.}  
  \par The edge data frame should contain all the relevant edge information.  \emph{The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network. The From and To vertices should be listed according to the names of the vertices in the vertex information data frame. Other variables may also be included for each edge, such as the edge weight or grouping variable.  As before, the variables of interest are columns in the data frame and the rows are each edge in the network.}\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}

  \subsection{Vertex Aesthetics}
  In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.  The \texttt{vlabel} aesthetic is the name or number assigned to each vertex in order to uniquely identify it.  This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be readable by the human eye.  The \texttt{vfill} aesthetic can be an identity color (e.g. \texttt{color = "red"} outside of the \texttt{aes} statement) to change the fill color of all vertices, or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it fills the vertices of the graph according to this grouping.  It can also fill the vertices according to the different values of  some numeric variable, such as degree.  The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_fill\_brewer}.  The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs.  \footnote{Sam's questions/notes: Is ``degree" going to be a column in our data frame? Or will we have to write something to detect degree size?} It an also be set to any other numeric variable associated with the vertices in the network.  The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.  It can also change the shape of the vertices according to some grouping variable.  Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10, or to some numeric variable of interest that takes on the values $[0,1]$.

\par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion. 
  \subsection{Edge Aesthetics}
  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for lines.  The \texttt{elabel} aesthetic is the name or number assigned to each edge.  This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.  The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to a grouping variable.  It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge. The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the many different types of dotted lines in \texttt{ggplot2}.  The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.  The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability.  Finally, the \texttt{directed} aesthetic is a logical value that identifies whether or not the network is directed.  When this value is true, the arrows change size, color, etc. with the previous aesthetics.

  \subsection{Layout Aesthetics}
  \emph{There are many layouts to choose from, and the default layout changes according the size of the graph.  For smaller networks, with less than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. **Now justify these two statements with research to be done.**} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fructerman-Reingold. Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.  This aesthetic takes only the identity colors, not additional variables.\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.  I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.}

\end{document}

### 2014-10-20 inst/writeup.Rnw
small changes in write up

@@
 \footnote{Sam's questions/notes: Is ``degree" going to be a column in our data frame?
-Or will we have to write something to detect degree size?} It an also be set to any other numeric variable associated with the vertices in the network.
+Or will we have to write something to detect degree size?} It can also be set to any other numeric variable associated with the vertices in the network.
 The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.

### 2014-10-20 R/first.R
add plotg

@@ -52,3 +52,35 @@ GeomNet <- proto::proto(ggplot2:::Geom, {
   required_aes <- c("x", "y")
   default_aes <- function(.) aes(shape=16, colour="black", size=10, fill = NA, alpha = NA) 
 })
+
+#' Moritz Marbach's function
+#' 
+#' on blog at http://sumtxt.wordpress.com/2011/07/02/visualizing-networks-with-ggplot2-in-r/
+#' updated to new ggplot2 reqs
+plotg <- function(net, value=NULL) {
+  m <- as.matrix.network.adjacency(net) # get sociomatrix #pkg: network
+  # get coordinates from Fruchterman and Reingold's force-directed placement algorithm.
+  plotcord <- data.frame(gplot.layout.fruchtermanreingold(m, NULL))  #pkg: sna
+  # or get it them from Kamada-Kawai's algorithm: 
+  # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL)) # pkg: sna
+  colnames(plotcord) = c("X1","X2")
+  edglist <- as.matrix.network.edgelist(net) #pkg: network
+  edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
+  plotcord$elements <- as.factor(get.vertex.attribute(net, "elements")) #pkg: network
+  colnames(edges) <-  c("X1","Y1","X2","Y2")
+  edges$midX  <- (edges$X1 + edges$X2) / 2
+  edges$midY  <- (edges$Y1 + edges$Y2) / 2
+  pnet <- ggplot()  + 
+    geom_segment(aes(x=X1, y=Y1, xend = X2, yend = Y2), 
+                 data=edges, size = 0.5, colour="grey") +
+    geom_point(aes(X1, X2,colour=elements), data=plotcord) +
+    scale_colour_brewer(palette="Set1") +
+    scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL) +
+    # discard default grid + titles in ggplot2 
+    theme(panel.background = element_blank(), legend.position="none", #changed opts to theme
+          axis.title.x = element_blank(), axis.title.y = element_blank(), #changed theme_blank to element_blank 
+          legend.background = element_rect(colour = NA),
+          panel.background = element_rect(fill = "white", colour = NA), 
+          panel.grid.minor = element_blank(), panel.grid.major = element_blank())
+  return(print(pnet))
+}

### 2014-10-20 R/first.R
messing broke it, but it installs

@@ -8,50 +8,126 @@
 minx <-  function(x, ...) return(min(x, ...))
 
 #' @export
+GeomNet <- proto::proto(ggplot2:::Geom, {
+  objname <- "net"
+  
+  draw_groups <- function(., ...) .$draw(...)
+  draw <- function(., data, coordinates, na.rm = FALSE, ...) {    
+    browser()
+    # net is adjacency matrix
+    #    if (empty(data)) return(zeroGrob())
+    if (empty(net)) return(zeroGrob())
+    
+    
+    m <- as.matrix.network.adjacency(net) # get sociomatrix #pkg: network
+    # get coordinates from Fruchterman and Reingold's force-directed placement algorithm.
+    plotcord <- data.frame(gplot.layout.fruchtermanreingold(m, NULL))  #pkg: sna
+    # or get it them from Kamada-Kawai's algorithm: 
+    # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL)) # pkg: sna
+    colnames(plotcord) = c("x","y")
+    nodes_grob <- GeomPoint$draw(plotcord, ...)
+    
+    edglist <- as.matrix.network.edgelist(net) #pkg: network
+    edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
+    plotcord$elements <- as.factor(get.vertex.attribute(net, "elements")) #pkg: network
+    colnames(edges) <-  c("x","y","xend","yend")
+    edges_grob <- GeomSegment(edges, ...)
+    #edges$midX  <- (edges$X1 + edges$X2) / 2
+    #edges$midY  <- (edges$Y1 + edges$Y2) / 2
+    
+    
+    
+    with(coord_transform(coordinates, data, scales), 
+         ggname(.$my_name(), grobTree(
+           nodes_grob,
+           edges_grob
+         )))
+  }
+  # draw_legend <- function(., data, ...) {
+  # data <- aesdefaults(data, .$default_aes(), list(...))
+  
+  # with(data,
+  # pointsGrob(0.5, 0.5, size=unit(size, "mm"), pch=shape, 
+  # gp=gpar(
+  # col=alpha(colour, alpha), 
+  # fill=alpha(fill, alpha), 
+  # fontsize = size * .pt)
+  # )
+  # )
+  # }
+  
+  browser()
+  default_stat <- function(.) StatIdentity
+  required_aes <- c("net")
[…]

### 2014-10-26 R/first.R
moving some stuff around; still spits out "non-function" error.

@@ -8,13 +8,24 @@
 minx <-  function(x, ...) return(min(x, ...))
 
 #' @export
+#' @example
+#' data(mtcars)
+#' library(ggplot2)
+#' ggplot(data=mtcars) + geom_net(aes(x=mpg, y=disp)) 
+geom_net <- function (mapping = NULL, data = NULL, stat = "identity", position = "identity",
+                      na.rm = FALSE, ...) {
+  browser()
+  GeomNet$new(mapping = mapping, data = data, stat = stat, position = position, 
+              na.rm = na.rm, ...)
+}
+
 GeomNet <- proto::proto(ggplot2:::Geom, {
   objname <- "net"
   
   draw_groups <- function(., ...) .$draw(...)
   draw <- function(., data, coordinates, na.rm = FALSE, ...) {    
     browser()
-    # net is adjacency matrix
+    # net is network object from network package
     #    if (empty(data)) return(zeroGrob())
     if (empty(net)) return(zeroGrob())
     
@@ -24,37 +35,33 @@ GeomNet <- proto::proto(ggplot2:::Geom, {
     plotcord <- data.frame(gplot.layout.fruchtermanreingold(m, NULL))  #pkg: sna
     # or get it them from Kamada-Kawai's algorithm: 
     # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL)) # pkg: sna
-    colnames(plotcord) = c("x","y")
+    colnames(plotcord) <- c("x","y")
     nodes_grob <- GeomPoint$draw(plotcord, ...)
     
     edglist <- as.matrix.network.edgelist(net) #pkg: network
     edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
     plotcord$elements <- as.factor(get.vertex.attribute(net, "elements")) #pkg: network
     colnames(edges) <-  c("x","y","xend","yend")
-    edges_grob <- GeomSegment(edges, ...)
-    #edges$midX  <- (edges$X1 + edges$X2) / 2
-    #edges$midY  <- (edges$Y1 + edges$Y2) / 2
+    edges_grob <- GeomSegment$draw(edges, ...) #add $draw like in nodes_grob
     
-    
-    
-    with(coord_transform(coordinates, data, scales), 
+    with(coord_transform(coordinates, data), #got rid of scales b/c we don't have it anywhere else
          ggname(.$my_name(), grobTree(
            nodes_grob,
[…]

### 2014-10-26 R/first.R
attempting to get rid of error. still no luck.

@@ -13,10 +13,10 @@ minx <-  function(x, ...) return(min(x, ...))
 #' library(ggplot2)
 #' ggplot(data=mtcars) + geom_net(aes(x=mpg, y=disp)) 
 geom_net <- function (mapping = NULL, data = NULL, stat = "identity", position = "identity",
-                      na.rm = FALSE, ...) {
+                        na.rm = FALSE, ...) {
   browser()
-  GeomNet$new(mapping = mapping, data = data, stat = stat, position = position, 
-              na.rm = na.rm, ...)
+  GeomNet$new(mapping = mapping, data = data, stat = stat, position = position,
+                na.rm = na.rm, ...)
 }
 
 GeomNet <- proto::proto(ggplot2:::Geom, {
@@ -25,10 +25,11 @@ GeomNet <- proto::proto(ggplot2:::Geom, {
   draw_groups <- function(., ...) .$draw(...)
   draw <- function(., data, coordinates, na.rm = FALSE, ...) {    
     browser()
+    data <- remove_missing(data, na.rm,
+                           c("net"), name = "geom_net")
     # net is network object from network package
     #    if (empty(data)) return(zeroGrob())
     if (empty(net)) return(zeroGrob())
-    
     
     m <- as.matrix.network.adjacency(net) # get sociomatrix #pkg: network
     # get coordinates from Fruchterman and Reingold's force-directed placement algorithm.
@@ -36,13 +37,13 @@ GeomNet <- proto::proto(ggplot2:::Geom, {
     # or get it them from Kamada-Kawai's algorithm: 
     # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL)) # pkg: sna
     colnames(plotcord) <- c("x","y")
-    nodes_grob <- GeomPoint$draw(plotcord, ...)
+    nodes_grob <- GeomPoint(plotcord, ...)
     
     edglist <- as.matrix.network.edgelist(net) #pkg: network
     edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
     plotcord$elements <- as.factor(get.vertex.attribute(net, "elements")) #pkg: network
     colnames(edges) <-  c("x","y","xend","yend")
-    edges_grob <- GeomSegment$draw(edges, ...) #add $draw like in nodes_grob
+    edges_grob <- GeomSegment(edges, ...)
     
     with(coord_transform(coordinates, data), #got rid of scales b/c we don't have it anywhere else
          ggname(.$my_name(), grobTree(

### 2014-10-26 CC/writeup.Rnw
Add write-up to github. Also made some additional comments about data structure.

--- 2014-10-17 inst/writeup.Rnw
@@
 
+\section{More Thoughts (26 Oct 14)}
+I was thinking some more about the data structure necessary to create a network, and I think it might be possible to just use one data frame.
+If we have a data frame (called \texttt{df} below) with ``from" and ``to" columns as the ``x" and ``y" columns, respectively, we could get the vertex names from that by just using an internal command like:
+
+<<r1, eval=FALSE>>=
+vnames <- unique(c(unique(df$x), unique(df$y)))
+@
+
+This could simply be the default for the \texttt{vlabel} aesthetic.
+Then, the \texttt{vlabel} aesthetic could really just be a logical one, which gets the label of \texttt{vnames} as above if \texttt{vlabel = TRUE} and gets nothing if \texttt{vlabel = FALSE}.
+To handle the rest of the aesthetics with just one dataframe,  the user would have to use \texttt{merge} or something of the like to join the vertex characteristics to the edge data frame and characteristics.
+An example of this with the blood data set is below:
+
+<<r2, echo=FALSE>>=
+all <- paste(sort(rep(c('A', "B", 'O', "AB"),2)), rep(c("+","-"),4), sep = '')
+A <- all[1:2]
+B <- all[5:6]
+AB <- all[3:4]
+O <- all[7:8]
+
+On <- cbind(from = rep('O-',8), to = all, letter = 'O')
+Op <- cbind(from = rep('O+',4), to = c(AB[1], A[1], B[1],O[1]), letter = 'O')
+Bn <- cbind(from = rep('B-',4), to = c(AB, B), letter = 'B')
+Bp <- cbind(from = rep('B+',2), to = c(AB[1], B[1]), letter = 'B')
+An <- cbind(from = rep('A-',4), to = c(AB, A), letter = 'A')
+Ap <- cbind(from = rep('A+',2), to = c(AB[1], A[1]), letter = 'A')
+ABn <- cbind(from = rep('AB-',2), to = AB, letter = 'AB')
+ABp <- cbind(from = 'AB+', to = 'AB+', letter = 'AB')
+
+blood <- data.frame(rbind(ABn, ABp, An, Ap, Bn, Bp, On, Op))
+
+blood.vertices <- unique(blood[,c(1,3)])
+names(blood.vertices) <- c("type", "group")
+blood.edges <- blood[,1:2]
+@
+<<r3, tidy=TRUE>>=
+blood.vertices
+head(blood.edges)
+
+names(blood.vertices) <- c('from', 'group')
+blood.df <- merge(blood.edges, blood.vertices, by = 'from')
+head(blood.df)
+@
+After this is done, the \texttt{vcolor} aesthetic could simply be set to \texttt{vcolor = group} and there wouldn't be any need for a separate data frame to be required to pass to \texttt{geom\_ net}.
+Furthermore, all of the edge aesthesics could simply be set to the variables that are columns in the \texttt{df}, which were determined prior to merging the vertex information with the edge information.
+For example, suppose we wanted to color the edges of the blood donation network by whether or not the blood in the ``from" vertex is going to the same or different blood group.
+i.e.
+the edge from O- to O+ is the same color as the edge from A- to A+, etc.
+but the edge from O- to AB+ is a different color.
+Then the data would have looked like:
+
+<<r4, tidy=TRUE>>=
+blood.edges$group_to <- c(rep('same',3),rep('diff',2),rep('same',2),'diff','same',rep('diff',2),rep('same',2),'diff','same',rep('diff',6),rep('same',2),rep('diff',3),'same')
+head(blood.edges)
+blood.df <- merge(blood.edges, blood.vertices, by = 'from')
+head(blood.df)
+@
+
+Then, the \texttt{ecolor} aesthestic would just be set to \texttt{ecolor = group\_to}.
 \end{document}

### 2015-03-11 CC/writeup.Rnw
Moving rnw file to correct location
 Please enter the commit message for your changes. Lines starting

--- 2014-10-20 inst/writeup.Rnw
@@
 This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be readable by the human eye.
-The \texttt{vfill} aesthetic can be an identity color (e.g.
-\texttt{color = "red"} outside of the \texttt{aes} statement) to change the fill color of all vertices, or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it fills the vertices of the graph according to this grouping.
-It can also fill the vertices according to the different values of  some numeric variable, such as degree.
-The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_fill\_brewer}.
-The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
+The \texttt{vcolor} aesthetic can be an identity color (e.g.
+\texttt{color = "red"} outside of the \texttt{aes} statement) to change the color of all vertices, or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.
+It can also color the vertices according to the different values of  some numeric variable, such as degree.
+The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.
+  %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
   The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs.

### 2015-03-11 CC/writeup.Rnw
couple of references included

@@
 \documentclass[11pt]{article}
+\usepackage{asa}
 \usepackage{xspace}
+\usepackage{natbib}
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 
+%\definecolor{Blue}{rgb}{0,0,0.5}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\sct}[1]{{\color{purple} #1}}
+
 \title{A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}}
@@
 \date{} % delete this line to display the current date
-
 %%% BEGIN DOCUMENT
@@
 \maketitle
+\begin{abstract}
+Abstract will be here
+\end{abstract}
 
@@
 We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree 10 twice the size of vertices of degree 5.
-Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by Gabor Csardi, \texttt{sna} by Carter T.
-Butts, and \texttt{network} by Carter T.
-Butts et al.
-but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network.
+Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by \citet{igraph}, \texttt{sna} by \citet{sna}, and \texttt{network} by \citet{network.jss, network} but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network.
 For instance, the \texttt{igraph} package allows for coloring vertices by groups but the user must assign the colors to each vertex individually as opposed to assigning color by a grouping factor variable.
@@
 
+\bibliographystyle{asa}
+\bibliography{references}
+
 \end{document}

### 2015-03-11 CC/writeup.Rnw
mini suggestion

@@
 \section{Introduction}
+\hh{XXX Start out with an example :) ... madmen? }
+
 \par The three necessary elements of any network visualization are the vertices, the edges, and the layout.
 But once those three items are visualized, we usually find our visualization to be lacking.
-We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree 10 twice the size of vertices of degree 5.
+We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.
 Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by \citet{igraph}, \texttt{sna} by \citet{sna}, and \texttt{network} by \citet{network.jss, network} but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network.
@@
 
-\par The \texttt{GGally} package by Barret Schloerke et al.
-contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these 3 necessary graph attributes.
-The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
+\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these \hh{XXX write out all numbers of size ten and under in words.
+} 3 necessary graph attributes.
+\hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
 Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.

### 2015-03-11 CC/writeup.Rnw
additional sections I'd like

@@
 Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
+%
   \subsection{Data Structure}
@@
 For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE.
-**Now justify these two statements with research to be done.**} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fructerman-Reingold.
+**Now justify these two statements with research to be done.** \hh{I don't think that you'd have to defend your choices of defaults.
+They are defaults, and can be changed.
+}} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fruchterman-Reingold.
 Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.
 This aesthetic takes only the identity colors, not additional variables.\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.
-I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.}
+I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.
+\hh{You're fine - those are things, which people are going to have strong opinions on without actual backup.
+You provide the environment for people to easily use network data in a familiar setting.}}
+
+\section{Many Many Examples}
+\section{Technical Details}
 

### 2015-03-16 CC/writeup.Rnw
what i have right now. working on it all day today.

@@
 %\definecolor{Blue}{rgb}{0,0,0.5}
+\definecolor{purple}{rgb}{.4,0,.8}
 \newcommand{\hh}[1]{{\color{magenta} #1}}
 \newcommand{\sct}[1]{{\color{purple} #1}}
+\newcommand{\sctq}[1]{{\color{red} #1}} %things in red are my questions for you!
 
@@
 
-\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these \hh{XXX write out all numbers of size ten and under in words.
-} 3 necessary graph attributes.
-\hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
+\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
+\hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} \sctq{Not quite sure what to say here.
+What follows is my guesstimate.} \sct{Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
+The vertices in a network and the edges between them compose the structure of a network, and being able to discover patterns among them visually is a key part of network analysis.
+Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
 Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
@@
   \par The edge data frame should contain all the relevant edge information.
-\emph{The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network.
+The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network.
 The From and To vertices should be listed according to the names of the vertices in the vertex information data frame.
 Other variables may also be included for each edge, such as the edge weight or grouping variable.
-As before, the variables of interest are columns in the data frame and the rows are each edge in the network.}\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}
+As before, the variables of interest are columns in the data frame and the rows are each edge in the network.%\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}
 
@@
 
-  \subsection{Layout Aesthetics}
-  \emph{There are many layouts to choose from, and the default layout changes according the size of the graph.
+  \subsection{Layout Argument}
+  \sct{There are many layouts to choose from, but the default layout is the Kamada-Kawai layout.  This is a force-directed layout for undirected networks \citep{kamadakawai}}.
+  %and the default layout changes according the size of the graph.
 For smaller networks, with less than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE.
 For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE.
-**Now justify these two statements with research to be done.** \hh{I don't think that you'd have to defend your choices of defaults.
+%**Now justify these two statements with research to be done.** \hh{I don't think that you'd have to defend your choices of defaults.
 They are defaults, and can be changed.
-}} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fruchterman-Reingold.
-Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.
-This aesthetic takes only the identity colors, not additional variables.\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.
+}}
+  The \texttt{layout} \sct{argument in the \texttt{geom\_net} function} can be changed to any of the layouts also found in the \texttt{sna} package, such as circular, eigen, or Fruchterman-Reingold.
+\sctq{Should I insert a table of layouts here?
+Or maybe not since it can't actually change the layout yet?
+And I took out the thing about the background color because I figure it can just be changed with \texttt{theme()}} %Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.
+This aesthetic takes only the identity colors, not additional variables.%\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.
 I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.
@@
 \section{Technical Details}
+\sct{
+In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
+The stat file performs all the necessary calculations on the two input data frames, and the geom file performs the drawing of the network layer.
+}
 

### 2015-03-16 CC/writeup.Rnw
done with section 3 for now.

@@
 \section{Technical Details}
+\sctq{I have no idea how detailed I should be.
+I'm going to attempt to be detailed without parsing each line in the code, but I do want to describe how everything works in detailed enough descriptions.
+I know you'll tell me if I'm too detailed or not detailed enough!}
 \sct{
 In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
-The stat file performs all the necessary calculations on the two input data frames, and the geom file performs the drawing of the network layer.
-}
+The stat file performs all the necessary calculations on the two input data frames, and the geom file performs the drawing of the network.
+  \subsection{Creating \texttt{stat\_net}}
+  In order to plot the network with only the edge connections and vertex names, we relied on the \texttt{sna} and \texttt{network} packages \citep{sna, network.jss}.
+First, we used the edges data frame to construct a network object with the \texttt{as.network()} function, then constructed an adjacency matrix from that network using the \texttt{as.matrix.network.adjacency()} function.
+This adjacency matrix is then passed to the chosen \texttt{gplot.layout.*()} function from the \texttt{sna} package.
+\emph{When the layout argument is changed in the \texttt{geom\_net()} function, the corresponding \texttt{gplot.layout.*()} function in \texttt{sna} is called with a \texttt{do.call()} statement.} This layout function produces a matrix of coordinates of the vertices, which we then transform into a data frame containing the coordinates of the edges using the \texttt{as.matrix.network.edgelist()} function in \texttt{network}.
+We are grateful to Moritz Marbach for his \texttt{gplot()} function which we used when creating this process \citep{marbach}.
+At the end of the previous routine, we have a data frame with \texttt{x, y, xend,} and \texttt{yend} columns describing the edges and another data frame with \texttt{x,y} columns describing the vertices.
+These two data frames, along with the many \texttt{aes()} and other argument options get passed on to the actual network geometry for drawing.
+  \subsection{Creating \texttt{geom\_net}}
+  We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
+Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.
+We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function.
+Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function.
+\emph{Finally, we use the \texttt{GeomText} draw function to add the vertex and edge labels when the \texttt{vlabel} and \texttt{elabel} arguments, respectively, are true.
+*insert more detail here after figuring it out*}
+} \sctq{This doesn't seem very long, but I already feel like there's too much detail! Should I add more?}
 

### 2015-03-16 CC/writeup.Rnw
write-up as of 6:45 on Monday.  new data was also added to dropbox.

@@
 \usepackage{natbib}
+\usepackage{url} 
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 
+% <<readingdata, message=FALSE>>=
+%  library(ggplot2)
+%  theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
+%  data(blood, football, theme_elements)
+%  email.net <- read.csv("~/Dropbox/ggplot2/email-data.csv") 
+%  email.net <- subset(email.net, day %in% c(9,13))
+%  email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to))))) 
+% @
+
 \section{Introduction}
@@
 Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
-%
+
   \subsection{Data Structure}
@@
 \section{Many Many Examples}
+\sctq{Format-wise, is it ok if each example is its own subsection?}\\
+\sct{
+In this section, we demonstrate the capabilities of \texttt{geom\_net} in a series of examples. 
+  \subsection{Blood Donation}
+  \sctq{I had to comment out all the code chunks for now. Not sure why but it won't compile if they're there.}
+  \emph{Note: would be best with labels and direction}
+% # <<blood_ex>>=
+% # ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
+% # @
+  \subsection{Email Network}
+  \emph{Note: would be best with labels and direction}
+% # <<email_ex>>=
+% # ggplot(data = email.net, aes(from_id = From, to_id = to, ealpha = 0.1,vsize = 2)) + geom_net(vertices = email.names) + expand_limits(x=c(-1,5), y = c(-1,5)) + facet_wrap(~day) + theme_net
+% # @
+  \subsection{\texttt{ggplot2} Theme Elements}
+  \emph{Note: would be best with labels and direction}
+% # <<theme_ex>>=
+% # #ggplot(data = theme_elements$edges, aes(from_id = child, to_id = parent)) + geom_net(vertices = theme_elements$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net
+% # @
+  \subsection{Madmen Network (again?? or no?)}
+  \sctq{leaving blank for now.}
+  \subsection{College Football}
+  \emph{Note: would be best with coloring vertices by conference} \citet{football}
+% # <<football_ex>>=
+% # ggplot(data = football$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = football$vertices) + expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) + theme_net
+% # @
+  \subsection{Les Mis\'{e}rables}
+  \emph{Haven't looked at this one yet. Just found it on MEJ Newmans website.  http://www-personal.umich.edu/~mejn/netdata/}
+}
+
 \section{Technical Details}
-\sctq{I have no idea how detailed I should be.
+\sctq{I have no idea exactly how detailed I should be.
 I'm going to attempt to be detailed without parsing each line in the code, but I do want to describe how everything works in detailed enough descriptions.

### 2015-03-17 CC/writeup.Rnw
example code

@@
   \emph{Note: would be best with labels and direction}
-% # <<blood_ex>>=
-% # ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
-% # @
+}
+<<blood_ex>>=
+library(ggplot2) # this needs to be the ggplot2 with geom_net
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
+@
+\sct{
   \subsection{Email Network}

### 2015-03-17 CC/writeup.Rnw
add les mis example code

--- 2015-03-16 CC/writeup.Rnw
@@
 \end{abstract}
-
-% <<readingdata, message=FALSE>>=
-%  library(ggplot2)
-%  theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
-%  data(blood, football, theme_elements)
-%  email.net <- read.csv("~/Dropbox/ggplot2/email-data.csv") 
-%  email.net <- subset(email.net, day %in% c(9,13))
-%  email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to))))) 
-% @
+<<chunk_setup, echo=FALSE>>=
+opts_chunk$set(echo=FALSE, message = FALSE)
+@
+<<readingdata, message=FALSE, echo=FALSE>>=
+ library(knitr)
+ library(ggplot2)
+ theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
+ data(blood, football, theme_elements)
+ email.net <- read.csv("~/Dropbox/ggplot2/email-data.csv") 
+ email.net <- subset(email.net, day %in% c(9,13))
+ email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to))))) 
+@
 
@@
 \sctq{Should I insert a table of layouts here?
-Or maybe not since it can't actually change the layout yet?
 And I took out the thing about the background color because I figure it can just be changed with \texttt{theme()}} %Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.
@@
   \sctq{I had to comment out all the code chunks for now. Not sure why but it won't compile if they're there.}
-  \emph{Note: would be best with labels and direction}
-% # <<blood_ex>>=
-% # ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
-% # @
+  \emph{Note: would be best with labels and direction}\\
+  In this network, there are eight nodes and 27 edges.  The nodes represent the eight different blood types possible in humans
+%<<blood_ex,highlight=TRUE>>=
+%ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
+%@
   \subsection{Email Network}
@@
 % # <<theme_ex>>=
-% # #ggplot(data = theme_elements$edges, aes(from_id = child, to_id = parent)) + geom_net(vertices = theme_elements$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net
+% # #ggplot(data = g$edges, aes(from_id = child, to_id = parent)) + geom_net(vertices = g$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net
 % # @
@@
   \subsection{Les Mis\'{e}rables}
-  \emph{Haven't looked at this one yet. Just found it on MEJ Newmans website.  http://www-personal.umich.edu/~mejn/netdata/}
+  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!}
+% <<lesmis_ex>>=
+% ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+% @
 }

### 2015-03-17 CC/writeup.Rnw

@@
 %@
+  \emph{Note: would be best with labels and direction}
+}
+<<blood_ex>>=
+library(ggplot2) # this needs to be the ggplot2 with geom_net
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
+@
+\sct{
   \subsection{Email Network}

### 2015-03-17 CC/writeup.Rnw
what I have as of finding that error.

--- 2015-03-17 CC/writeup.Rnw
@@
 <<chunk_setup, echo=FALSE>>=
-opts_chunk$set(echo=FALSE, message = FALSE)
+opts_chunk$set(tidy = TRUE, echo=FALSE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
 @
-<<readingdata, message=FALSE, echo=FALSE>>=
+<<readingdata>>=
  library(knitr)
- library(ggplot2)
+ library(ggplot2) # needs to be the version with geom_net! 
  theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
@@
  email.net <- read.csv("~/Dropbox/ggplot2/email-data.csv") 
- email.net <- subset(email.net, day %in% c(9,13))
  email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to))))) 
@@
 \section{Many Many Examples}
-\sctq{Format-wise, is it ok if each example is its own subsection?}\\
+\sctq{Format-wise, is it ok if each example is its own subsection?
+Also, can you check the references I added to the .bib file to make sure I'm creating them correctly?
+I've just been appending them.}\\
 \sct{
@@
   \subsection{Blood Donation}
-  \sctq{I had to comment out all the code chunks for now. Not sure why but it won't compile if they're there.}
   \emph{Note: would be best with labels and direction}\\
-  In this network, there are eight nodes and 27 edges.  The nodes represent the eight different blood types possible in humans
-%<<blood_ex,highlight=TRUE>>=
-%ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel later
-%@
+  In this directed network, there are eight vertices and 27 edges.
+The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive and negative types.
+\sctq{Can I cite wikipedia here?
+Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a persone whose blood type is that of a corresponding \emph{to} vertex.
+In the example below, loops are removed.
+Loops exist on every vertex in this example, as blood of matching ABO and RhD type can always be received.
+}
+<<blood_ex, echo=TRUE>>=
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, layout='circle') + expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
+@
+\sct{
   \subsection{Email Network}
-  \emph{Note: would be best with labels and direction}
-% # <<email_ex>>=
-% # ggplot(data = email.net, aes(from_id = From, to_id = to, ealpha = 0.1,vsize = 2)) + geom_net(vertices = email.names) + expand_limits(x=c(-1,5), y = c(-1,5)) + facet_wrap(~day) + theme_net
-% # @
+  \emph{Note: would be best with coloring and direction}
+  This email network comes from the 2014 VAST Challenge \citep{emailnet}.
+It is a directed network of emails between company employees with 55 vertices and 9,063 edges.
+Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees.
+The arrow of the directed edge points to the recipient(s) of the email.
+The network contains two business weeks of emails across the entire company.
+}
+<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
+ggplot(data = email.net, aes(from_id = From, to_id = to, )) + geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) + expand_limits(x=c(-1.75,1.75), y = c(-1.75,1.75)) + theme_net #add direction later
+@  
+\sct{This network is plotted in figure \ref{fig:email_ex}.
+This plot is very crowded, and it is difficult to distinguish structure in the network when it there is so much over-plotting.
+To remedy this, we facet the network by day to examine how the company's communications change throughout the week.
+}
+<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
+ggplot(data = email.net, aes(from_id = From, to_id = to, )) + geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) + expand_limits(x=c(-2,2), y = c(-2,2)) + facet_wrap(~day, nrow = 2) + theme_net #add direction later
+@
+\sct{This plot is shown in figure \ref{fig: email_ex_facet}.
+With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving?
+will know with direction option added later} several emails each day to a majority of the company.
+We can also see clusters that have formed on the perimeter of each day's network where the employees are communicating heavily with each other, but not much outside of their cluster.
+We can also see that there is heavier inter-cluster communication on different days.
+For example, day 17 has more concentrated in-group communication than day 15.
+Additionally, we can better distinguish the communication of the outlying employee, who only corresponds with up to four other complany employees on a given day.} \sctq{talk about clusters being different departments within the company?}
   \subsection{\texttt{ggplot2} Theme Elements}
-  \emph{Note: would be best with labels and direction}
-% # <<theme_ex>>=
-% # #ggplot(data = g$edges, aes(from_id = child, to_id = parent)) + geom_net(vertices = g$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net
-% # @
+  \emph{Note: would be best with labels and direction} \\
+  \sct{This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
+It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.
+There are 53 vertices and 36 edges in this network.
+There is an arrow from one vertex to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.
+For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.
+Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background.}
+<<theme_ex>>=
+ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) + geom_net(vertices = theme_elements$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel & direction later
+@
+\sct{The inheritance structure is plotted in figure \ref{fig:theme_ex}.
   \subsection{Madmen Network (again?? or no?)}

### 2015-03-18 CC/writeup.Rnw
haven't pushed in a while

@@
 <<chunk_setup, echo=FALSE>>=
-opts_chunk$set(tidy = TRUE, echo=FALSE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
+opts_chunk$set(tidy = TRUE, echo=FALSE, eval = FALSE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
 @
@@
  theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements)
- email.net <- read.csv("~/Dropbox/ggplot2/email-data.csv") 
+ data(blood, football, theme_elements, lesmis)
+ email.net <- read.csv("~/Desktop/NetworksResearch/ggnet/CC/data/email-data.csv") 
  email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to))))) 
@@
 Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background.}
-<<theme_ex>>=
+<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
 ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) + geom_net(vertices = theme_elements$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel & direction later
@@
   \subsection{College Football}
-  \emph{Note: would be best with coloring vertices by conference} \citet{football}
-% # <<football_ex>>=
-% # ggplot(data = football$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = football$vertices) + expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) + theme_net
-% # @
+  \emph{Note: would be best with coloring vertices by conference} 
+  This next example comes from M.E.J.
+Newman's network data web page \citep{football}.
+It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.
+There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
+<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
+ggplot(data = football$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = football$vertices, ealpha = 0.1) + expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) + theme_net
+@
+\sct{
+The network of football games is given in figure \ref{fig:football_ex}.
   \subsection{Les Mis\'{e}rables}
-  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!}
-% <<lesmis_ex>>=
-% ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
-% @
-}
+  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!}}
+<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+@
 

### 2015-03-18 CC/writeup.Rnw
all writing so far. eval is set to false in opts_chunk$set to avoid printing browser output. also added data files to data folder.

@@
  data(blood, football, theme_elements, lesmis)
+ load("~/Desktop/NetworksResearch/ggnet/CC/data/madmen.rda")
  email.net <- read.csv("~/Desktop/NetworksResearch/ggnet/CC/data/email-data.csv") 
@@
 \section{Introduction}
-\hh{XXX Start out with an example :) ... madmen? }
+\sct{At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.
+Here we refer to the lines as edges and the points as vertices, although these are also called nodes.
+These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study.
+Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
+The edges in a network can also be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
+The world wide web is an example of a directed network because one webpage can link to another, and not necessarily the other way around.
+Undirected edges, however, are simply connections between vertices.
+Coauthorship networks that encode information about academic publications are examples of undirected networks because if two people author a paper together, that creates a connection between them that is bidirectional.
 
+\par A social network is a network that everyone is a part of in one way or another.
+We do not refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
+To demonstrate the functionality of our geometry for plotting networks, we have chosen an example of a social network from the popular television show Mad Men.
+This network was compiled in \citet{madmen}.
+In this example network, there are 52 vertices and 87 edges.
+Each vertex is a character on the show, and there is an edge between every two characters who have had a romantic relationship.
+}
+<<madmen_ex, fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
+ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net
+@
+\sct{
+This network is shown in figure \ref{fig.cap:madmen_ex}.
+\emph{Notes: add names, color for gender} In the plot, we can see there is one central character who has many more relationships than any other character.
+This vertex represents the main character of the show, Don Draper.
+\par There are many kinds of networks, and networks are extensively studied across many disciplines.
+Many sociologists study social networks, and many biologists study protein networks.
+As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.
+We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}
+}
 \par The three necessary elements of any network visualization are the vertices, the edges, and the layout.
@@
 For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.
-Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background.}
+Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background.
+Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.
+}
 <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
@@
   \subsection{Les Mis\'{e}rables}
-  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!}}
+  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
+  This next network comes from \citet{lesmis}.
+It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.
+There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.
+There are 254 edges in this network.
+The edges are also weighted by the number of coappearances.
+The largest weighting is 31, between the characters Jean Valjean and Cosette.}
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), ealpha = 2/10, layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
 @

### 2015-03-19 CC/writeup.Rnw
more additions to examples

@@
 \newcommand{\sct}[1]{{\color{purple} #1}}
-\newcommand{\sctq}[1]{{\color{red} #1}} %things in red are my questions for you!
+\newcommand{\sctq}[1]{{\color{red} #1}} %things in red are questions for you or concerns that i have.
 
@@
 <<chunk_setup, echo=FALSE>>=
-opts_chunk$set(tidy = TRUE, echo=FALSE, eval = FALSE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
+opts_chunk$set(tidy = TRUE, echo=FALSE, eval = TRUE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
 @
@@
 }
-<<blood_ex, echo=TRUE>>=
+<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
 ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, layout='circle') + expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
@@
 \sct{
+This network is shown in figure \ref{fig:blood_ex}. 
   \subsection{Email Network}

### 2015-03-19 blood_example.R

@@ -1,0 +1,49 @@
+#Testing of the geom!
+
+all <- paste(sort(rep(c('A', "B", 'O', "AB"),2)), rep(c("+","-"),4), sep = '')
+A <- all[1:2]
+B <- all[5:6]
+AB <- all[3:4]
+O <- all[7:8]
+
+On <- cbind(from = rep('O-',8), to = all, letter = 'O')
+Op <- cbind(from = rep('O+',4), to = c(AB[1], A[1], B[1],O[1]), letter = 'O')
+Bn <- cbind(from = rep('B-',4), to = c(AB, B), letter = 'B')
+Bp <- cbind(from = rep('B+',2), to = c(AB[1], B[1]), letter = 'B')
+An <- cbind(from = rep('A-',4), to = c(AB, A), letter = 'A')
+Ap <- cbind(from = rep('A+',2), to = c(AB[1], A[1]), letter = 'A')
+ABn <- cbind(from = rep('AB-',2), to = AB, letter = 'AB')
+ABp <- cbind(from = 'AB+', to = 'AB+', letter = 'AB')
+
+blood <- data.frame(rbind(ABn, ABp, An, Ap, Bn, Bp, On, Op))
+type <- c('AB','A','B','O')
+
+blood.edges <- blood[-which(blood$from == blood$to),]
+blood.vertices <- data.frame(node_id=unique(c(as.character(blood$from), as.character(blood$to))))
+
+#node_id is a required aesthetic. #i have yet to match it with the nodes data in the geom code.
+ggplot(data = blood.edges, aes(from_id= from, to_id = to)) + geom_net(vertices = blood.vertices) + expand_limits(x = -5, y = -1)
+
+
+# #try again: 1/19
+# 
+# #get network coords first, outside of geom
+# net <- as.network(blood.edges, matrix.type = "edgelist") #from network package
+# m <- as.matrix.network.adjacency(net)
+# #change the stuff below later to include multiple layouts (in if else statements?)
+# vert.coord <- data.frame(sna::gplot.layout.kamadakawai(m, NULL)) # from sna package
+# vert.labels <- as.factor(unique(c(as.character(blood.edges$from),as.character(blood.edges$to))))
+# vert.coord$name <- sort(vert.labels) #needed here?
+# names(vert.coord) <- c('x','y','node_id')
+# edgelist <- as.matrix.network.edgelist(net)
[…]

### 2015-03-19 madmenexample.R

@@ -1,0 +1,10 @@
+#mad men example
+data(madmen, package="gcookbook")
+sex <- c("f", "m", "m", "f","m","f",rep("m",5),rep("f",15), "m", "m", "f","f","f","m","m","m","m",rep("f",8),"m","f")
+mm.vert <- data.frame(Name = as.factor(unique(c(madmen$Name1,madmen$Name2))), Gender = as.factor(sex))
+
+madmen <- list(edges = madmen, vertices = mm.vert)
+#save(madmen, file = "~/Desktop/NetworksResearch/ggnet/CC/data/madmen.rda")
+
+ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + 
+  expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net

### 2015-03-20 CC/writeup.Rnw
added protein network example

@@
 <<chunk_setup, echo=FALSE>>=
-opts_chunk$set(tidy = TRUE, echo=FALSE, eval = TRUE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
+opts_chunk$set(tidy = TRUE, echo=FALSE, eval = FALSE, message = FALSE, warning=FALSE,  fig.height = 4, fig.width = 4, fig.align = 'center')
 @
@@
 <<madmen_ex, fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net
+ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices, vlabel = TRUE) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net
 @
@@
 <<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, layout='circle') + expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE) + expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
 @
@@
 @
+\sct{
+  \subsection{Protein Interaction Network in Yeast}
+}
+<<read_data>>=
+yeast <- read.table("~/Desktop/NetworksResearch/ggnet/CC/data/protein.txt")
+yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
+@
+<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.">>=
+ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) + geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+@
 

### 2015-03-21 CC/writeup.Rnw
comments, and a figure included in the pdf

--- 2015-03-19 CC/writeup.Rnw
@@
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
-\geometry{a4paper} % or letter or a5paper or ... etc
+\geometry{letterpaper} % or letter or a5paper or ... etc
 % \geometry{landscape} % rotated page geometry
@@
 \end{abstract}
+\hh{XXX I've set tidy to FALSE; for some reason it's totally messing up the line width.}
 <<chunk_setup, echo=FALSE>>=
-opts_chunk$set(tidy = TRUE, echo=FALSE, eval = TRUE, message = FALSE, fig.height = 4, fig.width = 4, fig.align = 'center')
+opts_chunk$set(tidy = FALSE, echo=FALSE, eval = FALSE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center')
 @
@@
  data(blood, football, theme_elements, lesmis)
- load("~/Desktop/NetworksResearch/ggnet/CC/data/madmen.rda")
- email.net <- read.csv("~/Desktop/NetworksResearch/ggnet/CC/data/email-data.csv") 
+ load("data/madmen.rda") # relative path should work for both of us
+ email.net <- read.csv("data/email-data.csv")
  email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
@@
 \section{Introduction}
-\sct{At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.
-Here we refer to the lines as edges and the points as vertices, although these are also called nodes.
-These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study.
+At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.
+Here, we refer to the lines as edges and the points as vertices, although these are also called nodes.
+These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study \hh{XXX this would be a good point to cite a couple of papers that investigate networks}.
 Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
-The edges in a network can also be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
+The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
 The world wide web is an example of a directed network because one webpage can link to another, and not necessarily the other way around.
@@
 Each vertex is a character on the show, and there is an edge between every two characters who have had a romantic relationship.
-}
-<<madmen_ex, fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net
+
+<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
+ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
+           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
+  expand_limits(x = c(-7.5,7.5), y = c(-7.5,7.5)) + theme_net
 @
-\sct{
-This network is shown in figure \ref{fig.cap:madmen_ex}.
+
+\begin{figure}
+\centering
+\includegraphics{figure/madmen_ex-1.pdf}
+\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romatic relationship.}
+\end{figure}
+
+This network is shown in figure~\ref{fig.cap:madmen_ex}.
 \emph{Notes: add names, color for gender} In the plot, we can see there is one central character who has many more relationships than any other character.
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}
-}
+
 \par The three necessary elements of any network visualization are the vertices, the edges, and the layout.
@@
 <<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, layout='circle') + expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
+  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE) +
+  expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
 @
@@
 <<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to, )) + geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) + expand_limits(x=c(-1.75,1.75), y = c(-1.75,1.75)) + theme_net #add direction later
+ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
+  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
+  expand_limits(x=c(-1.75,1.75), y = c(-1.75,1.75)) +
+  theme_net #add direction later
 @
@@
 <<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to, )) + geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) + expand_limits(x=c(-2,2), y = c(-2,2)) + facet_wrap(~day, nrow = 2) + theme_net #add direction later
+ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
+  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
+  expand_limits(x=c(-2,2), y = c(-2,2)) +
+  facet_wrap(~day, nrow = 2) +
+  theme_net #add direction later
 @
@@
 }
-<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) + geom_net(vertices = theme_elements$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + theme_net #add vlabel & direction later
+<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.', eval=FALSE>>=
+ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
+  geom_net(vertices = theme_elements$vertices) +
+  expand_limits(x = c(-5,5), y = c(-5,5)) +
+  theme_net #add vlabel & direction later
 @
@@
 <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-ggplot(data = football$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = football$vertices, ealpha = 0.1) + expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) + theme_net
+ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
+  geom_net(vertices = football$vertices, ealpha = 0.1) +
+  expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) +
+  theme_net
 @
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, aes(esize = value), ealpha = 2/10, layout = 'fruchtermanreingold') + expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) +
+  geom_net(vertices = lesmis$vertices, aes(esize = value), ealpha = 2/10, layout = 'fruchtermanreingold') +
+  expand_limits(x = c(-25,25), y = c(-25,25)) +
+  theme_net
+@
+\sct{
+  \subsection{Protein Interaction Network in Yeast}
+}
+<<read_data>>=
+yeast <- read.table("data/protein.txt")
+yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
+@
+<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
+ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
+  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
+  expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
 @

### 2015-03-21 CC/writeup.Rnw
vertices need a label column

@@
 <<chunk_setup, echo=FALSE>>=
+library(knitr)
 opts_chunk$set(tidy = FALSE, echo=FALSE, eval = FALSE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center')
@@
 <<readingdata>>=
- library(knitr)
  library(ggplot2) # needs to be the version with geom_net!
@@
  load("data/madmen.rda") # relative path should work for both of us
+names(madmen$vertices)[1] <- "label"
+names(blood$vertices)[1] <- "label"
+
  email.net <- read.csv("data/email-data.csv")
- email.names <- data.frame(name = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
+ email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
+
 @

### 2015-03-21 CC/writeup.Rnw
edge id

@@
 
+This network is shown in figure~\ref{fig.cap:madmen_ex}. \emph{Notes: add names, color for gender}
+\hh{XXX if you show the code, you should acknowledge it in the write-up.
+I think, for the motivating example here, it would be OK to leave out the code ...
+for the examples later on, the code is much more important.}
+
 <<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
@@
 
-This network is shown in figure~\ref{fig.cap:madmen_ex}.
-\emph{Notes: add names, color for gender} In the plot, we can see there is one central character who has many more relationships than any other character.
-This vertex represents the main character of the show, Don Draper.
+ In the plot, we can see one central character who has many more relationships than any other character. This vertex represents the main character of the show, Don Draper.
 \par There are many kinds of networks, and networks are extensively studied across many disciplines.
@@
 As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.
-We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}
+We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
+\hh{XXX How widely used is ggplot2 at the moment? We could include some statistics from \url{http://cran-logs.rstudio.com/} to demonstrate this.}
 
@@
 \par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
-\hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} \sctq{Not quite sure what to say here.
-What follows is my guesstimate.} \sct{Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
+Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
 The vertices in a network and the edges between them compose the structure of a network, and being able to discover patterns among them visually is a key part of network analysis.
-Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
-Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
+Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.
+The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
+Our work builds off of the \texttt{ggnet} function and presents an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
 
   \subsection{Data Structure}
+  \hh{XXX we need a link between the two data frames: we have from_id and to_id in the edge data frame, but we also need an 'id' variable in the vertex data.
+We don't officially have that at the moment - in the code there is a soft requirement (new as of Mar 20) that one of the columns in the edge data set has to be called 'label' ...
+that's a bit stoopid, but I couldn't think of a better name.
+It would make more sense to call this id or edgeid or something along those lines - let's talk about it, and make it a formal requirement.
+}
   In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.

### 2015-03-22 CC/writeup.Rnw
small additions of examples and ggplot2 stats

@@
 \end{abstract}
-\hh{XXX I've set tidy to FALSE; for some reason it's totally messing up the line width.}
 <<chunk_setup, echo=FALSE>>=
@@
 Here, we refer to the lines as edges and the points as vertices, although these are also called nodes.
-These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study \hh{XXX this would be a good point to cite a couple of papers that investigate networks}.
-Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
+These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study.
+\sct{For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.
+Additionally, \citet{networkfailures} examine node failure in interdependent networks like power grids.
+Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C.
+elegans} worm are also etensively studied \citep{smallworld}.}  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
 The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
@@
 This network is shown in figure~\ref{fig.cap:madmen_ex}. \emph{Notes: add names, color for gender}
-\hh{XXX if you show the code, you should acknowledge it in the write-up.
-I think, for the motivating example here, it would be OK to leave out the code ...
-for the examples later on, the code is much more important.}
-
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men', echo=TRUE>>=
+<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
 ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
@@
            aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(-7.5,7.5), y = c(-7.5,7.5)) + theme_net
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
 @
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-\hh{XXX How widely used is ggplot2 at the moment? We could include some statistics from \url{http://cran-logs.rstudio.com/} to demonstrate this.}
+\sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times: approximately 3454 downloads per day.
+It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.}
 
@@
   \subsection{Data Structure}
-  \hh{XXX we need a link between the two data frames: we have from_id and to_id in the edge data frame, but we also need an 'id' variable in the vertex data.
+  \hh{XXX we need a link between the two data frames: we have from\_id and to\_id in the edge data frame, but we also need an 'id' variable in the vertex data.
 We don't officially have that at the moment - in the code there is a soft requirement (new as of Mar 20) that one of the columns in the edge data set has to be called 'label' ...
 that's a bit stoopid, but I couldn't think of a better name.
-It would make more sense to call this id or edgeid or something along those lines - let's talk about it, and make it a formal requirement.
-}
+It would make more sense to call this id or edgeid or something along those lines - let's talk about it, and make it a formal requirement.}
   In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
@@
   geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE) +
-  expand_limits(x = c(-2,2), y = c(-2,2)) + theme_net #add vlabel & direction later
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add vlabel & direction later
 @

### 2015-03-23 CC/writeup.Rnw
all figures in paper now.

@@
 \usepackage{url}
+\usepackage{pdflscape}
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, eval = FALSE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center')
+opts_chunk$set(tidy = FALSE, echo=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
 @
-<<readingdata>>=
+<<readingdata, eval=TRUE>>=
  library(ggplot2) # needs to be the version with geom_net!
@@
 names(blood$vertices)[1] <- "label"
-
+names(theme_elements$vertices)[1] <- "label"
+football$vertices$label <- as.factor(football$vertices$label)
+football$vertices$value <- as.factor(football$vertices$value)
  email.net <- read.csv("data/email-data.csv")
@@
 Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C.
-elegans} worm are also etensively studied \citep{smallworld}.}  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
+elegans} worm are also etensively studied \citep{smallworld}.}  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used \sct{and most complex} networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
 The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
-The world wide web is an example of a directed network because one webpage can link to another, and not necessarily the other way around.
+The world wide web is an example of a directed network because one webpage \sct{may} link to another, and not necessarily the other way around.
 Undirected edges, however, are simply connections between vertices.
@@
 \par A social network is a network that everyone is a part of in one way or another.
-We do not refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
+We do not \sct{necessarily} refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
 To demonstrate the functionality of our geometry for plotting networks, we have chosen an example of a social network from the popular television show Mad Men.
@@
 This network is shown in figure~\ref{fig.cap:madmen_ex}. \emph{Notes: add names, color for gender}
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
+<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men',cache=TRUE>>=
 ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
@@
            aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+  expand_limits(x = c(0,1.5), y = c(0,1)) + theme_net
 @
@@
 \centering
-\includegraphics{figure/madmen_ex-1.pdf}
+\includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
 \caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romatic relationship.}
@@
 ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add vlabel & direction later
+  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2)) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add direction later
 @
 \sct{
-This network is shown in figure \ref{fig:blood_ex}.
+This network is shown in figure \ref{fig.cap:blood_ex}.
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
+\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
+\end{figure}
+
 \subsection{Email Network}
@@
 <<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
+ggplot(data = email.net, aes(from_id = From, to_id = to)) +
   geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
-  expand_limits(x=c(-1.75,1.75), y = c(-1.75,1.75)) +
+  expand_limits(x=c(0,1), y = c(0,1)) +
   theme_net #add direction later
 @
-\sct{This network is plotted in figure \ref{fig:email_ex}.
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
+\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
+\end{figure}
+
+\sct{This network is plotted in figure \ref{fig.cap:email_ex}.
 This plot is very crowded, and it is difficult to distinguish structure in the network when it there is so much over-plotting.
@@
   geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
-  expand_limits(x=c(-2,2), y = c(-2,2)) +
+  expand_limits(x=c(0,1), y = c(0,1)) +
   facet_wrap(~day, nrow = 2) +
@@
 @
-\sct{This plot is shown in figure \ref{fig: email_ex_facet}.
+\begin{landscape}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
+\caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
+\end{figure}
+\end{landscape}
+\sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.
 With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving?
@@
 }
-<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.', eval=FALSE>>=
+<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
 ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices) +
-  expand_limits(x = c(-5,5), y = c(-5,5)) +
+  geom_net(vertices = theme_elements$vertices, vlabel = TRUE) +
+  expand_limits(x = c(0,1), y = c(0,1)) +
   theme_net #add vlabel & direction later
 @
-\sct{The inheritance structure is plotted in figure \ref{fig:theme_ex}.
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
+\caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
+\end{figure}
+
+\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
   \subsection{Madmen Network (again?? or no?)}
@@
 ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = football$vertices, ealpha = 0.1) +
-  expand_limits(x = c(-2.5,2.5), y = c(-2.5,2.5)) +
+  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = I(2), vlabel = TRUE,aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(value)])) +
+  expand_limits(x = c(0,1), y = c(0,1)) +
   theme_net
@@
 \sct{
-The network of football games is given in figure \ref{fig:football_ex}.
+The network of football games is given in figure \ref{fig.cap:football_ex}.
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
+\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
+\end{figure}
   \subsection{Les Mis\'{e}rables}
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = lesmis$vertices, aes(esize = value), ealpha = 2/10, layout = 'fruchtermanreingold') +
-  expand_limits(x = c(-25,25), y = c(-25,25)) +
-  theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
 @
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
+\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
+\end{figure}
 \sct{
   \subsection{Protein Interaction Network in Yeast}
+  This example of a protein interaction network comes from \citet{protein}.
+It is the complete protein-protein interaction network in the yeast species \emph{S.
+cerevisiae}.
+There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.
+These edges represent ``direct physical interactions" between any 2 proteins \citep[][p.
+42]{protein}.
 }
@@
   geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
-  expand_limits(x = c(-25,25), y = c(-25,25)) + theme_net
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
 @
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
+\caption{\label{fig.cap:yeast_ex} Complete protein-protein interaction network in \emph{S. cerevisiae}.}
+\end{figure}
 

### 2015-03-24 CC/writeup.Rnw
add pdf

@@
 %\definecolor{Blue}{rgb}{0,0,0.5}
+\usepackage{color}
 \definecolor{purple}{rgb}{.4,0,.8}
@@
 library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
+opts_chunk$set(tidy = FALSE, echo=FALSE, cache=TRUE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
 @
-<<readingdata, eval=TRUE>>=
+<<readingdata, eval=TRUE, echo=FALSE>>=
  library(ggplot2) # needs to be the version with geom_net!
@@
 \includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
-\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romatic relationship.}
+\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romantic relationship.}
 \end{figure}
@@
 ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = I(2), vlabel = TRUE,aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(value)])) +
+  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 5, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
   expand_limits(x = c(0,1), y = c(0,1)) +

### 2015-03-28 CC/writeup.Rnw
small changes

@@
 <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 5, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
+ggplot(data = football$edges, aes(from_id = from, to_id = to, elinetype = in.conf + 1)) +
+  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 2, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
   expand_limits(x = c(0,1), y = c(0,1)) +
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
 @

### 2015-03-28 CC/writeup.Rnw
more small changes.

@@
 library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, cache=TRUE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
+opts_chunk$set(tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
 @
 <<readingdata, eval=TRUE, echo=FALSE>>=
+setwd("~/Dropbox/ggplot2/") 
 library(ggplot2) # needs to be the version with geom_net!
@@
  data(blood, football, theme_elements, lesmis)
+names(blood$vertices)[1] <- 'label'
 load("data/madmen.rda") # relative path should work for both of us
-names(madmen$vertices)[1] <- "label"
-names(blood$vertices)[1] <- "label"
-names(theme_elements$vertices)[1] <- "label"
+names(madmen$vertices)[1] <- 'label'
+names(theme_elements$vertices)[1] <- 'label'
 football$vertices$label <- as.factor(football$vertices$label)
@@
 
-This network is shown in figure~\ref{fig.cap:madmen_ex}. \emph{Notes: add names, color for gender}
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men',cache=TRUE>>=
+This network is shown in figure~\ref{fig.cap:madmen_ex}.
+<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
 ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
@@
            aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.5), y = c(0,1)) + theme_net
+  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net
 @
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-\sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times: approximately 3454 downloads per day.
+\sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
 It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.}
@@
   In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.
-The \texttt{vlabel} aesthetic is the name or number assigned to each vertex in order to uniquely identify it.
-This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be readable by the human eye.
+The \texttt{vlabel} aesthetic is a logical value that prints the name or number assigned to each vertex in order to uniquely identify it on the plot.
+\sct{If this value is set to TRUE, the vertex data frame is required to have a column titled label.} This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be read clearly.
 The \texttt{vcolor} aesthetic can be an identity color (e.g.
-\texttt{color = "red"} outside of the \texttt{aes} statement) to change the color of all vertices, or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.
+\texttt{color = "red"}) to change the color of all vertices, \sct{XXX shouldn't we take this out?
+or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.}  \sct{Right now, the \texttt{vcolor} aesthetic does not work like a typical color aesthetic in \texttt{ggplot2}.
+It can, however, take HEX color characters and assign them to different levels of a factor variable.}  \sct{XXX We should probably take this out too.
 It can also color the vertices according to the different values of  some numeric variable, such as degree.
-The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.
+The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.}
   %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
   The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs.
-\footnote{Sam's questions/notes: Is ``degree" going to be a column in our data frame?
-Or will we have to write something to detect degree size?} It can also be set to any other numeric variable associated with the vertices in the network.
+\sct{These must be integer-valued columns in the vertex data frame.} It can also be set to any other numeric variable associated with the vertices in the network.
 The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.
-It can also change the shape of the vertices according to some grouping variable.
+It can also change the shape of the vertices according to some grouping variable, \sct{which is currently required to consist of integer values from 0 to 25.}.
 Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10, or to some numeric variable of interest that takes on the values $[0,1]$.
@@
   Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for lines.
+\sct{XXX take out this???
 The \texttt{elabel} aesthetic is the name or number assigned to each edge.
-This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.
-The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to a grouping variable.
-It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge.
-The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the many different types of dotted lines in \texttt{ggplot2}.
-The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.
-The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability.
+This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.}  The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to HEX colors assigned to levels of a grouping variable.
+%It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge.
+  The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the different types of lines in \texttt{ggplot2}.
+\sct{The column associated with this aesthetic must be integer valued and take on values from one to six.}  The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.
+\sct{The associated column can take on any numerical values.}  The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability.
+\sct{This property is especially useul for networks with a lot of connections and with vertices of relatively high degree.}  \sct{XXX should we take this out?
 Finally, the \texttt{directed} aesthetic is a logical value that identifies whether or not the network is directed.
 When this value is true, the arrows change size, color, etc.
-with the previous aesthetics.
+with the previous aesthetics.}
 
@@
 \sctq{Can I cite wikipedia here?
-Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a persone whose blood type is that of a corresponding \emph{to} vertex.
+Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a person whose blood type is that of a corresponding \emph{to} vertex.
 In the example below, loops are removed.
@@
 ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2)) +
+  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2), directed = TRUE) +
   expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add direction later
@@
 ggplot(data = email.net, aes(from_id = From, to_id = to)) +
-  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
+  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
   expand_limits(x=c(0,1), y = c(0,1)) +
@@
 ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
-  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2)) +
+  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
   expand_limits(x=c(0,1), y = c(0,1)) +
@@
 \sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.
-With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving?
+\sctq{XXX note for later: remove all the emails that go to everyone in the network.} With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving?
 will know with direction option added later} several emails each day to a majority of the company.
@@
 ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices, vlabel = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) +
-  theme_net #add vlabel & direction later
+  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
+  expand_limits(x = c(0,1.2), y = c(0,1)) +
+  theme_net
 @
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) + geom_net(vertices = lesmis$vertices, ealpha = 7/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
 @

### 2015-03-29 CC/writeup.Rnw
added description about the layer. compiled after removing unneccesary print statements.

@@
 <<readingdata, eval=TRUE, echo=FALSE>>=
-setwd("~/Dropbox/ggplot2/") 
 library(ggplot2) # needs to be the version with geom_net!
@@
 Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function.
-\emph{Finally, we use the \texttt{GeomText} draw function to add the vertex and edge labels when the \texttt{vlabel} and \texttt{elabel} arguments, respectively, are true.
-*insert more detail here after figuring it out*}
+\emph{Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE.
+We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default.
+We then created a new data frame for the label layer.
+This allows the vertex properties to be propogated through to the labels, including size and color.}
 } \sctq{This doesn't seem very long, but I already feel like there's too much detail! Should I add more?}

### 2015-03-29 CC/writeup.Rnw
couple of comments

@@
 \usepackage{url}
+\usepackage{afterpage}
 \usepackage{pdflscape}
@@
 \end{figure}
+\afterpage{\clearpage}
 
@@
 @
+
 \sct{
@@
 \end{landscape}
+
 \sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.
@@
 \end{figure}
+\afterpage{\clearpage}
+
   \subsection{Les Mis\'{e}rables}
@@
 \end{figure}
+\afterpage{\clearpage}
+
 \sct{
@@
 I know you'll tell me if I'm too detailed or not detailed enough!}
+\hh{You're doing just fine, keep going!}
 \sct{
@@
   \subsection{Creating \texttt{geom\_net}}
+\hh{mention that a network is constructed from different layers - the node layer of points, an edge layer of line segments and an optional text layer of labels. }
+
   We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
@@
 Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function.
-\emph{Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE.
+Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE.
 We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default.
 We then created a new data frame for the label layer.
-This allows the vertex properties to be propogated through to the labels, including size and color.}
-} \sctq{This doesn't seem very long, but I already feel like there's too much detail! Should I add more?}
+This allows the vertex properties to be propogated through to the labels, including size and color.
 

### 2015-03-29 CC/writeup.Rnw
fixed football example

--- 2015-03-29 CC/writeup.Rnw
@@
 football$vertices$value <- as.factor(football$vertices$value)
+football$edges$in.conf <- 0
+for (i in 1:nrow(football$edges)){
+  idx1 <- which(football$vertices$label == football$edges$from[i])
+  idx2 <- which(football$vertices$label == football$edges$to[i])
+  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
+    football$edges$in.conf[i] <- 1
+  }
+}
 email.net <- read.csv("data/email-data.csv")

### 2015-03-30 CC/writeup.Rnw
fix ex.
changed email example ... I like it!

@@
 <<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to)) +
-  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
+library(plyr)
+email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
+employee <- read.csv("data/EmployeeRecords.csv")
+employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
+  geom_net(vertices = employee,ealpha = 0.1, vsize = I(3), directed = TRUE, aes(vcolour=as.numeric(CurrentEmploymentType))) +
   expand_limits(x=c(0,1), y = c(0,1)) +
@@
 We then created a new data frame for the label layer.
-This allows the vertex properties to be propogated through to the labels, including size and color.
+This allows the vertex properties to be propogated through to the labels, including size and color.}
 

### 2015-03-30 CC/writeup.Rnw
bit cleaned up

--- 2015-03-29 CC/writeup.Rnw
@@
 email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-
+library(plyr)
+email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
+employee <- read.csv("data/EmployeeRecords.csv")
+employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+library(RColorBrewer)
+cols <- brewer.pal(8, name="Set1")
 @
@@
 <<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to)) +
-  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
+  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
+           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
   expand_limits(x=c(0,1), y = c(0,1)) +
-  theme_net #add direction later
+  theme_net
 @

### 2015-03-30 CC/writeup.Rnw
email example facetting

@@
 <<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
-ggplot(data = email.net, aes(from_id = From, to_id = to, )) +
-  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to, )) +
+  geom_net(vertices = employee,ealpha = 0.1, vsize = I(2), directed = TRUE,
+           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
   expand_limits(x=c(0,1), y = c(0,1)) +
   facet_wrap(~day, nrow = 2) +
-  theme_net #add direction later
+  theme_net
 @

### 2015-03-30 CC/writeup.Rnw
couple of comments

@@
 These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study.
-\sct{For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.
+For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.
 Additionally, \citet{networkfailures} examine node failure in interdependent networks like power grids.
 Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C.
-elegans} worm are also etensively studied \citep{smallworld}.}  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used \sct{and most complex} networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
+elegans} worm are also etensively studied \citep{smallworld}.
+Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used and most complex networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
 The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
-The world wide web is an example of a directed network because one webpage \sct{may} link to another, and not necessarily the other way around.
+The world wide web is an example of a directed network because one webpage may link to another, and not necessarily the other way around.
 Undirected edges, however, are simply connections between vertices.
@@
 \par A social network is a network that everyone is a part of in one way or another.
-We do not \sct{necessarily} refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
+We do not necessarily refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
 To demonstrate the functionality of our geometry for plotting networks, we have chosen an example of a social network from the popular television show Mad Men.
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-\sct{From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
-It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.}
+\hh{Just to give an idea of the popularity and the wide=spread use of \texttt{ggplot2}:} From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
+It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.
+\hh{This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.}
 
@@
   \subsection{Data Structure}
-  \hh{XXX we need a link between the two data frames: we have from\_id and to\_id in the edge data frame, but we also need an 'id' variable in the vertex data.
-We don't officially have that at the moment - in the code there is a soft requirement (new as of Mar 20) that one of the columns in the edge data set has to be called 'label' ...
-that's a bit stoopid, but I couldn't think of a better name.
-It would make more sense to call this id or edgeid or something along those lines - let's talk about it, and make it a formal requirement.}
+  \hh{variable label links the two data sets - let's talk about it, and make it a formal requirement.}
   In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
@@
   \subsection{Vertex Aesthetics}
+
   In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.
@@
   \subsection{Madmen Network (again?? or no?)}
-  \sctq{leaving blank for now.}
+  \sctq{leaving blank for now.} \hh{i don't think we need to repeat it, but you could show the code for the example here and give the reference to the example.
+are there any more details you wanted to talk about?}
   \subsection{College Football}
-  \emph{Note: would be best with coloring vertices by conference}
+
   This next example comes from M.E.J.
@@
 @
-\sct{
 The network of football games is given in figure \ref{fig.cap:football_ex}.
@@
 The edges are also weighted by the number of coappearances.
-The largest weighting is 31, between the characters Jean Valjean and Cosette.}
+The largest weighting is 31, between the characters Jean Valjean and Cosette.
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) + geom_net(vertices = lesmis$vertices, ealpha = 7/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) +
+  geom_net(vertices = lesmis$vertices, ealpha = 7/10,
+           layout = 'fruchtermanreingold', vlabel=TRUE) +
+  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
 @
@@
 \includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
-\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
+\caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
 \end{figure}
@@
 
-\sct{
   \subsection{Protein Interaction Network in Yeast}
@@
 42]{protein}.
-}
 <<read_data>>=
@@
 \hh{You're doing just fine, keep going!}
+
+\hh{XXX have a look at the ggplot2 help online and include something along these lines for both stat_net and geom_net}
+
 \sct{

### 2015-03-30 CC/writeup.Rnw
email example

@@
 }
-email.net <- read.csv("data/email-data.csv")
-email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-library(plyr)
-email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-employee <- read.csv("data/EmployeeRecords.csv")
-employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+load("data/email.rda")
+#email.net <- read.csv("data/email-data.csv")
+#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
+#library(plyr)
+#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
+#employee <- read.csv("data/EmployeeRecords.csv")
+#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+email.net <- email$edges
+employee <- email$nodes
 library(RColorBrewer)

### 2015-03-30 CC/writeup.Rnw
mini changes to bib file

--- 2015-03-30 CC/writeup.Rnw
@@
 load("data/madmen.rda") # relative path should work for both of us
-names(madmen$vertices)[1] <- 'label'
+#names(madmen$vertices)[1] <- 'label'
 names(theme_elements$vertices)[1] <- 'label'
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-\hh{Just to give an idea of the popularity and the wide=spread use of \texttt{ggplot2}:} From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
+\hh{Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}:} From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
 It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.

### 2015-03-31 CC/writeup.Rnw
many things added! lots of elaboration on examples, more detail about capabilities in the first section.

--- 2015-03-30 CC/writeup.Rnw
@@
  theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements, lesmis)
+ data(blood, football, theme_elements, lesmis,madmen, mm_directed,email)
 names(blood$vertices)[1] <- 'label'
-load("data/madmen.rda") # relative path should work for both of us
-names(madmen$vertices)[1] <- 'label'
+#load("data/madmen.rda") # relative path should work for both of us
+#load("data/mm_directed.rda")
+#names(madmen$vertices)[1] <- 'label'
 names(theme_elements$vertices)[1] <- 'label'
@@
 }
-email.net <- read.csv("data/email-data.csv")
-email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-library(plyr)
-email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-employee <- read.csv("data/EmployeeRecords.csv")
-employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+#load("data/email.rda")
+#email.net <- read.csv("data/email-data.csv")
+#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
+#library(plyr)
+#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
+#employee <- read.csv("data/EmployeeRecords.csv")
+#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
+email.net <- email$edges
+employee <- email$nodes
 library(RColorBrewer)
 cols <- brewer.pal(8, name="Set1")
+yeast <- read.table("data/protein.txt")
+yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
 @
@@
 
- In the plot, we can see one central character who has many more relationships than any other character. This vertex represents the main character of the show, Don Draper.
+ In the plot, we can see one central character who has many more relationships than any other character.
+This vertex represents the main character of the show, Don Draper, who is quite the ``ladies' man."  This fun example shows just how ubiquitous networks are.
 \par There are many kinds of networks, and networks are extensively studied across many disciplines.
@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-\hh{Just to give an idea of the popularity and the wide=spread use of \texttt{ggplot2}:} From January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
+Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}, from January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
 It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.
-\hh{This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.}
+This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.
 
@@
   \subsection{Data Structure}
-  \hh{variable label links the two data sets - let's talk about it, and make it a formal requirement.}
 In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
 The vertex data frame should contain all the relevant vertex information.
-\emph{The only necessary variables are the vertex names/labels and the vertex degree (if undirected) or the vertex in degree and out degree (if directed).
-The names and degrees are stored in columns and the rows are the vertices.
-Other values of interest, such as grouping variables, for each vertex should be stored as columns in this dataframe, with an observation for each vertex.}
+The only necessary variable is one called \texttt{label} which contains the vertex labels as they are encoded in the edges data set.
+This will be a formal requirement in the final version of the \texttt{geom\_net} function.
+Each row represents one vertex.
+Other values of interest, such as grouping variables or degree of each vertex should be stored as columns in this dataframe, with an observation for each vertex.
   \par The edge data frame should contain all the relevant edge information.
 The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network.
-The From and To vertices should be listed according to the names of the vertices in the vertex information data frame.
+The From and To vertices should match the names of the vertices in the vertex information data frame.
+The formal names for these columns are \texttt{from\_id} and \texttt{to\_id}, respectively.
 Other variables may also be included for each edge, such as the edge weight or grouping variable.
-As before, the variables of interest are columns in the data frame and the rows are each edge in the network.%\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}
+As before, the variables of interest are columns in the data frame and the rows are each edge in the network.
 
@@
 
-  In our geometry, the possible vertex aesthetics mimic the usual aesthetics in \texttt{ggplot2} for points.
-The \texttt{vlabel} aesthetic is a logical value that prints the name or number assigned to each vertex in order to uniquely identify it on the plot.
-\sct{If this value is set to TRUE, the vertex data frame is required to have a column titled label.} This aesthetic is most useful for smaller network objects where all vertex names can be printed on or near their corresponding vertices and still be read clearly.
+  In our geometry, we want to create all possible vertex aesthetic to mimic the usual aesthetics in \texttt{ggplot2} for points.
 The \texttt{vcolor} aesthetic can be an identity color (e.g.
-\texttt{color = "red"}) to change the color of all vertices, \sct{XXX shouldn't we take this out?
-or it can take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.}  \sct{Right now, the \texttt{vcolor} aesthetic does not work like a typical color aesthetic in \texttt{ggplot2}.
-It can, however, take HEX color characters and assign them to different levels of a factor variable.}  \sct{XXX We should probably take this out too.
-It can also color the vertices according to the different values of  some numeric variable, such as degree.
-The colors are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.}
+\texttt{color = "red"}) to change the color of all vertices, or, in the final version, it will take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.
+In the current version, the \texttt{vcolor} aesthetic does not work like a typical color aesthetic in \texttt{ggplot2}.
+It can, however, take HEX color characters and assign them to different levels of a factor variable.
+We use the \texttt{RColorBrewer} package to choose the colors for us in this paper \citep{RColorBrewer}.
+It will also color the vertices along a color gradient according to the different values of some numeric variable such as degree.
+The vertex color defaults to black.
+% are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.}
   %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
-  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to the degree of the vertices in undirected graphs or the in-degree (or out-degree) of directed graphs.
-\sct{These must be integer-valued columns in the vertex data frame.} It can also be set to any other numeric variable associated with the vertices in the network.
+  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to one of any numerical columns in the vertex data frame.
+Right now, the numerical values are strict point values: if the size of a vertex is two, then it will be two points in diameter.
+A point is the size of the default fontsize in the \texttt{grid} package.
+In our final version, the \texttt{vsize} will scale to minimum and maximum values as all other size aesthestics in \texttt{ggplot2} do.
 The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.
-It can also change the shape of the vertices according to some grouping variable, \sct{which is currently required to consist of integer values from 0 to 25.}.
-Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10, or to some numeric variable of interest that takes on the values $[0,1]$.
+It also changes the shape of the vertices according to some grouping variable, which is currently required to consist of integer values from 0 to 25.
+In the final version, it will be able to change shape based on different levels of a factor variable.
+Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10 or 0.5.
+It operates in exactly the same way as the \texttt{alpha} aesthetic in \texttt{geom\_point}.
+This aesthetic is very useful for networks with hundreds or thousands of vertices, which can easily crowd the static visualization.
+Making the vertices more transparent will better show the underlying structure in the network.
 
@@
   \subsection{Edge Aesthetics}
-  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for lines.
-\sct{XXX take out this???
-The \texttt{elabel} aesthetic is the name or number assigned to each edge.
-This aesthetic is most useful for visualizing random graphs or Markov processes, where each edge probability is of interest.}  The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to HEX colors assigned to levels of a grouping variable.
-%It can also be used to color the edges in random graphs or Markov processes according to the probability of each edge.
+  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for segments.
+The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to HEX colors assigned to levels of a grouping variable.
+In the final version, the \texttt{ecolor} aesthetic will automatically color according to levels of a grouping variable or according to a gradient scale associated with a numeric variable.
   The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the different types of lines in \texttt{ggplot2}.
-\sct{The column associated with this aesthetic must be integer valued and take on values from one to six.}  The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.
-\sct{The associated column can take on any numerical values.}  The \texttt{ealpha} aesthetic can change the edge transparency to a set value, or to some numeric variable of interest that takes on the values $[0,1]$, such as edge weight or probability.
-\sct{This property is especially useul for networks with a lot of connections and with vertices of relatively high degree.}  \sct{XXX should we take this out?
-Finally, the \texttt{directed} aesthetic is a logical value that identifies whether or not the network is directed.
-When this value is true, the arrows change size, color, etc.
-with the previous aesthetics.}
+Currently, the column associated with this aesthetic must be integer valued and take on values from zero to six.
+In the final version, it will change the edge linetypes according to levels of a grouping variable.
+The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.
+The associated column can take on any numerical values, and currently maps the values directly to points, just like the \texttt{vsize} aesthetic.
+In the final version, this aesthetic will work exactly like the sizing in \texttt{ggplot2}, by scaling sizes to the minimum and maximum values.
+The \texttt{ealpha} aesthetic can change the edge transparency to a set value on the interval $[0,1]$.
+This property is especially useul for large networks with a lot of connections or with vertices of relatively high degree.
+Finally, we will add an \texttt{elabel} aesthetic to print the name or number assigned to each edge on or next to it.
+This aesthetic will be very useful for visualizing random graphs or Markov processes where each edge probability is of interest.
 
-  \subsection{Layout Argument}
-  \sct{There are many layouts to choose from, but the default layout is the Kamada-Kawai layout.  This is a force-directed layout for undirected networks \citep{kamadakawai}}.
-  %and the default layout changes according the size of the graph.
-For smaller networks, with less than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE.
-For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE.
-%**Now justify these two statements with research to be done.** \hh{I don't think that you'd have to defend your choices of defaults.
-They are defaults, and can be changed.
-}}
-  The \texttt{layout} \sct{argument in the \texttt{geom\_net} function} can be changed to any of the layouts also found in the \texttt{sna} package, such as circular, eigen, or Fruchterman-Reingold.
-\sctq{Should I insert a table of layouts here?
-And I took out the thing about the background color because I figure it can just be changed with \texttt{theme()}} %Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.
-This aesthetic takes only the identity colors, not additional variables.%\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.
-I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.
-\hh{You're fine - those are things, which people are going to have strong opinions on without actual backup.
-You provide the environment for people to easily use network data in a familiar setting.}}
+  \subsection{Other Arguments}
+  Similar to the other geometries and functions in \texttt{ggplot2}, our geometry takes several arguments outside of the aesthetic mapping parameters.
+One such argument is the \texttt{vlabel} argument.
+The \texttt{vlabel} aesthetic is a logical value that, if set to \texttt{TRUE} prints the name or number associated with each vertex in the \texttt{label} column in order to identify it on the plot.
+In the final version of this geometry, we will add capabilities to adjust the size and positioning of the labels.
+For the moment, the color is set to the vertex color, the size is set to twice the vertex size, and all other possible parameters are set to their defaults in \texttt{geom\_text()}.
+The \texttt{vlabel} argument is most useful for smaller network objects where all vertex names can be printed on their corresponding vertices and still be read clearly.
+Next, the \texttt{directed} argument is a logical value that identifies whether or not the network is directed.
+When this value is true, arrows are created on the end of the line segment that corresponds to the \texttt{to\_id} value for each edge.
+The arrows match the edges in coloring, linetype, and size, and the default length of the arrow is set using the \texttt{unit} function in the \texttt{grid} package to 0.015 of the normalised parent coordinates of the plot.
+  Finally, the \texttt{layout} argument takes a character value corresponding to the possible layouts in the \texttt{sna} package that are created by the \texttt{gplot.layout.*()} family of functions.
+The default layout is the Kamada-Kawai layout.
+This is a force-directed layout for undirected networks \citep{kamadakawai}.
+There are, however, many other layouts possible.
+\begin{table}
+\centering
+\begin{tabular}{l|p{10cm}}
+Layout Name & Description \\
+\hline \hline
+\texttt{"adj"} & a version of \texttt{"mds"} which scales the raw adjacency matrix \\
+\texttt{"circle"}* & places vertices uniformly in a circle \\
+\texttt{"circrand"} & places vertices randomly in a circle \\ 
+\texttt{"eigen"} & places vertices based on the eigenstructure of the adjacency matrix \\
+\texttt{"fruchermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
+\texttt{"geodist"} & a version of \texttt{"mds"} which scales the matrix of geodesic distances \\
+\texttt{"hall"}* &  places vertices based on the last two eigenvectors of the Laplacian of the input matrix \\
+\texttt{"kamadakawai"} & generates a vertex layout using a version of the Kamada-Kawai force-directed placement algorithm \\
+\texttt{"mds"}  & places vertices based on a metric multidimensional scaling of a specified distance matrix \\
+\texttt{"random"} & places vertices randomly \\
+\texttt{"segeo"} & a version of \texttt{"mds"} which scales the squared euclidean distances between row-wise geodesic distances \\
+\texttt{"seham"} & a version of \texttt{"mds"} which scales the Hamming distance between rows/columns of the adjacency matrix \\
+\texttt{"target"} & produces a ``target diagram" or ``bullseye" layout using a force-directed placement algorithm.
+\end{tabular}
+\caption{All possible layouts to pass to \texttt{geom\_net}.
+All descriptions are from \citet{sna}.
+Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.
+All others use the default layout parameters in \texttt{sna} for now.}
+\label{table:layouts}
+\end{table}
+
+All layouts that \texttt{geom\_net} is currently capable of graphing are listed in table \ref{table:layouts}.
+There is not currently functionality to add layout parameters to the layout function within \texttt{geom\_net}, though this will be included in the final version.
+Therefore, all functions in table \ref{table:layouts} that are not marked with asterisks are using their respective default parameters according to the \texttt{sna} documentation for \texttt{gplot.layout}.
 
 \section{Many Many Examples}
-\sctq{Format-wise, is it ok if each example is its own subsection?
-Also, can you check the references I added to the .bib file to make sure I'm creating them correctly?
-I've just been appending them.}\\
-\sct{
-In this section, we demonstrate the capabilities of \texttt{geom\_net} in a series of examples.
+In this section, we demonstrate the current capabilities of \texttt{geom\_net} in a series of diverse examples.
   \subsection{Blood Donation}
-  \emph{Note: would be best with labels and direction}\\
   In this directed network, there are eight vertices and 27 edges.
-The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive and negative types.
-\sctq{Can I cite wikipedia here?
-Should I find a more academic source?}  The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a person whose blood type is that of a corresponding \emph{to} vertex.
-In the example below, loops are removed.
-Loops exist on every vertex in this example, as blood of matching ABO and RhD type can always be received.
-}
+The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive (+) and negative (-) types.
+The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a person whose blood type is that of a corresponding \emph{to} vertex.
+In the example below, loops are removed because loops exist on every vertex in this example, as blood between two people of matching ABO and RhD type can always be exchanged.
 <<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
 ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, layout='circle', vlabel = TRUE, vsize = I(2), directed = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net #add direction later
+  geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
+           vlabel = TRUE, vsize = I(3), directed = TRUE) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
 @
-
-\sct{
 This network is shown in figure \ref{fig.cap:blood_ex}.
+Here, we have used the aesthetics \texttt{vcolour} and \texttt{vsize} set to identity values to change the size and color of all vertices.
+We have also used the \texttt{layout} and \texttt{vlabel} arguments to change the default layout to a circle layout and to print the blood types, respectively.
+The circle layout places blood types of the same ABO type next to each other and spreads the vertices out far enough to distinguish between the various ``in" and ``out" types.
+You can tell clearly from this plot that the O- type is the universal donor: it has out degree of seven and in degree of zero.
+Additionally, you can see that the AB+ type is the universal recepient, with in degree of seven and out degree of zero.
+Anyone looking at this plot can quickly determine which type(s) of blood they can receive and which type(s) can receive their blood.
 \begin{figure}
@@
 \subsection{Email Network}
-  \emph{Note: would be best with coloring and direction}
   This email network comes from the 2014 VAST Challenge \citep{emailnet}.
@@
 The network contains two business weeks of emails across the entire company.
-}
-<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
+In order to better visualize the structure of the communication network between employees, all emails that were sent out to all employees are removed in the subsequent examples.
+<<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE>>=
 ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
@@
            aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) +
-  theme_net
+  expand_limits(x=c(0,1), y = c(0,1)) + theme_net
 @
@@
 
-\sct{This network is plotted in figure \ref{fig.cap:email_ex}.
-This plot is very crowded, and it is difficult to distinguish structure in the network when it there is so much over-plotting.
-To remedy this, we facet the network by day to examine how the company's communications change throughout the week.
-}
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.'>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to, )) +
-  geom_net(vertices = employee,ealpha = 0.1, vsize = I(2), directed = TRUE,
+This network is plotted in figure \ref{fig.cap:email_ex}.
+There are six distinct clusters in this network which almost perfectly correspond to the six different types of employee in this company: administration, engineering, executive, facilities, information technology, and security.
+Unfortunately, there is currently no legend to associate color with employment type.
+This will be remedied in the final version of our geometry.
+Additionally, the edges between employees in the same cluster are much darker than edges between employees in different clusters.
+This is due to the value of the \texttt{ealpha} aesthetic: more emails between two employees lead to darker edges.
+The value is set to 0.1 in this example, so that 10 or more emails between 2 employees results in a completely opaque edge.
+This pattern of heavy communication between employees of the same type is fairly unsurprising.
+To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
+<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE>>=
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) + 
+  geom_net(vertices = employee, ealpha = 0.2, vsize = I(2), directed = TRUE, 
            aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) +
-  facet_wrap(~day, nrow = 2) +
-  theme_net
+  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net
 @
@@
 \end{landscape}
-
-\sct{This plot is shown in figure \ref{fig.cap: email_ex_facet}.
-\sctq{XXX note for later: remove all the emails that go to everyone in the network.} With the facetting, we can see that there are between four and seven employees sending \sctq{or receiving?
-will know with direction option added later} several emails each day to a majority of the company.
-We can also see clusters that have formed on the perimeter of each day's network where the employees are communicating heavily with each other, but not much outside of their cluster.
-We can also see that there is heavier inter-cluster communication on different days.
-For example, day 17 has more concentrated in-group communication than day 15.
-Additionally, we can better distinguish the communication of the outlying employee, who only corresponds with up to four other complany employees on a given day.} \sctq{talk about clusters being different departments within the company?}
+This plot is shown in figure \ref{fig.cap: email_ex_facet}.
+With the facetting, we can see that there are several days where one or more departments do not communicate with any of the other departments.
+There are only two days, 13 and 15, without any isolated department communications.
+Facetting is one of the major benefits of creating a geometry for networks in \texttt{ggplot2}.
+Facetting quickly separates dense networks into separate subnetworks for easy visual comparison and  analyses.
   \subsection{\texttt{ggplot2} Theme Elements}
-  \emph{Note: would be best with labels and direction} \\
-  \sct{This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
+This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
 It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.
 There are 53 vertices and 36 edges in this network.
-There is an arrow from one vertex to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.
+Each vertex represents one possible theme option.
+There is an arrow from one theme option to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.
 For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.
-Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and thus no lines appear anywhere on the plot background.
+Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and no lines appear anywhere on the plot background.
 Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.
-}
 <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
@@
 
-\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
-  \subsection{Madmen Network (again?? or no?)}
-  \sctq{leaving blank for now.} \hh{i don't think we need to repeat it, but you could show the code for the example here and give the reference to the example.
-are there any more details you wanted to talk about?}
+The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
+In this plot, it is easy to quickly determine the parent and child vertices.
+Using this plot made creation of the \texttt{theme\_net} object used throughout these examples very simple.
+We just made set each of the major parent elements, \texttt{text}, \texttt{rect}, and \texttt{line} to \texttt{element\_blank()} and then set the aspect ratio equal to one.
+  \subsection{Mad Men Networks}
+  The following code creates the network example given in the introduction.
+We changed the vertex size and edge color for all vertices and edges, included vertex labels, and colored the vertices according to the character's gender.
+<<print_ex, ref.label="madmen_ex", eval=FALSE, echo=TRUE,cache=FALSE>>=
+@
+ There is another Mad Men network included in the \texttt{gcookbook} package \citep{madmen}.
+It is a directed network, also of romantic relationships between characters, but it also includes advances made by one character that were rejected by the other.
+For example, Roger Sterling made advances toward Betty Draper, but Betty refused him, and so there is a directed edge going from Roger to Betty, but not from Betty to Roger.
+If the advance is reciprocated, like between Sal Romano and the Bellhop, there are two directed edges between the two vertices.
+<<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
+ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) + 
+  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T, 
+           vsize = I(2.5), layout = 'fruchtermanreingold',
+           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) + 
+  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+@
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/madmen2_ex-1.pdf}
+\caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
+\end{figure}
+This network is shown in figure \ref{fig.cap:madmen2_ex}.
+This network is much more connected than the previous Mad Men example.
+It also allows us to see much more of the drama in the show.
+For instance, you can see that Roger made advances towards Betty, his business partner's wife, which is much more scandalous than what we can see in the first network.
+
   \subsection{College Football}
-
   This next example comes from M.E.J.
@@
 It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.
-There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
+There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.
+There is an additional variable in the vertex data frame corresponding to the conference each team belongs to, and there is an additional variable in the edge data frame that is equal to one if the game occured between teams in the same conference or zero if the game occured between teams in different conferences.
 <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-ggplot(data = football$edges, aes(from_id = from, to_id = to, elinetype = in.conf + 1)) +
-  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 2, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
-  expand_limits(x = c(0,1), y = c(0,1)) +
-  theme_net
+colors <- brewer.pal(12, name = 'Paired')
+ggplot(data = football$edges, aes(from_id = from, to_id = to, 
+                                  elinetype = in.conf + 1)) +
+  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE, 
+           aes(vcolour=colors[as.numeric(factor(value))])) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
 @
 The network of football games is given in figure \ref{fig.cap:football_ex}.
+Here, we have changed the \texttt{elinetype} aesthetic to correspond to games that occur between teams in the same conference or different conferences.
+These lines are dotted and solid, respectively.
+We have also assigned a different color to each conference, and the vertices and their labels are colored according to their conference.
+This coloring and changing of the linetypes make the structure of the game network easier to view.
+There is one conference consisting of Navy, Notre Dame, Utah State, Central Florida, and Connecticut, which is spread out all over the network, whereas most other conferences' teams are all very close to each other because they play within conference much more than they play out of conference.
+At the time, these five schools were all independents and did not have a home conference.
+Without the coloring capability, we would not have been able to pick out that difference as easily.
 \begin{figure}
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) +
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, 
+                                esize = value/mean(value))) +
   geom_net(vertices = lesmis$vertices, ealpha = 7/10,
@@
 42]{protein}.
-<<read_data>>=
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
-@
+
 <<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
@@
 
-\hh{XXX have a look at the ggplot2 help online and include something along these lines for both stat_net and geom_net}
+\hh{XXX have a look at the ggplot2 help online and include something along these lines for both stat\_net and geom\_net}
 

### 2015-04-01 CC/writeup.Rnw
done for today. i still have a few things to fill in near the end, and i lef some comments/questions

@@
 \begin{abstract}
-Abstract will be here
+There are many implementations of static network visualization in R, but none are equipped with the flexibility and functionality of \texttt{ggplot2}.
+\texttt{geom\_net} was created to fill this gap.
+Using two data frames to describe vertex and edge information, \texttt{geom\_net} makes use of the underlying structure of \texttt{ggplot2} to visualize networks.
+This makes it possible to easily facet networks according to covariates  or change aesthetics such as shape, size, or color according to additional edge or vertex information.
 \end{abstract}
@@
 Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}, from January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
-It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times.
+It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times\footnote{\texttt{ggplot2} usage statistics taken from \url{http://cran-logs.rstudio.com/}.}.
 This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.
@@
 In the final version, this aesthetic will work exactly like the sizing in \texttt{ggplot2}, by scaling sizes to the minimum and maximum values.
+This is one of two ways to visualize the number of edges between two vertices.
+The other way is through the \texttt{ealpha} aesthetic.
 The \texttt{ealpha} aesthetic can change the edge transparency to a set value on the interval $[0,1]$.
-This property is especially useul for large networks with a lot of connections or with vertices of relatively high degree.
+If there are multiple edge connections given between two vertices in the edge data frame, the \texttt{ealpha} aesthetic can be changed to plot opaque lines between vertices with many mutual connections and varying degrees of transparent lines between vertices with only a few or a single connection.
+This property is also especially useful for large networks with a lot of connections or with vertices of relatively high degree.
 Finally, we will add an \texttt{elabel} aesthetic to print the name or number assigned to each edge on or next to it.
@@
 All descriptions are from \citet{sna}.
-Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.
-All others use the default layout parameters in \texttt{sna} for now.}
+Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.}
 \label{table:layouts}
@@
 All layouts that \texttt{geom\_net} is currently capable of graphing are listed in table \ref{table:layouts}.
-There is not currently functionality to add layout parameters to the layout function within \texttt{geom\_net}, though this will be included in the final version.
-Therefore, all functions in table \ref{table:layouts} that are not marked with asterisks are using their respective default parameters according to the \texttt{sna} documentation for \texttt{gplot.layout}.
+The layouts not marked with an asterisk can also take their associated layout parameters.
+The \texttt{layout.par} argument takes a list of named parameters.
+For instance, if \texttt{layout = "random"}, we can also set \texttt{layout.par = list(dist = "normal")} to change the distribution of vertex placement from the default uniform to Gaussian.
 
@@
   \subsection{Les Mis\'{e}rables}
-  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
   This next network comes from \citet{lesmis}.
@@
 The largest weighting is 31, between the characters Jean Valjean and Cosette.
+This network is shown in figure \ref{fig.cap:lesmis_ex}.
+This network reveals many important character groups, as well as the main characters in the story.
+The main characters are the ones with high degree, like Valjean, Cosette, and Javert.
+There is an extremely wide edge connecting Valjean and Cosette, which implies that they are strongly connected; in fact, Cosette is Valjean's adopted daughter.
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
@@
   geom_net(vertices = lesmis$vertices, ealpha = 7/10, 
+           vcolour = I("red"), vsize = I(2.5),
            layout = 'fruchtermanreingold', vlabel=TRUE) +
@@
 42]{protein}.
+Because the network is very large, we randomly sampled 500 interactions and their associated proteins to plot it in figure \ref{fig.cap:yeast_ex}.
+We also demonstrate the layout capabilities by changing the layout to random and setting the distribution to \texttt{"uniang"}, which is a  ``gaussian donut" layout.
+Indeed, we see a nearly round area in the middle of the graph where the layout parameter has forced there to be no vertices.
 
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
-ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
+<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
+yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
+ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
+  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, vcolour = 'magenta',
+           layout = 'random', layout.par = list(dist = 'uniang')) +
   expand_limits(x = c(0,1), y = c(0,1)) + theme_net
@@
 \includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-\caption{\label{fig.cap:yeast_ex} Complete protein-protein interaction network in \emph{S. cerevisiae}.}
+\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
 \end{figure}
@@
 \section{Technical Details}
-\sctq{I have no idea exactly how detailed I should be.
-I'm going to attempt to be detailed without parsing each line in the code, but I do want to describe how everything works in detailed enough descriptions.
-I know you'll tell me if I'm too detailed or not detailed enough!}
-\hh{You're doing just fine, keep going!}
 
-\hh{XXX have a look at the ggplot2 help online and include something along these lines for both stat\_net and geom\_net}
-
-\sct{
 In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
@@
 These two data frames, along with the many \texttt{aes()} and other argument options get passed on to the actual network geometry for drawing.
+   \subsection{\texttt{stat\_net} Usage}
+<<usage_stat_net, eval=FALSE>>=
+stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", na.rm = FALSE, ...)
+#aes(colour = "black", size =3, layout="kamadakawai",layout.par = NA)
+@
+  The usage of the \texttt{stat\_net} arguments is as follows:
+  \sct{XXX adding detail tomorrow}
+  \begin{itemize}
+  \item mapping
+  \item data
+  \item vertices
+  \item geom
+  \item position
+  \item na.rm
+  \end{itemize}
+  \texttt{stat\_net} also accepts the following aesthetic values, none of which are required:
+  \begin{itemize}
+  \item layout 
+  \item layout.par
+  \end{itemize}
+
   \subsection{Creating \texttt{geom\_net}}
-\hh{mention that a network is constructed from different layers - the node layer of points, an edge layer of line segments and an optional text layer of labels. }
-
   We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
 Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.
-We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function.
-Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function.
+We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function to add a layer of segments to the plot.
+Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function to construct a point layer for the vertices.
 Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE.
+This adds a layer of text on top of the other two layers.
 We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default.
 We then created a new data frame for the label layer.
-This allows the vertex properties to be propogated through to the labels, including size and color.}
+This allows the vertex properties to be propogated through to the labels, including size and color.
+Finally, we added another if/else statement for the \texttt{directed} argument.
+This adds arrow arguments to the edges data frame which will put arrows pointing toward the ``to" vertex in a directed network corresponding to the \texttt{to\_id} value in the edges data frame.
+
+   \subsection{\texttt{geom\_net} Usage}
+<<usage_geom_net, eval=FALSE>>=
+geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
+#aes(ecolour = "grey20", vcolour = "black", vfill="black", fill = NA, esize = 1, vsize = 2, elinetype = 1, weight = NA, valpha = 1, vshape = 16, ealpha=1)
+@ 
+  The usage of the \texttt{geom\_net} arguments is as follows:
+  \sct{XXX adding detail tomorrow}
+  \begin{itemize}
+  \item mapping
+  \item data
+  \item vertices
+  \item vlabel
+  \item directed
+  \item stat
+  \end{itemize}
+  \texttt{stat\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
+  \begin{itemize}
+  \item from\_id*
+  \item to\_id*
+  \item ecolour
+  \item esize
+  \item elinetype
+  \item ealpha
+  \item vcolour
+  \item vfill
+  \item vsize
+  \item vshape
+  \item valpha
+  \end{itemize}
+
+\section{Future Work}
+As we discussed throughout the paper, there is still some work to be done before we submit it for publication in \emph{The R Journal}.
+We also hope to make it an official part of \texttt{ggplot2}.
+We believe that this is an important contribution to \texttt{ggplot2} and to the R user community because it fills in gaps in the network plotting capabilities of R.
+\sct{XXX That's all I got.
+Help?}
 

### 2015-04-01 madmenexample.R

@@ -8,3 +8,22 @@ madmen <- list(edges = madmen, vertices = mm.vert)
 
 ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + 
   expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net
+
+data(madmen2, package = "gcookbook")
+head(madmen2)
+
+label <- unique(c(madmen2$Name1,madmen2$Name2))
+Gender <- c('m','f',rep('m',2), rep('f',3), 'm', 'f', 'm','f',rep('m',2), 'f', rep('m',2), 'f', rep('m',2), rep('f',9),
+            rep('m',3), rep('f',4), 'm', 'f','m',rep('f',2), 'm', 'f','m','m','f','m',rep('f',6))
+
+vert = data.frame(label,Gender)
+head(vert)
+
+mm.directed <- list(edges = madmen2, vertices = vert)
+aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])
+
+ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) + 
+  geom_net(vertices = mm.directed$vertices, layout= "random", l.param = list(dist = 'normal'),directed = T, vlabel = T, vsize = I(2.5), aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) + 
+  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+
+save(mm.directed, file = "~/Dropbox/ggplot2/data/mm_directed.rda")

### 2015-04-01 CC/writeup.Rnw
conclusions/future work

@@
 \texttt{"eigen"} & places vertices based on the eigenstructure of the adjacency matrix \\
-\texttt{"fruchermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
+\texttt{"fruchtermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
 \texttt{"geodist"} & a version of \texttt{"mds"} which scales the matrix of geodesic distances \\
@@
 \section{Future Work}
+\hh{XXX  just be a bit more specific here: we (a) want to make the code a part of the ggplot2 world - either as a part of ggplot2 itself or within GGally.
+(b) we want to publish this paper in the R Journal.
+Things that we still need to do are discussed in the paper, they are: .... }
 As we discussed throughout the paper, there is still some work to be done  before we submit it for publication in \emph{The R Journal}.

### 2015-04-01 CC/writeup.Rnw
Ready to submit to committee! (?)

--- 2015-04-01 CC/writeup.Rnw
@@
 
-\section{Many Many Examples}
+\section{Usage Examples}
 In this section, we demonstrate the current capabilities of \texttt{geom\_net} in a series of diverse examples.
@@
 
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+<<yeast_code>>=
 sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
 yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
+@
+<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
 ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
@@
 <<usage_stat_net, eval=FALSE>>=
-stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", na.rm = FALSE, ...)
+stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
 #aes(colour = "black", size =3, layout="kamadakawai",layout.par = NA)
@@
   The usage of the \texttt{stat\_net} arguments is as follows:
-  \sct{XXX adding detail tomorrow}
   \begin{itemize}
-  \item mapping
-  \item data
-  \item vertices
-  \item geom
-  \item position
+  \item The \texttt{mapping} argument is the aesthetic mapping. Currently, there is no reason to ever change it from \texttt{NULL}.
+  \item The \texttt{data} argument is the edges data frame.  This is sent to \texttt{stat\_net} through the \texttt{ggplot()} function.
+  \item The \texttt{vertices} argument is the vertex data frame. This is sent to \texttt{stat\_net} through the \texttt{geom\_net} function.
+  \item The \texttt{geom} argument is the geometric object used to plot the data. It is by default set to \texttt{net}. 
+  \item The \texttt{position} argument is set to \texttt{identity} because we wish to plot the network as-is, with no adjustments.
   \item na.rm
@@
   \begin{itemize}
-  \item layout 
-  \item layout.par
+  \item The \texttt{layout} aesthetic is the layout algorithm to use to place the vertices of the network.
+The default is the Kamada-Kawai algorithm, and there are other options listed in \ref{table:layouts}.
+  \item The \texttt{layout.par} aesthetic is the layout arguments to pass to the layout function.
+It is set to \texttt{NA} by default, in which case the algorithm uses its set defaults in the \texttt{sna} package.
+This aesthetic takes a list of parameters according to the documentation by \citet{sna}.
   \end{itemize}
@@
   The usage of the \texttt{geom\_net} arguments is as follows:
-  \sct{XXX adding detail tomorrow}
   \begin{itemize}
-  \item mapping
-  \item data
-  \item vertices
-  \item vlabel
-  \item directed
-  \item stat
+  \item The \texttt{mapping} argument is the aesthetic mapping. It is constructed with \texttt{aes()} according to the aesthetic values that can be passed to \texttt{geom\_net}. 
+  \item The \texttt{data} argument is the edges data frame.  There is usually no need to use it in \texttt{geom\_net()} because it will have already been provided in the \texttt{ggplot()} function.
+  \item The \texttt{vertices} argument is the vertex data frame.
+The vertex data frame should be provided to this argument in \texttt{geom\_net}.
+This data frame is required to have a column named \texttt{label} providing the labels of all vertices in the network.
+  \item The \texttt{vlabel} argument is a logical value which plots the label layer on top of the edge and vertex layers when it is set to \texttt{TRUE}. 
+  \item The \texttt{directed} argument is a logical value which adds arrows to the edge layer pointing to the \texttt{to\_id} vertex when it is set to \texttt{TRUE}.  
+  \item The \texttt{stat} argument is the statistical transformation used on the data for the network layer. It is always set to \texttt{net}. 
   \end{itemize}
-  \texttt{stat\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
+  \texttt{geom\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
   \begin{itemize}
-  \item from\_id*
-  \item to\_id*
-  \item ecolour
-  \item esize
-  \item elinetype
-  \item ealpha
-  \item vcolour
-  \item vfill
-  \item vsize
-  \item vshape
-  \item valpha
+  \item The \texttt{from\_id}* and \texttt{to\_id}* aesthetics each take one column which, when combined, provide the edges in a network.
+If the network is undirected, then their order does not matter.
+If the network is directed, then the \emph{from} vertex names should all be in the \texttt{from\_id} column, and their corresponding \emph{to} vertex names should all be in the \texttt{to\_id} column.
+  \item The \texttt{ecolour} aesthetic changes the color of the edges. Its default is \texttt{grey20}
+  \item The \texttt{esize} aesthetic changes the size (thickness) of the edges. Its default is 1. 
+  \item The \texttt{elinetype} aesthetic changes the linetype of the edges. Its default is 1. 
+  \item The \texttt{ealpha} aesthetic changes the transparency of the edges. Its default is 1. 
+  \item The \texttt{vcolour} aesthetic changes the color of the vertices. Its default is \texttt{black}. 
+  \item The \texttt{vfill} aesthetic changes the ouline color of the vertices. Its default is \texttt{black}. 
+  \item The \texttt{vsize} aesthetic changes the size (diameter) of the vertices. Its default is 2.
+  \item The \texttt{vshape} aesthetic changes the shape of the vertices. Its default is 16 (solid circular points). 
+  \item The \texttt{valpha} aesthetic changes the transparency of the vertices. Its default is 1. 
   \end{itemize}
@@
 \section{Future Work}
-As we discussed throughout the paper, there is still some work to be done before we submit it for publication in \emph{The R Journal}.
-We also hope to make it an official part of \texttt{ggplot2}.
+As we discussed throughout the paper, there is still some work to be done.
+The primary change that we need to make is to add the capability of mapping aesthetic values to the plot based on variables directly.
+Right now, we have the ability to do this indirectly, but we want to add this functionality so that it is easier to change the \texttt{ecolor}, \texttt{esize}, \texttt{elinetype}, \texttt{vcolor}, \texttt{vfill}, \texttt{vsize}, and \texttt{vshape} by mapping variables to these aesthetics.
+We also want to include the legends for each of these aesthetics like we would normally see in \texttt{ggplot2}.
+We also hope to make it an official part of the \texttt{ggplot2} world, by either adding it to \texttt{ggplot2} itself or by adding it to the \texttt{GGally} package.
 We believe that this is an important contribution to \texttt{ggplot2} and to the R user community because it fills in gaps in the network plotting capabilities of R.
-\sct{XXX That's all I got.
-Help?}
+Ultimately, our goal is to publish this paper in \textit{The R Journal}.
 

### 2015-04-01 CC/writeup.Rnw
final version (for now)

@@
 To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE>>=
+<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE,eval=FALSE>>=
 ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
@@
 \centering
-\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
+\includegraphics[width=8in]{figure/email_ex_facet-1.pdf}
 \caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
@@
 ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, vcolour = 'magenta',
-           layout = 'random', layout.par = list(dist = 'uniang')) +
+  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
+           vcolour = 'magenta', layout = 'random', 
+           layout.par = list(dist = 'uniang')) +
   expand_limits(x = c(0,1), y = c(0,1)) + theme_net

### 2015-04-02 CC/writeup.Rnw
add data, installation instructions, R files, and examples :wq

@@
 stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
-#aes(colour = "black", size =3, layout="kamadakawai",layout.par = NA)
 @
@@
 geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
-#aes(ecolour = "grey20", vcolour = "black", vfill="black", fill = NA, esize = 1, vsize = 2, elinetype = 1, weight = NA, valpha = 1, vshape = 16, ealpha=1)
 @

### 2015-04-02 Examples/blood_example.R

@@ -1,0 +1,13 @@
+#Blood Donation
+
+load('data/blood.rda')
+
+library(network)
+library(sna)
+library(ggplot2)
+
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
+  geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
+           vlabel = TRUE, vsize = I(3), directed = TRUE) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+

### 2015-04-02 Examples/email_example.R

@@ -1,0 +1,23 @@
+#Email Network from http://vacommunity.org/VAST+Challenge+2014%3A+Mini-Challenge+1#Download_ the_Datasets_Entry_Forms_and_Documentation
+
+load('data/email.rda')
+email.net <- email$edges
+employee <- email$nodes
+
+library(network)
+library(sna)
+library(ggplot2)
+library(RColorBrewer)
+cols <- brewer.pal(8, name="Set1")
+
+#no facets
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
+  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
+           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
+  expand_limits(x=c(0,1), y = c(0,1)) + theme_net
+
+#facet by day 
+ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
+  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
+           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
+  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net

### 2015-04-02 Examples/football_example.R

@@ -1,0 +1,26 @@
+#College Football Games in the Fall 2000 regular season
+# Source: http://www-personal.umich.edu/~mejn/netdata/
+
+load('data/football.rda')
+football$vertices$label <- as.factor(football$vertices$label)
+football$vertices$value <- as.factor(football$vertices$value)
+football$edges$in.conf <- 0
+for (i in 1:nrow(football$edges)){
+  idx1 <- which(football$vertices$label == football$edges$from[i])
+  idx2 <- which(football$vertices$label == football$edges$to[i])
+  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
+    football$edges$in.conf[i] <- 1
+  }
+}
+
+library(network)
+library(sna)
+library(ggplot2)
+library(RColorBrewer)
+colors <- brewer.pal(12, name = 'Paired')
+
+ggplot(data = football$edges, aes(from_id = from, to_id = to,
+                                  elinetype = in.conf + 1)) +
+  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
+           aes(vcolour=colors[as.numeric(factor(value))])) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net

### 2015-04-02 Examples/lesmis_example.R

@@ -1,0 +1,12 @@
+# Les Misérables example
+# Source: http://www-personal.umich.edu/~mejn/netdata/
+
+load('data/lesmis.rda')
+
+library(network)
+library(sna)
+library(ggplot2)
+
+ggplot(data = lesmis$edges, aes(from_id = from, to_id = to, esize = value/mean(value))) +
+  geom_net(vertices = lesmis$vertices, ealpha = 7/10,vcolour = I("red"), vsize = I(2.5), layout = 'fruchtermanreingold', vlabel=TRUE) +
+  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net

### 2015-04-02 Examples/madmenexample.R

@@ -1,0 +1,12 @@
+#Madmen Relationships
+
+load('data/madmen.rda')
+
+library(network)
+library(sna)
+library(ggplot2)
+
+ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
+           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
+  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net

### 2015-04-02 Examples/madmenexample2.R

@@ -1,0 +1,13 @@
+#Attempted Madmen Relationships
+
+load('data/mm_directed.rda')
+
+library(network)
+library(sna)
+library(ggplot2)
+
+ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
+           vsize = I(2.5), layout = 'fruchtermanreingold',
+           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
+  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net

### 2015-04-02 Examples/protein_example.R

@@ -1,0 +1,17 @@
+
+load('data/protein.rda')
+
+#subset the data
+sample.edges <- protein$edges[sample(nrow(protein$edges), 500),]
+yeast.samp <- list(edges = sample.edges, 
+                   vertices = data.frame(label = unique(c(sample.edges$from,sample.edges$to))))
+
+library(network)
+library(sna)
+library(ggplot2)
+
+ggplot(data = yeast.samp$edges, aes(from_id = from, to_id = to)) +
+  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
+           vcolour = 'magenta', layout = 'random', 
+           layout.par = list(dist = 'uniang')) +
+  expand_limits(x = c(0,1), y = c(0,1)) + theme_net

### 2015-04-02 Examples/theme_elements_example.R

@@ -1,0 +1,13 @@
+#Theme Inheritance structure in ggplot2: see bottom of this page: http://docs.ggplot2.org/current/theme.html
+
+load('data/theme_elements.rda')
+names(theme_elements$vertices)[1] <- 'label'
+
+library(network)
+library(sna)
+library(ggplot2)
+
+
+ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
+  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) + 
+  expand_limits(x = c(0,1.2), y = c(0,1)) + theme_net

### 2015-04-02 R/example.R

@@ -1,24 +1,0 @@
-#example data
-all <- paste(sort(rep(c('A', "B", 'O', "AB"),2)), rep(c("+","-"),4), sep = '')
-A <- all[1:2]
-B <- all[5:6]
-AB <- all[3:4]
-O <- all[7:8]
-
-On <- cbind(from = rep('O-',8), to = all, group = 'O')
-Op <- cbind(from = rep('O+',4), to = c(AB[1], A[1], B[1],O[1]), group = 'O')
-Bn <- cbind(from = rep('B-',4), to = c(AB, B), group = 'B')
-Bp <- cbind(from = rep('B+',2), to = c(AB[1], B[1]), group = 'B')
-An <- cbind(from = rep('A-',4), to = c(AB, A), group = 'A')
-Ap <- cbind(from = rep('A+',2), to = c(AB[1], A[1]), group = 'A')
-ABn <- cbind(from = rep('AB-',2), to = AB, group = 'AB')
-ABp <- cbind(from = 'AB+', to = 'AB+', group = 'AB')
-
-blood.edges <- data.frame(rbind(ABn, ABp, An, Ap, Bn, Bp, On, Op))[,1:2]
-blood.nodes <- unique(data.frame(rbind(ABn, ABp, An, Ap, Bn, Bp, On, Op))[,c(1,3)])
-names(blood.nodes) <- c("type", "group")
-blood.nodes$in.degree <- c(4,8,2,4,2,4,1,2)
-blood.nodes$out.degree <- c(2,1,4,2,4,2,8,4)
-
-#ggplot() + geom_net(nodes = blood.nodes, edges = blood.edges, 
-#                     aes(vlabel = type, vcolor = group, vsize = in.degree))

### 2015-04-02 R/first.R

@@ -1,160 +1,0 @@
-#' test function
-#' 
-#' this is what the function does
-#' @export
-#' @param x numeric vector
-#' @param ... addiional parameters passed on to min
-#' @return minimum of the vector x
-minx <-  function(x, ...) return(min(x, ...))
-
-#' @export
-#' @example
-#' data(mtcars)
-#' library(ggplot2)
-#' ggplot(data=mtcars) + geom_net(aes(x=mpg, y=disp)) 
-geom_net <- function (mapping = NULL, data = NULL, stat = "identity", position = "identity",
-                        na.rm = FALSE, ...) {
-  browser()
-  GeomNet$new(mapping = mapping, data = data, stat = stat, position = position,
-                na.rm = na.rm, ...)
-}
-
-GeomNet <- proto::proto(ggplot2:::Geom, {
-  objname <- "net"
-  
-  draw_groups <- function(., ...) .$draw(...)
-  draw <- function(., data, coordinates, na.rm = FALSE, ...) {    
-    browser()
-    data <- remove_missing(data, na.rm,
-                           c("net"), name = "geom_net")
-    # net is network object from network package
-    #    if (empty(data)) return(zeroGrob())
-    if (empty(net)) return(zeroGrob())
-    
-    m <- as.matrix.network.adjacency(net) # get sociomatrix #pkg: network
-    # get coordinates from Fruchterman and Reingold's force-directed placement algorithm.
-    plotcord <- data.frame(gplot.layout.fruchtermanreingold(m, NULL))  #pkg: sna
-    # or get it them from Kamada-Kawai's algorithm: 
-    # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL)) # pkg: sna
-    colnames(plotcord) <- c("x","y")
-    nodes_grob <- GeomPoint(plotcord, ...)
-    
-    edglist <- as.matrix.network.edgelist(net) #pkg: network
-    edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
-    plotcord$elements <- as.factor(get.vertex.attribute(net, "elements")) #pkg: network
-    colnames(edges) <-  c("x","y","xend","yend")
-    edges_grob <- GeomSegment(edges, ...)
-    
-    with(coord_transform(coordinates, data), #got rid of scales b/c we don't have it anywhere else
-         ggname(.$my_name(), grobTree(
-           nodes_grob,
-           edges_grob
-         )))
-  }
-   draw_legend <- function(., data, ...) {
[…]

### 2015-04-02 blood_example.R

@@ -1,49 +1,0 @@
-#Testing of the geom!
-
-all <- paste(sort(rep(c('A', "B", 'O', "AB"),2)), rep(c("+","-"),4), sep = '')
-A <- all[1:2]
-B <- all[5:6]
-AB <- all[3:4]
-O <- all[7:8]
-
-On <- cbind(from = rep('O-',8), to = all, letter = 'O')
-Op <- cbind(from = rep('O+',4), to = c(AB[1], A[1], B[1],O[1]), letter = 'O')
-Bn <- cbind(from = rep('B-',4), to = c(AB, B), letter = 'B')
-Bp <- cbind(from = rep('B+',2), to = c(AB[1], B[1]), letter = 'B')
-An <- cbind(from = rep('A-',4), to = c(AB, A), letter = 'A')
-Ap <- cbind(from = rep('A+',2), to = c(AB[1], A[1]), letter = 'A')
-ABn <- cbind(from = rep('AB-',2), to = AB, letter = 'AB')
-ABp <- cbind(from = 'AB+', to = 'AB+', letter = 'AB')
-
-blood <- data.frame(rbind(ABn, ABp, An, Ap, Bn, Bp, On, Op))
-type <- c('AB','A','B','O')
-
-blood.edges <- blood[-which(blood$from == blood$to),]
-blood.vertices <- data.frame(node_id=unique(c(as.character(blood$from), as.character(blood$to))))
-
-#node_id is a required aesthetic. #i have yet to match it with the nodes data in the geom code.
-ggplot(data = blood.edges, aes(from_id= from, to_id = to)) + geom_net(vertices = blood.vertices) + expand_limits(x = -5, y = -1)
-
-
-# #try again: 1/19
-# 
-# #get network coords first, outside of geom
-# net <- as.network(blood.edges, matrix.type = "edgelist") #from network package
-# m <- as.matrix.network.adjacency(net)
-# #change the stuff below later to include multiple layouts (in if else statements?)
-# vert.coord <- data.frame(sna::gplot.layout.kamadakawai(m, NULL)) # from sna package
-# vert.labels <- as.factor(unique(c(as.character(blood.edges$from),as.character(blood.edges$to))))
-# vert.coord$name <- sort(vert.labels) #needed here?
-# names(vert.coord) <- c('x','y','node_id')
-# edgelist <- as.matrix.network.edgelist(net)
[…]

### 2015-04-02 geom-net.r

@@ -1,4 +1,4 @@
-#' Add heatmap of 2d bin counts.
+#' Geom for visualizing a network.
 #'
 #' @section Aesthetics:
 #' \Sexpr[results=rd,stage=build]{ggplot2:::rd_aesthetics("geom", "bin2d")}
@@ -36,33 +36,124 @@
 #' }
 #'
 #'
-#' g <- network(150, directed=FALSE, density=0.03)
+#' g <-
+#' (150, directed=FALSE, density=0.03)
 #' classes <- rbinom(150,1,0.5) + rbinom(150,1,0.5) + rbinom(150,1,0.5)
 #' set.vertex.attribute(g, "elements", classes)
 #' plotg(g)
+#'
+#' # make a theme for networks:
+#' theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
+#' data(blood) #3/16 - added to the data file in dropbox. a list with an edges df and a vertices df.
+#' #node_id is a required aesthetic. #i have yet to match it with the nodes data in the geom code.
+#' ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5))
+
+#' #ecolour by group_to doesnt work
+#' ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices, aes(ecolour = group_to)) + expand_limits(x = c(-5,5), y = c(-5,5))
+#' #facetting seems to work?
+#' ggplot(data = blood$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = blood$vertices) + expand_limits(x = c(-5,5), y = c(-5,5)) + facet_wrap(~group_to)
+#' #Mad Men Example
+#' sex <- c("m", "f", "m", "m", "f", "f", "f", "m", "f", "m", "f", "m", "m", "f", "m", "m", "f", "m", "m",
+#' rep("f",9), "m", "m", "m", "f", "f", "f", "f", "m", "f", "m", "f", "f", "m", "f", "m", "m", "f",
+#' "m", "f", "f", "f", "f", "f", "f")
+#' data(madmen2, package="gcookbook")
+#' mm.vert <- data.frame(Name = as.factor(unique(c(madmen2$Name1,madmen2$Name2))), Gender = as.factor(sex))
[…]

### 2015-04-02 madmenexample.R

@@ -1,29 +1,0 @@
-#mad men example
-data(madmen, package="gcookbook")
-sex <- c("f", "m", "m", "f","m","f",rep("m",5),rep("f",15), "m", "m", "f","f","f","m","m","m","m",rep("f",8),"m","f")
-mm.vert <- data.frame(Name = as.factor(unique(c(madmen$Name1,madmen$Name2))), Gender = as.factor(sex))
-
-madmen <- list(edges = madmen, vertices = mm.vert)
-#save(madmen, file = "~/Desktop/NetworksResearch/ggnet/CC/data/madmen.rda")
-
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) + geom_net(vertices = madmen$vertices) + 
-  expand_limits(x = c(-10,10), y = c(-10,10)) + theme_net
-
-data(madmen2, package = "gcookbook")
-head(madmen2)
-
-label <- unique(c(madmen2$Name1,madmen2$Name2))
-Gender <- c('m','f',rep('m',2), rep('f',3), 'm', 'f', 'm','f',rep('m',2), 'f', rep('m',2), 'f', rep('m',2), rep('f',9),
-            rep('m',3), rep('f',4), 'm', 'f','m',rep('f',2), 'm', 'f','m','m','f','m',rep('f',6))
-
-vert = data.frame(label,Gender)
-head(vert)
-
-mm.directed <- list(edges = madmen2, vertices = vert)
-aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])
-
-ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) + 
-  geom_net(vertices = mm.directed$vertices, layout= "random", l.param = list(dist = 'normal'),directed = T, vlabel = T, vsize = I(2.5), aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) + 
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-
-save(mm.directed, file = "~/Dropbox/ggplot2/data/mm_directed.rda")

### 2015-04-02 stat-net.r

@@ -18,46 +18,72 @@
 #' \donttest{
 #' #put in example here.
 #' }
-stat_net <- function (mapping = NULL, data = NULL, geom = "net", position = "identity", scales, ...) {
-  StatNet$new(mapping = mapping, data = data, geom = geom,
-  position = position, layout = NULL, ...)
+stat_net <- function (mapping = NULL, data = NULL, vertices = NULL,  geom = "net", position = "identity", ...) {
+#  print("new net stat")
+
+  StatNet$new(mapping = mapping, data = data, vertices = vertices, geom = geom,
+  position = position, ...)
 }
 
 StatNet <- proto(Stat, {
   objname <- "net"
+  default_geom <- function(.) GeomNet
+  default_aes <- function(.) aes(colour = "black", size =3, layout="kamadakawai",layout.par = NA)
+  #required_aes <- c("x", "y")
+  #required_aes <- c("from_id", "to_id")
 
-  calculate_groups <- function(., data, scales, ...){
-    .super$calculate_groups(., data, scales,...)
+  calculate_groups <- function(., data, coordinates=coordinates, vertices = vertices,  ...){
+#    print("calculate groups net")
+#    browser()
+
+    data$group <-1
+    .super$calculate_groups(., data, coordinates=coordinates, vertices = vertices, ...)
   }
 
 #calculate function: extracts smaller data frame from original data using just req params. then computes some other stuff.
-  calculate <- function(., data, scales, na.rm = FALSE,  layout = TRUE, ...) {  #add requirements here later.
-    browser()
+  calculate <- function(., data, na.rm = FALSE,  coordinates=coordinates, vertices = vertices, ...) {  #add requirements here later.
+#    print("calculate net")
+#    browser()
 
     #compute x and y using layout method if x,y are not there. then can assume that x and y are there from now on.
-    library(network,sna) #how/where to call network, sna, pkg here?
+    require(network)
+    require(sna) #how/where to call network, sna, pkg here?
     stopifnot(is.data.frame(data))
-    #first column of data frame needs to be the 'from' vertex.
-    #second column of data frame needs to be the 'to' vertex. other cols can be anything
-    if (!is.null(data$from)) data[,1] <- data$from
-    if (!is.null(data$to)) data[,2] <- data$to
-    edges <- data[,c('from','to')]
[…]

### 2015-04-02 geom-net.r
update headers

@@ -1,72 +1,6 @@
 #' Geom for visualizing a network.
-#'
-#' @section Aesthetics:
-#' \Sexpr[results=rd,stage=build]{ggplot2:::rd_aesthetics("geom", "bin2d")}
-#'
-#' @export
-#' @inheritParams geom_point
-#' @examples
-#' This is Mortiz Marbach's plotg function, from his post at http://sumtxt.wordpress.com/2011/07/02/visualizing-networks-with-ggplot2-in-r/.
-#' plotg <- function(net, value=NULL) {
-#'  m <- as.matrix.network.adjacency(net) # get sociomatrix
-#'   # get coordinates from Fruchterman and Reingold's force-directed placement algorithm.
-#'  plotcord <- data.frame(gplot.layout.fruchtermanreingold(m, NULL))
-#'   # or get it them from Kamada-Kawai's algorithm:
-#'  # plotcord <- data.frame(gplot.layout.kamadakawai(m, NULL))
-#'colnames(plotcord) = c("X1","X2")
-#'edglist <- as.matrix.network.edgelist(net)
-#'edges <- data.frame(plotcord[edglist[,1],], plotcord[edglist[,2],])
-#'plotcord$elements <- as.factor(get.vertex.attribute(net, "elements"))
-#'colnames(edges) <-  c("X1","Y1","X2","Y2")
-#'edges$midX  <- (edges$X1 + edges$X2) / 2
-#'edges$midY  <- (edges$Y1 + edges$Y2) / 2
-#'pnet <- ggplot()  +
-#'  geom_segment(aes(x=X1, y=Y1, xend = X2, yend = Y2),
-#'               data=edges, size = 0.5, colour="grey") +
-#'  geom_point(aes(X1, X2,colour=elements), data=plotcord) +
-#'  scale_colour_brewer(palette="Set1") +
-#'  scale_x_continuous(breaks = NA) + scale_y_continuous(breaks = NA) +
-#'  # discard default grid + titles in ggplot2
-#'  opts(panel.background = theme_blank()) + opts(legend.position="none")+
-#'  opts(axis.title.x = theme_blank(), axis.title.y = theme_blank()) +
-#'  opts( legend.background = theme_rect(colour = NA)) +
-#'  opts(panel.background = theme_rect(fill = "white", colour = NA)) +
-#'  opts(panel.grid.minor = theme_blank(), panel.grid.major = theme_blank())
-#' return(print(pnet))
-#' }
-#'
-#'
-#' g <-
[…]

### 2015-04-02 stat-net.r

@@ -1,23 +1,7 @@
 #' 2d density estimation.
-#'
-#' @section Aesthetics:
-#' \Sexpr[results=rd,stage=build]{ggplot2:::rd_aesthetics("stat", "density2d")}
-#'
-#' @param contour If \code{TRUE}, contour the results of the 2d density
-#'   estimation
-#' @param n number of grid points in each direction
-#' @param ... other arguments passed on to \code{\link{kde2d}}
-#' @param na.rm If \code{FALSE} (the default), removes missing values with
-#'    a warning.  If \code{TRUE} silently removes missing values.
-#' @inheritParams stat_identity
-#' @return A data frame in the same format as \code{\link{stat_contour}}
-#'
-#@importFrom network
-#' @export
-#' @examples
-#' \donttest{
-#' #put in example here.
-#' }
+#' #Documentation to be added
+#' 
+#' 
 stat_net <- function (mapping = NULL, data = NULL, vertices = NULL,  geom = "net", position = "identity", ...) {
 #  print("new net stat")
 

### 2015-10-01 inst/CC/writeup.Rnw
reviving ggnet as package

--- 2015-04-02 CC/writeup.Rnw
@@
 cols <- brewer.pal(8, name="Set1")
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
+load("data/protein.rda")
+
+#yeast <- list(edges = protein, vertices = data.frame(label = min(yeast):max(yeast)))
+yeast <- protein
 @

### 2015-10-01 inst/CC/writeup.tex
reviving ggnet as package

--- 2015-04-02 CC/writeup.Rnw
@@
 %%% BEGIN DOCUMENT
+\usepackage{Sweave}
 \begin{document}
@@
 \end{abstract}
-<<chunk_setup, echo=FALSE>>=
-library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
-@
-<<readingdata, eval=TRUE, echo=FALSE>>=
-library(ggplot2) # needs to be the version with geom_net!
- theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements, lesmis,madmen, mm_directed,email)
-names(blood$vertices)[1] <- 'label'
-#load("data/madmen.rda") # relative path should work for both of us
-#load("data/mm_directed.rda")
-#names(madmen$vertices)[1] <- 'label'
-names(theme_elements$vertices)[1] <- 'label'
-football$vertices$label <- as.factor(football$vertices$label)
-football$vertices$value <- as.factor(football$vertices$value)
-football$edges$in.conf <- 0
-for (i in 1:nrow(football$edges)){
-  idx1 <- which(football$vertices$label == football$edges$from[i])
-  idx2 <- which(football$vertices$label == football$edges$to[i])
-  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
-    football$edges$in.conf[i] <- 1
-  }
-}
-#load("data/email.rda")
-#email.net <- read.csv("data/email-data.csv")
-#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-#library(plyr)
-#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-#employee <- read.csv("data/EmployeeRecords.csv")
-#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
-email.net <- email$edges
-employee <- email$nodes
-library(RColorBrewer)
-cols <- brewer.pal(8, name="Set1")
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
-@
-
-\section{Introduction}
-At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.
-Here, we refer to the lines as edges and the points as vertices, although these are also called nodes.
-These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study.
-For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.
-Additionally, \citet{networkfailures} examine node failure in interdependent networks like power grids.
-Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C.
-elegans} worm are also etensively studied \citep{smallworld}.
-Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used and most complex networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.
-The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
-The world wide web is an example of a directed network because one webpage may link to another, and not necessarily the other way around.
-Undirected edges, however, are simply connections between vertices.
-Coauthorship networks that encode information about academic publications are examples of undirected networks because if two people author a paper together, that creates a connection between them that is bidirectional.
-
-\par A social network is a network that everyone is a part of in one way or another.
-We do not necessarily refer here to social media like Facebook or LinkedIn, but rather to the connections we form with other people.
-To demonstrate the functionality of our geometry for plotting networks, we have chosen an example of a social network from the popular television show Mad Men.
-This network was compiled in \citet{madmen}.
-In this example network, there are 52 vertices and 87 edges.
-Each vertex is a character on the show, and there is an edge between every two characters who have had a romantic relationship.
-
-This network is shown in figure~\ref{fig.cap:madmen_ex}.
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net
-@
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
-\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romantic relationship.}
-\end{figure}
-\afterpage{\clearpage}
-
- In the plot, we can see one central character who has many more relationships than any other character.
-This vertex represents the main character of the show, Don Draper, who is quite the ``ladies' man."  This fun example shows just how ubiquitous networks are.
-\par There are many kinds of networks, and networks are extensively studied across many disciplines.
-Many sociologists study social networks, and many biologists study protein networks.
-As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.
-We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}, from January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
-It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times\footnote{\texttt{ggplot2} usage statistics taken from \url{http://cran-logs.rstudio.com/}.}.
-This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.
-
-\par The three necessary elements of any network visualization are the vertices, the edges, and the layout.
-But once those three items are visualized, we usually find our visualization to be lacking.
-We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.
-Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by \citet{igraph}, \texttt{sna} by \citet{sna}, and \texttt{network} by \citet{network.jss, network} but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network.
-For instance, the \texttt{igraph} package allows for coloring vertices by groups but the user must assign the colors to each vertex individually as opposed to assigning color by a grouping factor variable.
-\\
-
-\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
-Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
-The vertices in a network and the edges between them compose the structure of a network, and being able to discover patterns among them visually is a key part of network analysis.
-Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.
-The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
-Our work builds off of the \texttt{ggnet} function and presents an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
-
-  \subsection{Data Structure}
-In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
-The vertex data frame should contain all the relevant vertex information.
-The only necessary variable is one called \texttt{label} which contains the vertex labels as they are encoded in the edges data set.
-This will be a formal requirement in the final version of the \texttt{geom\_net} function.
-Each row represents one vertex.
-Other values of interest, such as grouping variables or degree of each vertex should be stored as columns in this dataframe, with an observation for each vertex.
-  \par The edge data frame should contain all the relevant edge information.
-The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network.
-The From and To vertices should match the names of the vertices in the vertex information data frame.
-The formal names for these columns are \texttt{from\_id} and \texttt{to\_id}, respectively.
-Other variables may also be included for each edge, such as the edge weight or grouping variable.
-As before, the variables of interest are columns in the data frame and the rows are each edge in the network.
-
-  \subsection{Vertex Aesthetics}
-
-  In our geometry, we want to create all possible vertex aesthetic to mimic the usual aesthetics in \texttt{ggplot2} for points.
-The \texttt{vcolor} aesthetic can be an identity color (e.g.
-\texttt{color = "red"}) to change the color of all vertices, or, in the final version, it will take a factor variable which identifies each vertex as belonging to one of several groups, and it colors the vertices of the graph according to this grouping.
-In the current version, the \texttt{vcolor} aesthetic does not work like a typical color aesthetic in \texttt{ggplot2}.
-It can, however, take HEX color characters and assign them to different levels of a factor variable.
-We use the \texttt{RColorBrewer} package to choose the colors for us in this paper \citep{RColorBrewer}.
-It will also color the vertices along a color gradient according to the different values of some numeric variable such as degree.
-The vertex color defaults to black.
-% are set to the default color palette in \texttt{ggplot2}, and the palette used can be changed in the usual way within the \texttt{ggplot2} framework, by using \texttt{scale\_color\_brewer}.}
-  %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
-  The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to one of any numerical columns in the vertex data frame.
-Right now, the numerical values are strict point values: if the size of a vertex is two, then it will be two points in diameter.
-A point is the size of the default fontsize in the \texttt{grid} package.
-In our final version, the \texttt{vsize} will scale to minimum and maximum values as all other size aesthestics in \texttt{ggplot2} do.
-The \texttt{vshape} aesthetic is for changing the shape of the vertices from the default circle to other shapes, like square or triangle.
-It also changes the shape of the vertices according to some grouping variable, which is currently required to consist of integer values from 0 to 25.
-In the final version, it will be able to change shape based on different levels of a factor variable.
-Finally, the \texttt{valpha} aesthetic will change the vertex transparency to a set value, like 1/10 or 0.5.
-It operates in exactly the same way as the \texttt{alpha} aesthetic in \texttt{geom\_point}.
-This aesthetic is very useful for networks with hundreds or thousands of vertices, which can easily crowd the static visualization.
-Making the vertices more transparent will better show the underlying structure in the network.
-
-\par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion.
-  \subsection{Edge Aesthetics}
-  Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for segments.
-The \texttt{ecolor} aesthetic changes the color of all of the edges according to the identity function or can change the color according to HEX colors assigned to levels of a grouping variable.
-In the final version, the \texttt{ecolor} aesthetic will automatically color according to levels of a grouping variable or according to a gradient scale associated with a numeric variable.
-  The \texttt{elinetype} aesthetic can also be used with a grouping variable to change the line type of each edge in the graph from solid to one of the different types of lines in \texttt{ggplot2}.
-Currently, the column associated with this aesthetic must be integer valued and take on values from zero to six.
-In the final version, it will change the edge linetypes according to levels of a grouping variable.
-The \texttt{esize} aesthetic can also be set to the weight or probability of each edge, or it can be changed for all edges with the identity function.
-The associated column can take on any numerical values, and currently maps the values directly to points, just like the \texttt{vsize} aesthetic.
-In the final version, this aesthetic will work exactly like the sizing in \texttt{ggplot2}, by scaling sizes to the minimum and maximum values.
-This is one of two ways to visualize the number of edges between two vertices.
-The other way is through the \texttt{ealpha} aesthetic.
-The \texttt{ealpha} aesthetic can change the edge transparency to a set value on the interval $[0,1]$.
-If there are multiple edge connections given between two vertices in the edge data frame, the \texttt{ealpha} aesthetic can be changed to plot opaque lines between vertices with many mutual connections and varying degrees of transparent lines between vertices with only a few or a single connection.
-This property is also especially useful for large networks with a lot of connections or with vertices of relatively high degree.
-Finally, we will add an \texttt{elabel} aesthetic to print the name or number assigned to each edge on or next to it.
-This aesthetic will be very useful for visualizing random graphs or Markov processes where each edge probability is of interest.
-
-  \subsection{Other Arguments}
-  Similar to the other geometries and functions in \texttt{ggplot2}, our geometry takes several arguments outside of the aesthetic mapping parameters.
-One such argument is the \texttt{vlabel} argument.
-The \texttt{vlabel} aesthetic is a logical value that, if set to \texttt{TRUE} prints the name or number associated with each vertex in the \texttt{label} column in order to identify it on the plot.
-In the final version of this geometry, we will add capabilities to adjust the size and positioning of the labels.
-For the moment, the color is set to the vertex color, the size is set to twice the vertex size, and all other possible parameters are set to their defaults in \texttt{geom\_text()}.
-The \texttt{vlabel} argument is most useful for smaller network objects where all vertex names can be printed on their corresponding vertices and still be read clearly.
-Next, the \texttt{directed} argument is a logical value that identifies whether or not the network is directed.
-When this value is true, arrows are created on the end of the line segment that corresponds to the \texttt{to\_id} value for each edge.
-The arrows match the edges in coloring, linetype, and size, and the default length of the arrow is set using the \texttt{unit} function in the \texttt{grid} package to 0.015 of the normalised parent coordinates of the plot.
-  Finally, the \texttt{layout} argument takes a character value corresponding to the possible layouts in the \texttt{sna} package that are created by the \texttt{gplot.layout.*()} family of functions.
-The default layout is the Kamada-Kawai layout.
-This is a force-directed layout for undirected networks \citep{kamadakawai}.
-There are, however, many other layouts possible.
-\begin{table}
-\centering
-\begin{tabular}{l|p{10cm}}
-Layout Name & Description \\
-\hline \hline
-\texttt{"adj"} & a version of \texttt{"mds"} which scales the raw adjacency matrix \\
-\texttt{"circle"}* & places vertices uniformly in a circle \\
-\texttt{"circrand"} & places vertices randomly in a circle \\
-\texttt{"eigen"} & places vertices based on the eigenstructure of the adjacency matrix \\
-\texttt{"fruchtermanreingold"} & uses a variant of Fruchterman and Reingold's force-directed placement algorithm \\
-\texttt{"geodist"} & a version of \texttt{"mds"} which scales the matrix of geodesic distances \\
-\texttt{"hall"}* &  places vertices based on the last two eigenvectors of the Laplacian of the input matrix \\
-\texttt{"kamadakawai"} & generates a vertex layout using a version of the Kamada-Kawai force-directed placement algorithm \\
-\texttt{"mds"}  & places vertices based on a metric multidimensional scaling of a specified distance matrix \\
-\texttt{"random"} & places vertices randomly \\
-\texttt{"segeo"} & a version of \texttt{"mds"} which scales the squared euclidean distances between row-wise geodesic distances \\
-\texttt{"seham"} & a version of \texttt{"mds"} which scales the Hamming distance between rows/columns of the adjacency matrix \\
-\texttt{"target"} & produces a ``target diagram" or ``bullseye" layout using a force-directed placement algorithm.
-\end{tabular}
-\caption{All possible layouts to pass to \texttt{geom\_net}.
-All descriptions are from \citet{sna}.
-Note that the layouts marked with an asterisk (*) are the only layouts that do not take layout parameters.}
-\label{table:layouts}
-\end{table}
-
-All layouts that \texttt{geom\_net} is currently capable of graphing are listed in table \ref{table:layouts}.
-The layouts not marked with an asterisk can also take their associated layout parameters.
-The \texttt{layout.par} argument takes a list of named parameters.
-For instance, if \texttt{layout = "random"}, we can also set \texttt{layout.par = list(dist = "normal")} to change the distribution of vertex placement from the default uniform to Gaussian.
-
-\section{Usage Examples}
-In this section, we demonstrate the current capabilities of \texttt{geom\_net} in a series of diverse examples.
-  \subsection{Blood Donation}
-  In this directed network, there are eight vertices and 27 edges.
-The vertices represent the eight different blood types in humans that are most important for donation: the ABO blood types A, B, AB, and O, combined with the RhD positive (+) and negative (-) types.
-The edges are directed: a person whose blood type is that of a \emph{from} vertex can to donate blood to a person whose blood type is that of a corresponding \emph{to} vertex.
-In the example below, loops are removed because loops exist on every vertex in this example, as blood between two people of matching ABO and RhD type can always be exchanged.
-<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
-           vlabel = TRUE, vsize = I(3), directed = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-This network is shown in figure \ref{fig.cap:blood_ex}.
-Here, we have used the aesthetics \texttt{vcolour} and \texttt{vsize} set to identity values to change the size and color of all vertices.
-We have also used the \texttt{layout} and \texttt{vlabel} arguments to change the default layout to a circle layout and to print the blood types, respectively.
-The circle layout places blood types of the same ABO type next to each other and spreads the vertices out far enough to distinguish between the various ``in" and ``out" types.
-You can tell clearly from this plot that the O- type is the universal donor: it has out degree of seven and in degree of zero.
-Additionally, you can see that the AB+ type is the universal recepient, with in degree of seven and out degree of zero.
-Anyone looking at this plot can quickly determine which type(s) of blood they can receive and which type(s) can receive their blood.
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
-\caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
-\end{figure}
-\afterpage{\clearpage}
-
-\subsection{Email Network}
-  This email network comes from the 2014 VAST Challenge \citep{emailnet}.
-It is a directed network of emails between company employees with 55 vertices and 9,063 edges.
-Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees.
-The arrow of the directed edge points to the recipient(s) of the email.
-The network contains two business weeks of emails across the entire company.
-In order to better visualize the structure of the communication network between employees, all emails that were sent out to all employees are removed in the subsequent examples.
-<<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
-\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
-\end{figure}
-\afterpage{\clearpage}
-
-This network is plotted in figure \ref{fig.cap:email_ex}.
-There are six distinct clusters in this network which almost perfectly correspond to the six different types of employee in this company: administration, engineering, executive, facilities, information technology, and security.
-Unfortunately, there is currently no legend to associate color with employment type.
-This will be remedied in the final version of our geometry.
-Additionally, the edges between employees in the same cluster are much darker than edges between employees in different clusters.
-This is due to the value of the \texttt{ealpha} aesthetic: more emails between two employees lead to darker edges.
-The value is set to 0.1 in this example, so that 10 or more emails between 2 employees results in a completely opaque edge.
-This pattern of heavy communication between employees of the same type is fairly unsurprising.
-To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE,eval=FALSE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee, ealpha = 0.2, vsize = I(2), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net
-@
-\begin{landscape}
-\begin{figure}
-\centering
-\includegraphics[width=8in]{figure/email_ex_facet-1.pdf}
-\caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
-\end{figure}
-\end{landscape}
-This plot is shown in figure \ref{fig.cap: email_ex_facet}.
-With the facetting, we can see that there are several days where one or more departments do not communicate with any of the other departments.
-There are only two days, 13 and 15, without any isolated department communications.
-Facetting is one of the major benefits of creating a geometry for networks in \texttt{ggplot2}.
-Facetting quickly separates dense networks into separate subnetworks for easy visual comparison and  analyses.
-  \subsection{\texttt{ggplot2} Theme Elements}
-This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
-It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.
-There are 53 vertices and 36 edges in this network.
-Each vertex represents one possible theme option.
-There is an arrow from one theme option to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.
-For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.
-Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and no lines appear anywhere on the plot background.
-Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.
-<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
-  expand_limits(x = c(0,1.2), y = c(0,1)) +
-  theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
-\caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
-\end{figure}
-\afterpage{\clearpage}
-
-The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
-In this plot, it is easy to quickly determine the parent and child vertices.
-Using this plot made creation of the \texttt{theme\_net} object used throughout these examples very simple.
-We just made set each of the major parent elements, \texttt{text}, \texttt{rect}, and \texttt{line} to \texttt{element\_blank()} and then set the aspect ratio equal to one.
-  \subsection{Mad Men Networks}
-  The following code creates the network example given in the introduction.
-We changed the vertex size and edge color for all vertices and edges, included vertex labels, and colored the vertices according to the character's gender.
-<<print_ex, ref.label="madmen_ex", eval=FALSE, echo=TRUE,cache=FALSE>>=
-@
- There is another Mad Men network included in the \texttt{gcookbook} package \citep{madmen}.
-It is a directed network, also of romantic relationships between characters, but it also includes advances made by one character that were rejected by the other.
-For example, Roger Sterling made advances toward Betty Draper, but Betty refused him, and so there is a directed edge going from Roger to Betty, but not from Betty to Roger.
-If the advance is reciprocated, like between Sal Romano and the Bellhop, there are two directed edges between the two vertices.
-<<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
-ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
-           vsize = I(2.5), layout = 'fruchtermanreingold',
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/madmen2_ex-1.pdf}
-\caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
-\end{figure}
-This network is shown in figure \ref{fig.cap:madmen2_ex}.
-This network is much more connected than the previous Mad Men example.
-It also allows us to see much more of the drama in the show.
-For instance, you can see that Roger made advances towards Betty, his business partner's wife, which is much more scandalous than what we can see in the first network.
-
-  \subsection{College Football}
-  This next example comes from M.E.J.
-Newman's network data web page \citep{football}.
-It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.
-There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.
-There is an additional variable in the vertex data frame corresponding to the conference each team belongs to, and there is an additional variable in the edge data frame that is equal to one if the game occured between teams in the same conference or zero if the game occured between teams in different conferences.
-<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-colors <- brewer.pal(12, name = 'Paired')
-ggplot(data = football$edges, aes(from_id = from, to_id = to,
-                                  elinetype = in.conf + 1)) +
-  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
-           aes(vcolour=colors[as.numeric(factor(value))])) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-The network of football games is given in figure \ref{fig.cap:football_ex}.
-Here, we have changed the \texttt{elinetype} aesthetic to correspond to games that occur between teams in the same conference or different conferences.
-These lines are dotted and solid, respectively.
-We have also assigned a different color to each conference, and the vertices and their labels are colored according to their conference.
-This coloring and changing of the linetypes make the structure of the game network easier to view.
-There is one conference consisting of Navy, Notre Dame, Utah State, Central Florida, and Connecticut, which is spread out all over the network, whereas most other conferences' teams are all very close to each other because they play within conference much more than they play out of conference.
-At the time, these five schools were all independents and did not have a home conference.
-Without the coloring capability, we would not have been able to pick out that difference as easily.
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
-\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
-\end{figure}
-\afterpage{\clearpage}
-
-  \subsection{Les Mis\'{e}rables}
-  This next network comes from \citet{lesmis}.
-It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.
-There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.
-There are 254 edges in this network.
-The edges are also weighted by the number of coappearances.
-The largest weighting is 31, between the characters Jean Valjean and Cosette.
-This network is shown in figure \ref{fig.cap:lesmis_ex}.
-This network reveals many important character groups, as well as the main characters in the story.
-The main characters are the ones with high degree, like Valjean, Cosette, and Javert.
-There is an extremely wide edge connecting Valjean and Cosette, which implies that they are strongly connected; in fact, Cosette is Valjean's adopted daughter.
-<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to,
-                                esize = value/mean(value))) +
-  geom_net(vertices = lesmis$vertices, ealpha = 7/10,
-           vcolour = I("red"), vsize = I(2.5),
-           layout = 'fruchtermanreingold', vlabel=TRUE) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
-\caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
-\end{figure}
-\afterpage{\clearpage}
-
-  \subsection{Protein Interaction Network in Yeast}
-  This example of a protein interaction network comes from \citet{protein}.
-It is the complete protein-protein interaction network in the yeast species \emph{S.
-cerevisiae}.
-There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.
-These edges represent ``direct physical interactions" between any 2 proteins \citep[][p.
-42]{protein}.
-Because the network is very large, we randomly sampled 500 interactions and their associated proteins to plot it in figure \ref{fig.cap:yeast_ex}.
-We also demonstrate the layout capabilities by changing the layout to random and setting the distribution to \texttt{"uniang"}, which is a  ``gaussian donut" layout.
-Indeed, we see a nearly round area in the middle of the graph where the layout parameter has forced there to be no vertices.
-
-<<yeast_code>>=
-sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
-@
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
-           vcolour = 'magenta', layout = 'random', 
-           layout.par = list(dist = 'uniang')) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
-\end{figure}
-\afterpage{\clearpage}
-
-\section{Technical Details}
-
-In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
-The stat file performs all the necessary calculations on the two input data frames, and the geom file performs the drawing of the network.
-  \subsection{Creating \texttt{stat\_net}}
-  In order to plot the network with only the edge connections and vertex names, we relied on the \texttt{sna} and \texttt{network} packages \citep{sna, network.jss}.
-First, we used the edges data frame to construct a network object with the \texttt{as.network()} function, then constructed an adjacency matrix from that network using the \texttt{as.matrix.network.adjacency()} function.
-This adjacency matrix is then passed to the chosen \texttt{gplot.layout.*()} function from the \texttt{sna} package.
-\emph{When the layout argument is changed in the \texttt{geom\_net()} function, the corresponding \texttt{gplot.layout.*()} function in \texttt{sna} is called with a \texttt{do.call()} statement.} This layout function produces a matrix of coordinates of the vertices, which we then transform into a data frame containing the coordinates of the edges using the \texttt{as.matrix.network.edgelist()} function in \texttt{network}.
-We are grateful to Moritz Marbach for his \texttt{gplot()} function which we used when creating this process \citep{marbach}.
-At the end of the previous routine, we have a data frame with \texttt{x, y, xend,} and \texttt{yend} columns describing the edges and another data frame with \texttt{x,y} columns describing the vertices.
-These two data frames, along with the many \texttt{aes()} and other argument options get passed on to the actual network geometry for drawing.
-   \subsection{\texttt{stat\_net} Usage}
-<<usage_stat_net, eval=FALSE>>=
-stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
-@
-  The usage of the \texttt{stat\_net} arguments is as follows:
-  \begin{itemize}
-  \item The \texttt{mapping} argument is the aesthetic mapping. Currently, there is no reason to ever change it from \texttt{NULL}.
-  \item The \texttt{data} argument is the edges data frame.  This is sent to \texttt{stat\_net} through the \texttt{ggplot()} function.
-  \item The \texttt{vertices} argument is the vertex data frame. This is sent to \texttt{stat\_net} through the \texttt{geom\_net} function.
-  \item The \texttt{geom} argument is the geometric object used to plot the data. It is by default set to \texttt{net}. 
-  \item The \texttt{position} argument is set to \texttt{identity} because we wish to plot the network as-is, with no adjustments.
-  \item na.rm
-  \end{itemize}
-  \texttt{stat\_net} also accepts the following aesthetic values, none of which are required:
-  \begin{itemize}
-  \item The \texttt{layout} aesthetic is the layout algorithm to use to place the vertices of the network.
-The default is the Kamada-Kawai algorithm, and there are other options listed in \ref{table:layouts}.
-  \item The \texttt{layout.par} aesthetic is the layout arguments to pass to the layout function.
-It is set to \texttt{NA} by default, in which case the algorithm uses its set defaults in the \texttt{sna} package.
-This aesthetic takes a list of parameters according to the documentation by \citet{sna}.
-  \end{itemize}
-
-  \subsection{Creating \texttt{geom\_net}}
-  We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
-Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.
-We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function to add a layer of segments to the plot.
-Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function to construct a point layer for the vertices.
-Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE.
-This adds a layer of text on top of the other two layers.
-We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default.
-We then created a new data frame for the label layer.
-This allows the vertex properties to be propogated through to the labels, including size and color.
-Finally, we added another if/else statement for the \texttt{directed} argument.
-This adds arrow arguments to the edges data frame which will put arrows pointing toward the ``to" vertex in a directed network corresponding to the \texttt{to\_id} value in the edges data frame.
-
-   \subsection{\texttt{geom\_net} Usage}
-<<usage_geom_net, eval=FALSE>>=
-geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
-@
-  The usage of the \texttt{geom\_net} arguments is as follows:
-  \begin{itemize}
-  \item The \texttt{mapping} argument is the aesthetic mapping. It is constructed with \texttt{aes()} according to the aesthetic values that can be passed to \texttt{geom\_net}. 
-  \item The \texttt{data} argument is the edges data frame.  There is usually no need to use it in \texttt{geom\_net()} because it will have already been provided in the \texttt{ggplot()} function.
-  \item The \texttt{vertices} argument is the vertex data frame.
-The vertex data frame should be provided to this argument in \texttt{geom\_net}.
-This data frame is required to have a column named \texttt{label} providing the labels of all vertices in the network.
-  \item The \texttt{vlabel} argument is a logical value which plots the label layer on top of the edge and vertex layers when it is set to \texttt{TRUE}. 
-  \item The \texttt{directed} argument is a logical value which adds arrows to the edge layer pointing to the \texttt{to\_id} vertex when it is set to \texttt{TRUE}.  
-  \item The \texttt{stat} argument is the statistical transformation used on the data for the network layer. It is always set to \texttt{net}. 
-  \end{itemize}
-  \texttt{geom\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
-  \begin{itemize}
-  \item The \texttt{from\_id}* and \texttt{to\_id}* aesthetics each take one column which, when combined, provide the edges in a network.
-If the network is undirected, then their order does not matter.
-If the network is directed, then the \emph{from} vertex names should all be in the \texttt{from\_id} column, and their corresponding \emph{to} vertex names should all be in the \texttt{to\_id} column.
-  \item The \texttt{ecolour} aesthetic changes the color of the edges. Its default is \texttt{grey20}
-  \item The \texttt{esize} aesthetic changes the size (thickness) of the edges. Its default is 1. 
-  \item The \texttt{elinetype} aesthetic changes the linetype of the edges. Its default is 1. 
-  \item The \texttt{ealpha} aesthetic changes the transparency of the edges. Its default is 1. 
-  \item The \texttt{vcolour} aesthetic changes the color of the vertices. Its default is \texttt{black}. 
-  \item The \texttt{vfill} aesthetic changes the ouline color of the vertices. Its default is \texttt{black}. 
-  \item The \texttt{vsize} aesthetic changes the size (diameter) of the vertices. Its default is 2.
-  \item The \texttt{vshape} aesthetic changes the shape of the vertices. Its default is 16 (solid circular points). 
-  \item The \texttt{valpha} aesthetic changes the transparency of the vertices. Its default is 1. 
-  \end{itemize}
-
-\section{Future Work}
-As we discussed throughout the paper, there is still some work to be done.
-The primary change that we need to make is to add the capability of mapping aesthetic values to the plot based on variables directly.
-Right now, we have the ability to do this indirectly, but we want to add this functionality so that it is easier to change the \texttt{ecolor}, \texttt{esize}, \texttt{elinetype}, \texttt{vcolor}, \texttt{vfill}, \texttt{vsize}, and \texttt{vshape} by mapping variables to these aesthetics.
-We also want to include the legends for each of these aesthetics like we would normally see in \texttt{ggplot2}.
-We also hope to make it an official part of the \texttt{ggplot2} world, by either adding it to \texttt{ggplot2} itself or by adding it to the \texttt{GGally} package.
-We believe that this is an important contribution to \texttt{ggplot2} and to the R user community because it fills in gaps in the network plotting capabilities of R.
-Ultimately, our goal is to publish this paper in \textit{The R Journal}.
-
-\bibliographystyle{asa}
-\bibliography{references}
-
-\end{document}

### 2015-10-01 R/geom-net.r

@@ -1,0 +1,74 @@
+#' Geom for network visualization
+#'
+#' @inheritParams ggplot2::stat_identity
+#' @param na.rm If \code{FALSE} (the default), removes missing values with
+#'    a warning. If \code{TRUE} silently removes missing values.
+#'
+#' @export
+#' @examples
+#' library(ggplot2)
+#' data(blood)
+#' p <- ggplot(data = blood$edges, aes(from_id = from, to_id = to))
+#' p + geom_net()
+#' p + geom_net(vertices = blood$vertices)
+#' p + geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
+#' vlabel = TRUE, vsize = I(3), directed = TRUE) +
+#' expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+#'
+geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
+                        layout="kamadakawai", layout.par=list(), vertices=NULL, vlabel=FALSE, ...) {
+browser()
+    layer(
+    geom = GeomNet, mapping = mapping,  data = data, stat = stat,
+    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
+    params = list(layout=layout, layout.par=layout.par, vertices=vertices, vlabel=vlabel, ...)
+  )
+}
+
+#' @importFrom grid grobTree
+#' @export
+GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
+  required_aes = c("x", "y"),
+
+  default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
+                             shape = 19, colour = "black", size = 0.75, fill = NA,
+                             alpha = NA, stroke = 0.5, linewidth=1, angle=0),
+  draw_key = ggplot2::draw_key_point,
+
+  draw_panel = function(data, panel_scales, coord, vlabel=FALSE) {
+browser()
+    vertices <- data.frame(
+      data$vertices[[1]],
+      colour = data$colour[1],
+      shape = data$shape[1],
+      size = 5*data$size[1],
+      stroke = data$outlier.stroke[1] %||% data$stroke[1],
+      fill = NA,
+      alpha = NA,
+      stringsAsFactors = FALSE,
+      row.names=NULL)
+
+    if (any(vlabel)) {
[…]

### 2015-10-01 R/stat-net.r

@@ -1,0 +1,95 @@
+"%||%" <- function(a, b) {
+  if (!is.null(a)) a else b
+}
+
+#' @export
+StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
+  required_aes = c("from_id", "to_id"),
+  non_missing_aes = "weight",
+
+  setup_params = function(data, params) {
+    print(str(params))
+
+    params
+  },
+
+  setup_data = function(data, params) {
+    browser()
+  #  data <- remove_missing(data, na.rm, "y", name = "stat_net", finite = TRUE)
+
+    # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
+#     plyr::ddply(data, "group", transform,
+#                 ymin = min(y),
+#                 ymax = max(y),
+#                 xmin = x - width / 2,
+#                 xmax = x + width / 2
+#     )
+    data
+  },
+#  compute_group = function(data, scales, params, na.rm = FALSE,
+#                           layout="kamadakawai", layout.par=list(), vertices=NULL,  ...) {
+#    data
+#  },
+  compute_panel = function(self, data, scales, params, na.rm = FALSE,
+                           layout="kamadakawai", layout.par=list(), vertices=NULL) {
+browser()
+    edges <- data[,c('from_id', 'to_id')]
+    #3/17 - the next two lines are the source of the deletion of lone vertices.
+    net <- network::as.network(edges, matrix.type = "edgelist") #from network package
+    m <- network::as.matrix.network.adjacency(net)
+
+    #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
+
+    layoutFun <- paste('gplot.layout.',layout,sep='')
+    vert.coord <- data.frame(do.call(layoutFun, list(m, layout.par = layout.par)))
+
+    vert.coord$label <- row.names(m)
+    vert.coord$X1 <- scale(vert.coord$X1, center=min(vert.coord$X1), scale=diff(range(vert.coord$X1))) # center nodes
+    vert.coord$X2 <- scale(vert.coord$X2, center=min(vert.coord$X2), scale=diff(range(vert.coord$X2)))
+    names(vert.coord) <- c("x", "y", "label")
+
+    edgelist <- network::as.matrix.network.edgelist(net) #network pkg
+    edge.coord <- data.frame(vert.coord[edgelist[,1],], vert.coord[edgelist[,2],], row.names=NULL)
+    names(edge.coord) <- c('x','y', "from", 'xend','yend', "to")
[…]

### 2015-10-01 R/stat_dcommon.R

@@ -1,0 +1,61 @@
+StatDensityCommon <- ggplot2::ggproto("StatDensityCommon", ggplot2::Stat,
+  required_aes = "x",
+
+  setup_params = function(data, params) {
+    browser()
+    if (!is.null(params$bandwidth))
+      return(params)
+
+    xs <- split(data$x, data$group)
+    bws <- vapply(xs, bw.nrd0, numeric(1))
+    bw <- mean(bws)
+    message("Picking bandwidth of ", signif(bw, 3))
+
+    list(bandwidth = bw)
+  },
+
+  compute_group = function(data, scales, bandwidth = 1) {
+    d <- density(data$x, bw = bandwidth)
+    data.frame(x = d$x, y = d$y)
+  }
+)
+
+#' @export
+stat_density_common <- function(mapping = NULL, data = NULL, geom = "line",
+                                position = "identity", show.legend = NA,
+                                inherit.aes = TRUE, bandwidth = NULL,
+                                ...) {
+  layer(
+    stat = StatDensityCommon, data = data, mapping = mapping, geom = geom,
+    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
+    params = list(bandwidth = bandwidth, ...)
+  )
+}
+
+
+#' @export
+StatLm <- ggplot2::ggproto("StatLm", ggplot2::Stat,
+                  required_aes = c("x", "y"),
+
+                  compute_group = function(data, scales, params, n = 100, formix = y ~ x) {
+                    rng <- range(data$x, na.rm = TRUE)
+                    grid <- data.frame(x = seq(rng[1], rng[2], length = n))
+
+                    mod <- lm(formix, data = data)
+                    grid$y <- predict(mod, newdata = grid)
+
+                    grid
+                  }
+)
+
+#' @export
+stat_lm <- function(mapping = NULL, data = NULL, geom = "line",
+                    position = "identity", show.legend = NA,
+                    inherit.aes = TRUE, n = 50, formix = y ~ x,
+                    ...) {
+  layer(
+    stat = StatLm, data = data, mapping = mapping, geom = geom,
+    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
+    params = list(n = n, formix = formix, ...)
+  )
+}

### 2015-10-01 geom-net.r

@@ -1,93 +1,0 @@
-#' Geom for visualizing a network.
-#' #Documentation to be added.
-#' 
-
-geom_net <- function (mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...) {
-#browser()
-#  print("new net geom")
-  vmapping <- mapping[grep("^v", names(mapping))]
-  if (length(vmapping) > 0) {
-    names(vmapping) <- gsub("^v", "", names(vmapping))
-    p <- GeomBar$new(data=vertices, mapping = vmapping, ...)
-    vertices <- data.frame(p$data, p$compute_aesthetics(data=vertices, ggplot()))
-#browser()
- #   scales <- p$scales
-  #  vertices <- lapply(vertices, scales_transform_df, scales = scales)
-
-  }
-
-  idx <- grep("^v", names(mapping))
-  if (length(idx) > 0) mapping <- mapping[-idx]
-  GeomNet$new(vertices = vertices, mapping = mapping,
-              data = data, vlabel = vlabel, directed = directed, stat = stat,  ...)
-}
-
-GeomNet <- proto(Geom, {
-  objname <- "net"
-  #error: argument "coordinates" is missing, with no default
-  #draw <- function(., data, coordinates, ...) {
-  draw <- function(., data, vlabel = vlabel, vertices = vertices, directed = directed, ...) {
-#    print("draw net")
-
-    edges <- data.frame(data$edges[[1]])
-    #added 2/2 from geom_boxplot
-# only take the values from data, if there are no columns of the same name in edges
-    aes <- setdiff(c("ecolour", "elinetype", "ealpha", "esize"), names(edges))
-    for (x in aes) {
-      edges[,x] <- data[, x]
-    }
-    # get "e" out of the aesthetics' name
-    idx <- which(names(edges) %in% c("ecolour", "elinetype", "ealpha", "esize"))
-    names(edges)[idx] <- gsub("^e", "", names(edges)[idx])
-
-# need to be merged, but it's a bit tricky ...
-    if (!is.null(vertices))
-      vertices <- merge(data$vertices[[1]], vertices, by="label")
-    else vertices <- data$vertices[[1]]
-
-    aes <- setdiff(c("colour", "shape", "fill", "alpha", "size"), names(vertices))
-    for (x in aes) {
-      vertices[,x] <- data[, paste("v",x, sep="")]
-    }
-    vertices$group = data$group
[…]

### 2015-10-01 stat-net.r

@@ -1,73 +1,0 @@
-#' 2d density estimation.
-#' #Documentation to be added
-#' 
-#' 
-stat_net <- function (mapping = NULL, data = NULL, vertices = NULL,  geom = "net", position = "identity", ...) {
-#  print("new net stat")
-
-  StatNet$new(mapping = mapping, data = data, vertices = vertices, geom = geom,
-  position = position, ...)
-}
-
-StatNet <- proto(Stat, {
-  objname <- "net"
-  default_geom <- function(.) GeomNet
-  default_aes <- function(.) aes(colour = "black", size =3, layout="kamadakawai",layout.par = NA)
-  #required_aes <- c("x", "y")
-  #required_aes <- c("from_id", "to_id")
-
-  calculate_groups <- function(., data, coordinates=coordinates, vertices = vertices,  ...){
-#    print("calculate groups net")
-#    browser()
-
-    data$group <-1
-    .super$calculate_groups(., data, coordinates=coordinates, vertices = vertices, ...)
-  }
-
-#calculate function: extracts smaller data frame from original data using just req params. then computes some other stuff.
-  calculate <- function(., data, na.rm = FALSE,  coordinates=coordinates, vertices = vertices, ...) {  #add requirements here later.
-#    print("calculate net")
-#    browser()
-
-    #compute x and y using layout method if x,y are not there. then can assume that x and y are there from now on.
-    require(network)
-    require(sna) #how/where to call network, sna, pkg here?
-    stopifnot(is.data.frame(data))
-    edges <- data[,c('from_id', 'to_id')]
-    #3/17 - the next two lines are the source of the deletion of lone vertices.
-    net <- network::as.network(edges, matrix.type = "edgelist") #from network package
-    m <- network::as.matrix.network.adjacency(net)
-   # 3/17 attempt to fix issue. error: argument "vertices" is missing with no default. took out vertices above.
-   #  iso.vert.idx <- which(as.character(data$vertices[,1]) %in% rownames(m))  #require first column in df to be the names of vertices
-      #paste "layout name" instead of 'kamadakawai' (2/15)
-   # vert.coord <- data.frame(do.call(paste('sna::gplot.layout.',layout,sep=''), list(m, NULL)))
-  if (is.null(list(...)$layout))
-    layout <- .$default_aes()$layout
- else layout <- list(...)$layout
[…]

### 2015-10-01 R/geom-net.r
blood data: vertices have now label variable

@@ -11,6 +11,7 @@
 #' p <- ggplot(data = blood$edges, aes(from_id = from, to_id = to))
 #' p + geom_net()
 #' p + geom_net(vertices = blood$vertices)
+#' p + geom_net(vertices = blood$vertices, vlabel=blood$vertices$rho=="neg")
 #' p + geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
 #' vlabel = TRUE, vsize = I(3), directed = TRUE) +
 #' expand_limits(x = c(0,1), y = c(0,1)) + theme_net

### 2015-10-01 R/stat-net.r

@@ -39,6 +39,7 @@ browser()
     m <- network::as.matrix.network.adjacency(net)
 
     #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
+    require(sna)
 
     layoutFun <- paste('gplot.layout.',layout,sep='')
     vert.coord <- data.frame(do.call(layoutFun, list(m, layout.par = layout.par)))

### 2015-10-01 R/geom-net.r
defaults changed, parameters added

@@ -12,17 +12,20 @@
 #' p + geom_net()
 #' p + geom_net(vertices = blood$vertices)
 #' p + geom_net(vertices = blood$vertices, vlabel=blood$vertices$rho=="neg")
-#' p + geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
-#' vlabel = TRUE, vsize = I(3), directed = TRUE) +
-#' expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+#' p + geom_net(vertices = blood$vertices, vcolour = "orange", layout = 'circle',
+#' vlabel = TRUE, vsize = 6, directed = TRUE) +  theme_net
+#'
 #'
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
-                        layout="kamadakawai", layout.par=list(), vertices=NULL, vlabel=FALSE, ...) {
+                      layout="kamadakawai", layout.par=list(), vertices=NULL, vlabel=FALSE, vcolour = NULL, vsize=NULL, vshape=NULL,
+                      directed = FALSE, ...) {
 browser()
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
-    params = list(layout=layout, layout.par=layout.par, vertices=vertices, vlabel=vlabel, ...)
+    params = list(layout=layout, layout.par=layout.par, vertices=vertices,
+                  vlabel=vlabel, vcolour=vcolour, vsize=vsize, vshape=vshape,
+                  directed=directed, ...)
   )
 }
 
@@ -32,31 +35,41 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   required_aes = c("x", "y"),
 
   default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
-                             shape = 19, colour = "black", size = 0.75, fill = NA,
+                             shape = 19, colour = "grey60", size = 0.75, fill = NA,
                              alpha = NA, stroke = 0.5, linewidth=1, angle=0),
   draw_key = ggplot2::draw_key_point,
 
-  draw_panel = function(data, panel_scales, coord, vlabel=FALSE) {
[…]

### 2015-10-01 R/geom-net.r
for some strange reason ggplot2 things that characters are missings

@@ -15,7 +15,13 @@
 #' p + geom_net(vertices = blood$vertices, vcolour = "orange", layout = 'circle',
 #' vlabel = TRUE, vsize = 6, directed = TRUE) +  theme_net
 #'
-#'
+#' #Madmen Relationships
+#' data(madmen)
+#' p <- ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2))
+#' p + geom_net()
+#' p + geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, colour="grey30",
+#'              vcolour=c("#FF69B4", "#0099ff")[as.numeric(madmen$vertices$Gender)])
+
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
                       layout="kamadakawai", layout.par=list(), vertices=NULL, vlabel=FALSE, vcolour = NULL, vsize=NULL, vshape=NULL,
                       directed = FALSE, ...) {

### 2015-10-01 R/geom-net.r
madmen example

@@ -18,9 +18,9 @@
 #' #Madmen Relationships
 #' data(madmen)
 #' p <- ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2))
-#' p + geom_net()
-#' p + geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, colour="grey30",
-#'              vcolour=c("#FF69B4", "#0099ff")[as.numeric(madmen$vertices$Gender)])
+#' p + geom_net(vlabel= TRUE)
+#' p + geom_net(vertices = madmen$vertices, vsize=5, vlabel= TRUE, colour="grey30",
+#'              aes(vcolour=c("#FF69B4", "#0099ff")[as.numeric(madmen$vertices$Gender)]))
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
                       layout="kamadakawai", layout.par=list(), vertices=NULL, vlabel=FALSE, vcolour = NULL, vsize=NULL, vshape=NULL,
@@ -45,6 +45,12 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
                              alpha = NA, stroke = 0.5, linewidth=1, angle=0),
   draw_key = ggplot2::draw_key_point,
 
+  setup_data = function(data, params, mapping) {
+    browser()
+
+    data
+  },
+
   draw_panel = function(data, panel_scales, coord, vlabel=FALSE, vcolour=NULL, vsize=NULL, vshape=NULL, directed=FALSE) {
 browser()
     vertices <- data.frame(
@@ -62,7 +68,8 @@ browser()
     if (any(vlabel)) {
       if (length(vlabel) == 1) vlabel <- rep(vlabel, nrow(vertices))
       labels <- subset(vertices, vlabel == TRUE)
-      labels$x <- labels$x + 0.05
+      labels$y <- labels$y + 0.02
+      labels$vjust <- 1
       labels$angle <- data$outlier.angle[1] %||% data$angle[1]
       labels$colour="grey30"
 

### 2015-10-01 R/stat-net.r

@@ -57,7 +57,9 @@ browser()
 
     # vertices data set gets stuffed into the first element of each panel
     if (!is.null(vertices)) {
-      vert.coord <- merge(vert.coord, vertices, by="label")
+      vertices$.order <- 1:nrow(vertices)
+      vert.coord <- merge(vertices, vert.coord, by="label")
+      vert.coord <- vert.coord[order(vert.coord$.order),]
     }
 
 #     vert.coord$edges <- NA

### 2015-10-05 R/geom-net.r
added football example

@@ -81,6 +81,16 @@
 #' p + geom_net(layout="fruchtermanreingold")
 #' p + geom_net(layout="fruchtermanreingold", label=TRUE, vjust=-0.5)
 #' p + geom_net(layout="fruchtermanreingold", label=TRUE, vjust=-0.5, aes(linewidth=value/5))
+#'
+#' ## College Football Games in the Fall 2000 regular season
+#' # Source: http://www-personal.umich.edu/~mejn/netdata/
+#' data(football)
+#' ftnet <- merge(football$edges, football$vertices, by.x="from", by.y="label", all=TRUE)
+#' p <- ggplot(data=ftnet, aes(from_id=from, to_id=to))
+#' p + geom_net(aes(colour=value), linewidth=0.75, size=4.5, ecolour="grey80") +
+#'   scale_colour_brewer("Conference", palette="Paired") + theme_net() +
+#'   theme(legend.position="bottom")
+
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
                       layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", directed = FALSE, arrowsize=1,
                       labelcolour=NULL, ...) {
@@ -135,24 +145,24 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
       stringsAsFactors = FALSE
     )
     edges <- unique(subset(edges, !is.na(xend)))
-  
+
     arrow = NULL
     if (directed) arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="closed")
 
 #     vertices <- data.frame(
 #       x = unique(c(data$x, data$xend)),
 #       y = unique(c(data$y, data$yend)))
-#     vertices$colour = data$colour[unique(c(match(data$x, vertices$x), 
+#     vertices$colour = data$colour[unique(c(match(data$x, vertices$x),
 #                                            match(data$xend, vertices$x)))]
-#     vertices$shape = data$shape[unique(c(match(data$x, vertices$x), 
+#     vertices$shape = data$shape[unique(c(match(data$x, vertices$x),
 #                                           match(data$xend, vertices$x)))]
-#     vertices$size = data$size[unique(c(match(data$x, vertices$x), 
+#     vertices$size = data$size[unique(c(match(data$x, vertices$x),
 #                                          match(data$xend, vertices$x)))]
[…]

### 2015-10-05 R/theme-net.r

@@ -1,4 +1,4 @@
-#' theme_net 
+#' theme_net
 #' Themes set the general aspect of the plot such as the colour of the
 #' background, gridlines, the size and colour of fonts.
 #'
@@ -10,7 +10,7 @@
 #' \item{\code{theme_net}}{
 #' The usual ggplot2 theme with no background, no axes, and an aspect ratio of 1 for better
 #' viewing of networks and graphs.}
-#'
+#' }
 #' @examples
 #' library(ggplot2)
 #' data(blood)
@@ -18,16 +18,14 @@
 #' p <- ggplot(data = bloodnet, aes(from_id = from, to_id = to))
 #' p + geom_net()
 #' p + geom_net() + theme_net()
-#' 
+#'
 #' @name ggtheme
 NULL
-
-
 #' @export
 #' @rdname ggtheme
 theme_net <- function(base_size = 11, base_family = "") {
   half_line <- base_size / 2
-  
+
   theme(
     # Elements in this first block aren't used directly, but are inherited
     # by others
@@ -41,7 +39,7 @@ theme_net <- function(base_size = 11, base_family = "") {
       lineheight = 0.9, hjust = 0.5, vjust = 0.5, angle = 0,
       margin = margin()
     ),
-    
+
     axis.line =          element_blank(),
     axis.text =          element_blank(),
     axis.ticks =         element_blank(),
@@ -60,7 +58,7 @@ theme_net <- function(base_size = 11, base_family = "") {
     legend.direction =   NULL,
     legend.justification = "center",
     legend.box =         NULL,
-    
+
     panel.background =   element_blank(),
     panel.border =       element_blank(),
     panel.grid = element_blank(),
@@ -68,14 +66,14 @@ theme_net <- function(base_size = 11, base_family = "") {
     panel.margin.x =     NULL,
     panel.margin.y =     NULL,
     panel.ontop    =     FALSE,
-    
+
     strip.background =   element_rect(fill = "grey85", colour = NA),
     strip.text =         element_text(colour = "grey10", size = rel(0.8)),
     strip.text.x =       element_text(margin = margin(t = half_line, b = half_line)),
     strip.text.y =       element_text(angle = -90, margin = margin(l = half_line, r = half_line)),
     strip.switch.pad.grid = unit(0.1, "cm"),
     strip.switch.pad.wrap = unit(0.1, "cm"),
-    
+
     plot.background =    element_blank(),
     plot.title =         element_text(
       size = rel(1.2),

### 2015-10-05 inst/paper.Rnw
including geom_net version of examples into the paper

--- 2015-04-02 CC/writeup.Rnw
@@
 <<readingdata, eval=TRUE, echo=FALSE>>=
-library(ggplot2) # needs to be the version with geom_net!
- theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements, lesmis,madmen, mm_directed,email)
-names(blood$vertices)[1] <- 'label'
-#load("data/madmen.rda") # relative path should work for both of us
-#load("data/mm_directed.rda")
-#names(madmen$vertices)[1] <- 'label'
-names(theme_elements$vertices)[1] <- 'label'
-football$vertices$label <- as.factor(football$vertices$label)
-football$vertices$value <- as.factor(football$vertices$value)
-football$edges$in.conf <- 0
-for (i in 1:nrow(football$edges)){
-  idx1 <- which(football$vertices$label == football$edges$from[i])
-  idx2 <- which(football$vertices$label == football$edges$to[i])
-  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
-    football$edges$in.conf[i] <- 1
-  }
-}
-#load("data/email.rda")
-#email.net <- read.csv("data/email-data.csv")
-#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-#library(plyr)
-#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-#employee <- read.csv("data/EmployeeRecords.csv")
-#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
-email.net <- email$edges
-employee <- email$nodes
+library(ggplot2) # needs to be version ≥ 1.0.1.9003
+library(ggnet) # install from github
+data(blood, football, theme_elements, lesmis,madmen, mm_directed, email)
+
+# data(protein) # uncomment it once protein.rda is in data folder
 library(RColorBrewer)
 cols <- brewer.pal(8, name="Set1")
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
+yeast <- protein
+
+# set up networks for all of the data sets
+MMnet <- merge(madmen$edges, madmen$vertices, by.x="Name1", by.y="label", all=TRUE)
+bloodnet <- merge(blood$edges, blood$vertices, by.x="from", by.y="label", all=TRUE)
+
+employee <- data.frame(expand.grid(
+              label=unique(email$nodes$label), day=unique(email$edges$day)))
+employee <- merge(employee, email$nodes, by="label")
+emailnet <- merge(subset(email$edges, nrecipients < 54), employee, by.x=c("From", "day"), by.y=c("label", "day"), all=TRUE)
+
 @
 
+\hh{todo list:
+\begin{enumerate}
+\item there is no protein.rda in the data folder - could you please put it there?
+\item include installation instruction for package ggnet
+\end{enumerate}
+}
 \section{Introduction}
@@
 This network is shown in figure~\ref{fig.cap:madmen_ex}.
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net
+
+\begin{figure}[hbtp]
+\centering
+<<madmen_ex, dependson='readingdata', fig.cap = 'Network of Relationships in Mad Men', out.width='\\textwidth', fig.width=9, fig.height=9>>=
+ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(aes(colour=Gender), size=4, label= TRUE, vjust=-0.6,
+           ecolour="grey50") +
+  theme_net() +
+  scale_colour_manual(values=c( "#FF69B4", "#0099ff")) +
+  xlim(c(-0.05, 1.05)) + theme(legend.position="bottom")
 @
-
-\begin{figure}
-\centering
 \includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
@@
  %The \texttt{vcolor} aesthetic operates in the same way as the \texttt{vfill} aesthetic, but it changes the outline color of the vertex instead of the fill color of the vertex, which is in line with the \texttt{ggplot2} use of the \texttt{color} and \texttt{fill} aesthetics.
+
 The \texttt{vsize} aesthetic can also be set to an identity value to increase or decrease the size of all vertices in the graph, or it can be set to one of any numerical columns in the vertex data frame.
@@
 In the example below, loops are removed because loops exist on every vertex in this example, as blood between two people of matching ABO and RhD type can always be exchanged.
-<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
-           vlabel = TRUE, vsize = I(3), directed = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
+ggplot(data = bloodnet, aes(from_id = from, to_id = to)) +
+  geom_net(colour="red", layout="circle", label=TRUE, size=3,
+           directed=TRUE, vjust=-0.5, arrowsize=1.5) + theme_net()
 @
@@
 \centering
-\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
+\includegraphics[width=0.6\textwidth]{figure/blood_ex-1.pdf}
 \caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
@@
 In order to better visualize the structure of the communication network between employees, all emails that were sent out to all employees are removed in the subsequent examples.
-<<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + theme_net
-@
+%
 \begin{figure}
 \centering
+<<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE, out.width='\\textwidth'>>=
+ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
+  geom_net(aes(colour= CurrentEmploymentType),
+           linewidth=0.5, alpha=0.1, size=4) +
+  scale_colour_brewer(palette="Set1") + theme_net() +
+  theme(legend.position="bottom")
+@
 \includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
@@
 To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE,eval=FALSE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee, ealpha = 0.2, vsize = I(2), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net
-@
 \begin{landscape}
@@
 \centering
-\includegraphics[width=8in]{figure/email_ex_facet-1.pdf}
+%\includegraphics[width=8in]{figure/email_ex_facet-1.pdf}
+<<email_ex_facet, fig.height=5, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
+ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
+  geom_net(aes(colour= CurrentEmploymentType), fiteach=TRUE,
+           linewidth=0.5, alpha=0.2, size=2) +
+  scale_colour_brewer(palette="Set1") + theme_net() +
+  theme(legend.position="bottom") + facet_wrap(~day, nrow = 2)
+@
+\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
 \caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
@@
 \end{landscape}
-This plot is shown in figure \ref{fig.cap: email_ex_facet}.
+%' This plot is shown in figure \ref{fig.cap: email_ex_facet}.
 With the facetting, we can see that there are several days where one or more departments do not communicate with any of the other departments.
@@
 Facetting quickly separates dense networks into separate subnetworks for easy visual comparison and  analyses.
-  \subsection{\texttt{ggplot2} Theme Elements}
-This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
+%'   \subsection{\texttt{ggplot2} Theme Elements}
+%' This example comes from the \texttt{theme()} help page in the \texttt{ggplot2} documentation \citep{ggplot}.
 It is a directed network which shows the structure of the inheritance of theme options in the construction of a \texttt{ggplot2} plot.
@@
 Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.
-<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
-  expand_limits(x = c(0,1.2), y = c(0,1)) +
-  theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
-\caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
-\end{figure}
-\afterpage{\clearpage}
-
-The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
+%' <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
+%' ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
+%'   geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
+%'   expand_limits(x = c(0,1.2), y = c(0,1)) +
+%'   theme_net
+%' @
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
+%' \caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
+%' \end{figure}
+%' \afterpage{\clearpage}
+%'
+%' The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
 In this plot, it is easy to quickly determine the parent and child vertices.
@@
 We just made set each of the major parent elements, \texttt{text}, \texttt{rect}, and \texttt{line} to \texttt{element\_blank()} and then set the aspect ratio equal to one.
-  \subsection{Mad Men Networks}
-  The following code creates the network example given in the introduction.
+%'   \subsection{Mad Men Networks}
+%'   The following code creates the network example given in the introduction.
 We changed the vertex size and edge color for all vertices and edges, included vertex labels, and colored the vertices according to the character's gender.
-<<print_ex, ref.label="madmen_ex", eval=FALSE, echo=TRUE,cache=FALSE>>=
-@
- There is another Mad Men network included in the \texttt{gcookbook} package \citep{madmen}.
+%' <<print_ex, ref.label="madmen_ex", eval=FALSE, echo=TRUE,cache=FALSE>>=
+%' @
+%'  There is another Mad Men network included in the \texttt{gcookbook} package \citep{madmen}.
 It is a directed network, also of romantic relationships between characters, but it also includes advances made by one character that were rejected by the other.
@@
 If the advance is reciprocated, like between Sal Romano and the Bellhop, there are two directed edges between the two vertices.
-<<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
-ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
-           vsize = I(2.5), layout = 'fruchtermanreingold',
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/madmen2_ex-1.pdf}
-\caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
-\end{figure}
-This network is shown in figure \ref{fig.cap:madmen2_ex}.
+%' <<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
+%' ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
+%'   geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
+%'            vsize = I(2.5), layout = 'fruchtermanreingold',
+%'            aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
+%'   expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+%' @
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/madmen2_ex-1.pdf}
+%' \caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
+%' \end{figure}
+%' This network is shown in figure \ref{fig.cap:madmen2_ex}.
 This network is much more connected than the previous Mad Men example.
@@
 For instance, you can see that Roger made advances towards Betty, his business partner's wife, which is much more scandalous than what we can see in the first network.
-
-  \subsection{College Football}
-  This next example comes from M.E.J.
+%'
+%'   \subsection{College Football}
+%'   This next example comes from M.E.J.
 Newman's network data web page \citep{football}.
@@
 There is an additional variable in the vertex data frame corresponding to the conference each team belongs to, and there is an additional variable in the edge data frame that is equal to one if the game occured between teams in the same conference or zero if the game occured between teams in different conferences.
-<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-colors <- brewer.pal(12, name = 'Paired')
-ggplot(data = football$edges, aes(from_id = from, to_id = to,
-                                  elinetype = in.conf + 1)) +
-  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
-           aes(vcolour=colors[as.numeric(factor(value))])) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-The network of football games is given in figure \ref{fig.cap:football_ex}.
+%' <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
+%' colors <- brewer.pal(12, name = 'Paired')
+%' ggplot(data = football$edges, aes(from_id = from, to_id = to,
+%'                                   elinetype = in.conf + 1)) +
+%'   geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
+%'            aes(vcolour=colors[as.numeric(factor(value))])) +
+%'   expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+%' @
+%' The network of football games is given in figure \ref{fig.cap:football_ex}.
 Here, we have changed the \texttt{elinetype} aesthetic to correspond to games that occur between teams in the same conference or different conferences.
@@
 Without the coloring capability, we would not have been able to pick out that difference as easily.
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
-\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
-\end{figure}
-\afterpage{\clearpage}
-
-  \subsection{Les Mis\'{e}rables}
-  This next network comes from \citet{lesmis}.
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
+%' \caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
+%' \end{figure}
+%' \afterpage{\clearpage}
+%'
+%'   \subsection{Les Mis\'{e}rables}
+%'   This next network comes from \citet{lesmis}.
 It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.
@@
 There is an extremely wide edge connecting Valjean and Cosette, which implies that they are strongly connected; in fact, Cosette is Valjean's adopted daughter.
-<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to,
-                                esize = value/mean(value))) +
-  geom_net(vertices = lesmis$vertices, ealpha = 7/10,
-           vcolour = I("red"), vsize = I(2.5),
-           layout = 'fruchtermanreingold', vlabel=TRUE) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
-\caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
-\end{figure}
-\afterpage{\clearpage}
-
-  \subsection{Protein Interaction Network in Yeast}
-  This example of a protein interaction network comes from \citet{protein}.
+%' <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
+%' ggplot(data = lesmis$edges, aes(from_id = from, to_id = to,
+%'                                 esize = value/mean(value))) +
+%'   geom_net(vertices = lesmis$vertices, ealpha = 7/10,
+%'            vcolour = I("red"), vsize = I(2.5),
+%'            layout = 'fruchtermanreingold', vlabel=TRUE) +
+%'   expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+%' @
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
+%' \caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
+%' \end{figure}
+%' \afterpage{\clearpage}
+%'
+%'   \subsection{Protein Interaction Network in Yeast}
+%'   This example of a protein interaction network comes from \citet{protein}.
 It is the complete protein-protein interaction network in the yeast species \emph{S.
@@
 Indeed, we see a nearly round area in the middle of the graph where the layout parameter has forced there to be no vertices.
-
-<<yeast_code>>=
-sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
-@
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
-           vcolour = 'magenta', layout = 'random', 
-           layout.par = list(dist = 'uniang')) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
-\end{figure}
-\afterpage{\clearpage}
-
-\section{Technical Details}
-
-In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
+%'
+%' <<yeast_code>>=
+%' sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
+%' yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
+%' @
+%' <<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+%' ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
+%'   geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5,
+%'            vcolour = 'magenta', layout = 'random',
+%'            layout.par = list(dist = 'uniang')) +
+%'   expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+%' @
+%'
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
+%' \caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
+%' \end{figure}
+%' \afterpage{\clearpage}
+%'
+%'
+%' \section{Technical Details}
+%'
+%' In order to construct the \texttt{geom} in \texttt{ggplot2}, we needed two \texttt{R} script files: a `stat' file and a `geom' file.
 The stat file performs all the necessary calculations on the two input data frames, and the geom file performs the drawing of the network.
-  \subsection{Creating \texttt{stat\_net}}
-  In order to plot the network with only the edge connections and vertex names, we relied on the \texttt{sna} and \texttt{network} packages \citep{sna, network.jss}.
+%'   \subsection{Creating \texttt{stat\_net}}
+%'   In order to plot the network with only the edge connections and vertex names, we relied on the \texttt{sna} and \texttt{network} packages \citep{sna, network.jss}.
 First, we used the edges data frame to construct a network object with the \texttt{as.network()} function, then constructed an adjacency matrix from that network using the \texttt{as.matrix.network.adjacency()} function.
@@
 These two data frames, along with the many \texttt{aes()} and other argument options get passed on to the actual network geometry for drawing.
-   \subsection{\texttt{stat\_net} Usage}
-<<usage_stat_net, eval=FALSE>>=
-stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
-@
-  The usage of the \texttt{stat\_net} arguments is as follows:
-  \begin{itemize}
-  \item The \texttt{mapping} argument is the aesthetic mapping. Currently, there is no reason to ever change it from \texttt{NULL}.
-  \item The \texttt{data} argument is the edges data frame.  This is sent to \texttt{stat\_net} through the \texttt{ggplot()} function.
-  \item The \texttt{vertices} argument is the vertex data frame. This is sent to \texttt{stat\_net} through the \texttt{geom\_net} function.
-  \item The \texttt{geom} argument is the geometric object used to plot the data. It is by default set to \texttt{net}. 
-  \item The \texttt{position} argument is set to \texttt{identity} because we wish to plot the network as-is, with no adjustments.
-  \item na.rm
-  \end{itemize}
-  \texttt{stat\_net} also accepts the following aesthetic values, none of which are required:
-  \begin{itemize}
-  \item The \texttt{layout} aesthetic is the layout algorithm to use to place the vertices of the network.
+%'    \subsection{\texttt{stat\_net} Usage}
+%' <<usage_stat_net, eval=FALSE>>=
+%' stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
+%' @
+%'   The usage of the \texttt{stat\_net} arguments is as follows:
+%'   \begin{itemize}
+%'   \item The \texttt{mapping} argument is the aesthetic mapping. Currently, there is no reason to ever change it from \texttt{NULL}.
+%'   \item The \texttt{data} argument is the edges data frame.  This is sent to \texttt{stat\_net} through the \texttt{ggplot()} function.
+%'   \item The \texttt{vertices} argument is the vertex data frame. This is sent to \texttt{stat\_net} through the \texttt{geom\_net} function.
+%'   \item The \texttt{geom} argument is the geometric object used to plot the data. It is by default set to \texttt{net}.
+%'   \item The \texttt{position} argument is set to \texttt{identity} because we wish to plot the network as-is, with no adjustments.
+%'   \item na.rm
+%'   \end{itemize}
+%'   \texttt{stat\_net} also accepts the following aesthetic values, none of which are required:
+%'   \begin{itemize}
+%'   \item The \texttt{layout} aesthetic is the layout algorithm to use to place the vertices of the network.
 The default is the Kamada-Kawai algorithm, and there are other options listed in \ref{table:layouts}.
-  \item The \texttt{layout.par} aesthetic is the layout arguments to pass to the layout function.
+%'   \item The \texttt{layout.par} aesthetic is the layout arguments to pass to the layout function.
 It is set to \texttt{NA} by default, in which case the algorithm uses its set defaults in the \texttt{sna} package.
 This aesthetic takes a list of parameters according to the documentation by \citet{sna}.
-  \end{itemize}
-
-  \subsection{Creating \texttt{geom\_net}}
-  We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
+%'   \end{itemize}
+%'
+%'   \subsection{Creating \texttt{geom\_net}}
+%'   We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.
 Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.
@@
 This adds arrow arguments to the edges data frame which will put arrows pointing toward the ``to" vertex in a directed network corresponding to the \texttt{to\_id} value in the edges data frame.
-
-   \subsection{\texttt{geom\_net} Usage}
-<<usage_geom_net, eval=FALSE>>=
-geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
-@
-  The usage of the \texttt{geom\_net} arguments is as follows:
-  \begin{itemize}
-  \item The \texttt{mapping} argument is the aesthetic mapping. It is constructed with \texttt{aes()} according to the aesthetic values that can be passed to \texttt{geom\_net}. 
-  \item The \texttt{data} argument is the edges data frame.  There is usually no need to use it in \texttt{geom\_net()} because it will have already been provided in the \texttt{ggplot()} function.
-  \item The \texttt{vertices} argument is the vertex data frame.
+%'
+%'    \subsection{\texttt{geom\_net} Usage}
+%' <<usage_geom_net, eval=FALSE>>=
+%' geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
+%' @
+%'   The usage of the \texttt{geom\_net} arguments is as follows:
+%'   \begin{itemize}
+%'   \item The \texttt{mapping} argument is the aesthetic mapping. It is constructed with \texttt{aes()} according to the aesthetic values that can be passed to \texttt{geom\_net}.
+%'   \item The \texttt{data} argument is the edges data frame.  There is usually no need to use it in \texttt{geom\_net()} because it will have already been provided in the \texttt{ggplot()} function.
+%'   \item The \texttt{vertices} argument is the vertex data frame.
 The vertex data frame should be provided to this argument in \texttt{geom\_net}.
 This data frame is required to have a column named \texttt{label} providing the labels of all vertices in the network.
-  \item The \texttt{vlabel} argument is a logical value which plots the label layer on top of the edge and vertex layers when it is set to \texttt{TRUE}. 
-  \item The \texttt{directed} argument is a logical value which adds arrows to the edge layer pointing to the \texttt{to\_id} vertex when it is set to \texttt{TRUE}.  
-  \item The \texttt{stat} argument is the statistical transformation used on the data for the network layer. It is always set to \texttt{net}. 
-  \end{itemize}
-  \texttt{geom\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
-  \begin{itemize}
-  \item The \texttt{from\_id}* and \texttt{to\_id}* aesthetics each take one column which, when combined, provide the edges in a network.
+%'   \item The \texttt{vlabel} argument is a logical value which plots the label layer on top of the edge and vertex layers when it is set to \texttt{TRUE}.
+%'   \item The \texttt{directed} argument is a logical value which adds arrows to the edge layer pointing to the \texttt{to\_id} vertex when it is set to \texttt{TRUE}.
+%'   \item The \texttt{stat} argument is the statistical transformation used on the data for the network layer. It is always set to \texttt{net}.
+%'   \end{itemize}
+%'   \texttt{geom\_net} also accepts the following aesthetic values, and the ones marked with an asterisk are required:
+%'   \begin{itemize}
+%'   \item The \texttt{from\_id}* and \texttt{to\_id}* aesthetics each take one column which, when combined, provide the edges in a network.
 If the network is undirected, then their order does not matter.
 If the network is directed, then the \emph{from} vertex names should all be in the \texttt{from\_id} column, and their corresponding \emph{to} vertex names should all be in the \texttt{to\_id} column.
-  \item The \texttt{ecolour} aesthetic changes the color of the edges. Its default is \texttt{grey20}
-  \item The \texttt{esize} aesthetic changes the size (thickness) of the edges. Its default is 1. 
-  \item The \texttt{elinetype} aesthetic changes the linetype of the edges. Its default is 1. 
-  \item The \texttt{ealpha} aesthetic changes the transparency of the edges. Its default is 1. 
-  \item The \texttt{vcolour} aesthetic changes the color of the vertices. Its default is \texttt{black}. 
-  \item The \texttt{vfill} aesthetic changes the ouline color of the vertices. Its default is \texttt{black}. 
-  \item The \texttt{vsize} aesthetic changes the size (diameter) of the vertices. Its default is 2.
-  \item The \texttt{vshape} aesthetic changes the shape of the vertices. Its default is 16 (solid circular points). 
-  \item The \texttt{valpha} aesthetic changes the transparency of the vertices. Its default is 1. 
-  \end{itemize}
-
-\section{Future Work}
-As we discussed throughout the paper, there is still some work to be done.
+%'   \item The \texttt{ecolour} aesthetic changes the color of the edges. Its default is \texttt{grey20}
+%'   \item The \texttt{esize} aesthetic changes the size (thickness) of the edges. Its default is 1.
+%'   \item The \texttt{elinetype} aesthetic changes the linetype of the edges. Its default is 1.
+%'   \item The \texttt{ealpha} aesthetic changes the transparency of the edges. Its default is 1.
+%'   \item The \texttt{vcolour} aesthetic changes the color of the vertices. Its default is \texttt{black}.
+%'   \item The \texttt{vfill} aesthetic changes the ouline color of the vertices. Its default is \texttt{black}.
+%'   \item The \texttt{vsize} aesthetic changes the size (diameter) of the vertices. Its default is 2.
+%'   \item The \texttt{vshape} aesthetic changes the shape of the vertices. Its default is 16 (solid circular points).
+%'   \item The \texttt{valpha} aesthetic changes the transparency of the vertices. Its default is 1.
+%'   \end{itemize}
+%'
+%' \section{Future Work}
+%' As we discussed throughout the paper, there is still some work to be done.
 The primary change that we need to make is to add the capability of mapping aesthetic values to the plot based on variables directly.

### 2015-10-05 R/geom-net.r

@@ -72,6 +72,9 @@
 #' ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
 #'   geom_net(aes(colour= CitizenshipCountry), linewidth=0.5, fiteach=TRUE) + scale_colour_brewer(palette="Set2") +
 #'   facet_wrap(~day, nrow=2) + theme(legend.position="bottom")
+#' ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
+#'   geom_net(aes(colour= CurrentEmploymentType), linewidth=0.5, fiteach=FALSE) + scale_colour_brewer(palette="Set2") +
+#'   facet_wrap(~day, nrow=2) + theme(legend.position="bottom")
 #'
 #' ## Les Miserables example
 #'

### 2015-10-05 R/geom-net.r
exclude self references from drawing

@@ -135,6 +135,8 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
 
   draw_panel = function(data, panel_scales, coord,  ecolour="grey60",
                         directed=FALSE, arrowsize=1, label=FALSE, labelcolour=NULL) {
+
+    data$self <- data$to == data$from
     edges <- data.frame(
       x = data$x,
       xend = data$xend,
@@ -145,8 +147,11 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
       alpha = data$alpha,
       linetype=data$linetype,
       stroke = data$stroke,
+      self = data$self,
       stringsAsFactors = FALSE
     )
+    selfies <- subset(edges, self == TRUE)
+    edges <- subset(edges, self != TRUE) # what are we going to do with self references?
     edges <- unique(subset(edges, !is.na(xend)))
 
     arrow = NULL

### 2015-10-05 inst/paper.Rnw
all bot the protein example

--- 2015-04-02 CC/writeup.Rnw
@@
 <<readingdata, eval=TRUE, echo=FALSE>>=
-library(ggplot2) # needs to be the version with geom_net!
- theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements, lesmis,madmen, mm_directed,email)
-names(blood$vertices)[1] <- 'label'
-#load("data/madmen.rda") # relative path should work for both of us
-#load("data/mm_directed.rda")
-#names(madmen$vertices)[1] <- 'label'
-names(theme_elements$vertices)[1] <- 'label'
-football$vertices$label <- as.factor(football$vertices$label)
-football$vertices$value <- as.factor(football$vertices$value)
-football$edges$in.conf <- 0
+library(ggplot2) # needs to be version ≥ 1.0.1.9003
+library(ggnet) # install from github
+data(blood, football, theme_elements, lesmis,madmen, mm_directed, email)
+
+data(protein)
+library(RColorBrewer)
+cols <- brewer.pal(8, name="Set1")
+yeast <- protein
+
+# set up networks for all of the data sets
+MMnet <- merge(madmen$edges, madmen$vertices, by.x="Name1", by.y="label", all=TRUE)
+MM2net <- merge(mm.directed$edges, mm.directed$vertices, by.x="Name1", by.y="label", all=TRUE)
+bloodnet <- merge(blood$edges, blood$vertices, by.x="from", by.y="label", all=TRUE)
+
+employee <- data.frame(expand.grid(
+              label=unique(email$nodes$label), day=unique(email$edges$day)))
+employee <- merge(employee, email$nodes, by="label")
+emailnet <- merge(subset(email$edges, nrecipients < 54), employee, by.x=c("From", "day"), by.y=c("label", "day"), all=TRUE)
+
+TEnet <- merge(theme_elements$edges, theme_elements$vertices, by.x="parent", by.y="name", all=TRUE)
+
+football$edges$in.conf <- 1
 for (i in 1:nrow(football$edges)){
@@
   if (football$vertices$value[idx1] == football$vertices$value[idx2]){
-    football$edges$in.conf[i] <- 1
+    football$edges$in.conf[i] <- 0
   }
 }
-#load("data/email.rda")
-#email.net <- read.csv("data/email-data.csv")
-#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-#library(plyr)
-#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-#employee <- read.csv("data/EmployeeRecords.csv")
-#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
-email.net <- email$edges
-employee <- email$nodes
-library(RColorBrewer)
-cols <- brewer.pal(8, name="Set1")
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
+ftnet <- merge(football$edges, football$vertices, by.x="from", by.y="label", all=TRUE)
+
 @
 
+\hh{todo list:
+\begin{enumerate}
+\item there is no protein.rda in the data folder - could you please put it there?
+\item include installation instruction for package ggnet
+\end{enumerate}
+}
 \section{Introduction}
@@
 To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE,eval=FALSE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee, ealpha = 0.2, vsize = I(2), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net
-@
-\begin{landscape}
+%\begin{landscape}
 \begin{figure}
 \centering
-\includegraphics[width=8in]{figure/email_ex_facet-1.pdf}
+<<email_ex_facet, fig.height=5, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
+ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
+  geom_net(aes(colour= CurrentEmploymentType), fiteach=TRUE,
+           linewidth=0.5, alpha=0.2, size=2) +
+  scale_colour_brewer(palette="Set1") + theme_net() +
+  theme(legend.position="bottom") + facet_wrap(~day, nrow = 2)
+@
+\includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
 \caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
 \end{figure}
-\end{landscape}
+%\end{landscape}
 This plot is shown in figure \ref{fig.cap: email_ex_facet}.
@@
 <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
-  expand_limits(x = c(0,1.2), y = c(0,1)) +
-  theme_net
+ggplot(data = TEnet, aes(from_id = parent, to_id = child)) +
+  geom_net(directed = TRUE, label = TRUE, vjust=-.5) +
+  theme_net() + xlim(c(-0.05, 1.05))
 @
@@
 <<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
-ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
-           vsize = I(2.5), layout = 'fruchtermanreingold',
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+ggplot(data = MM2net, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(aes(colour=Gender), directed=TRUE, label=TRUE,
+           size=2.5, layout='fruchtermanreingold') +
+  scale_colour_manual(values=c( "#FF69B4", "#0099ff")) +
+  theme_net() + xlim(c(-.1, 1.1)) + theme(legend.position="bottom")
 @
@@
 <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-colors <- brewer.pal(12, name = 'Paired')
-ggplot(data = football$edges, aes(from_id = from, to_id = to,
-                                  elinetype = in.conf + 1)) +
-  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
-           aes(vcolour=colors[as.numeric(factor(value))])) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+sid <- which(ftnet$from %in% c("Navy", "NotreDame", "UtahState", "CentralFlorida", "Connecticut"))
+ftnet$schools <- ""
+ftnet$schools[sid] <- ftnet$from[sid]
+
+ggplot(data=ftnet, aes(from_id=from, to_id=to)) +
+  geom_net(aes(colour=value, linewidth=(2-in.conf)/3,  label=schools),
+           size=4, ecolour="grey30", vjust=-0.5,
+           alpha=0.3, layout='fruchtermanreingold') +
+  scale_colour_brewer("Conference", palette="Paired") + theme_net() +
+  theme(legend.position="bottom") + xlim(c(-0.05, 1.05))
 @
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to,
-                                esize = value/mean(value))) +
-  geom_net(vertices = lesmis$vertices, ealpha = 7/10,
-           vcolour = I("red"), vsize = I(2.5),
-           layout = 'fruchtermanreingold', vlabel=TRUE) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
+lesmisnet <- merge(lesmis$edges, lesmis$vertices, by.x="from", by.y="label", all=TRUE)
+
+ggplot(data = lesmisnet, aes(from_id = from, to_id = to, linewidth = value/5)) +
+  geom_net(alpha = 7/10, colour = "red", size = 2.5,
+           layout = 'fruchtermanreingold', label=TRUE) +
+  theme_net()
 @
@@
 
-  \subsection{Protein Interaction Network in Yeast}
-  This example of a protein interaction network comes from \citet{protein}.
+%'   \subsection{Protein Interaction Network in Yeast}
+%'   This example of a protein interaction network comes from \citet{protein}.
 It is the complete protein-protein interaction network in the yeast species \emph{S.
@@
 Indeed, we see a nearly round area in the middle of the graph where the layout parameter has forced there to be no vertices.
-
-<<yeast_code>>=
-sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
-@
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
-           vcolour = 'magenta', layout = 'random', 
-           layout.par = list(dist = 'uniang')) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
-\end{figure}
-\afterpage{\clearpage}
-
+%'
+%' <<yeast_code>>=
+%' sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
+%' yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
+%' @
+%' <<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+%' ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
+%'   geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5,
+%'            vcolour = 'magenta', layout = 'random',
+%'            layout.par = list(dist = 'uniang')) +
+%'   expand_limits(x = c(0,1), y = c(0,1)) + theme_net
+%' @
+%'
+%' \begin{figure}
+%' \centering
+%' \includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
+%' \caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
+%' \end{figure}
+%' \afterpage{\clearpage}
+%'
+%'
 \section{Technical Details}

### 2015-10-06 inst/paper.Rnw
start on paper edits

@@
 \begin{enumerate}
-\item there is no protein.rda in the data folder - could you please put it there?
 \item include installation instruction for package ggnet
+\item the biggest change to the previous implementation is that there is no longer a vertices data set.
+Instead, all of the vertex information has to be included in the edge dataset.
+The assumption is that any vertex information included relates to the `from\_id' column.
+If the user does not include ALL vertex information in the dataset (i.e.
+if some vertices only show up under the `to\_id' column), there will be segment ends without a dot in the graphic.
+As a measure against that, we suggest to merge edges and vertex information along the `from\_id' column using the parameter setting `all = TRUE'.
+This ensures all vertices to be included; if a vertex is a sink, the corresponding `to\_id' value is set to NA.
+This is accommodated for both in the layout and the drawing of the graph.
+\item the parameters have changed names: all of the vertex parameters have the regular names, edge parameters start with `e'. `esize' is re-named to `linewidth'.
+Some parameters might not be implemented yet, we need to get a good overview of what is missing and what works.
+\item we could draw selfies, but at the moment we don't. We need to discuss this.
 \end{enumerate}
 }
+\tableofcontents
+\newpage
 \section{Introduction}
@@
 
-%'   \subsection{Protein Interaction Network in Yeast}
-%'   This example of a protein interaction network comes from \citet{protein}.
+  \subsection{Protein Interaction Network in Yeast}
+  This example of a protein interaction network comes from \citet{protein}.
 It is the complete protein-protein interaction network in the yeast species \emph{S.
@@
 Indeed, we see a nearly round area in the middle of the graph where the layout parameter has forced there to be no vertices.
-%'
-%' <<yeast_code>>=
-%' sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-%' yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
-%' @
-%' <<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-%' ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-%'   geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5,
-%'            vcolour = 'magenta', layout = 'random',
-%'            layout.par = list(dist = 'uniang')) +
-%'   expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-%' @
-%'
-%' \begin{figure}
-%' \centering
-%' \includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-%' \caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
-%' \end{figure}
-%' \afterpage{\clearpage}
-%'
-%'
+
+<<yeast_code>>=
+yeast <- protein
+sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
+yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$from,sample.edges$to))))
+
+proteinnet <- merge(yeast.samp$vertices, yeast.samp$edges, by.x="label", by.y="from", all=TRUE)
+@
+<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+ggplot(data = proteinnet, aes(from_id = label, to_id = to)) +
+  geom_net(alpha = .25, colour = 'magenta', layout = 'random',
+           layout.par = list(dist = 'uniang')) +
+  theme_net()
+@
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
+\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
+\end{figure}
+\afterpage{\clearpage}
+
 \section{Technical Details}

### 2015-10-06 inst/paper.Rnw
alpha still not showing

@@
 ggplot(data = proteinnet, aes(from_id = label, to_id = to)) +
-  geom_net(alpha = .25, colour = 'magenta', layout = 'random',
+  geom_net(alpha=0.25, ealpha = .1, colour = 'magenta', layout = 'random',
            layout.par = list(dist = 'uniang')) +

### 2015-10-06 R/geom-net.r

@@ -2,7 +2,8 @@
 #'
 #' @inheritParams ggplot2::stat_identity
 #' @param stat character string of the network stat corresponding to geom_net.
-#' @param alpha numeric value of alpha blending of edges.
+#' @param alpha numeric value of alpha blending of vertices.
+#' @param ealpha numeric value of alpha blending of edges.
 #' @param na.rm If \code{FALSE} (the default), removes missing values with
 #'    a warning. If \code{TRUE} silently removes missing values.
 #' @param layout character value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?gplot.layout} in the package sna for other choices.
@@ -95,13 +96,13 @@
 #'   theme(legend.position="bottom")
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
-                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", directed = FALSE, arrowsize=1,
+                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, directed = FALSE, arrowsize=1,
                       labelcolour=NULL, ...) {
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
     params = list(layout=layout, layout.par=layout.par, fiteach=fiteach, label=label,
-                  ecolour = ecolour,  directed=directed,
+                  ecolour = ecolour, ealpha=ealpha, directed=directed,
                   arrowsize=arrowsize, labelcolour=labelcolour, ...)
   )
 }
@@ -113,8 +114,8 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
 
   default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
                              shape = 19, colour = "grey30",
-                             size = 4, fill = NA,
-                             alpha = NA, stroke = 0.5, linewidth=1, angle=0, vjust=0),
+                             size = 4, fill = NA, alpha = NA, stroke = 0.5,
+                             linewidth=1, angle=0, vjust=0),
 
   draw_key = function(data, params)  {
     with(data, grobTree(
[…]

### 2015-10-06 inst/paper.Rnw
warning added in case not all vertex information is included

@@
 \begin{enumerate}
+\item change format to comply with journal
 \item include installation instruction for package ggnet
@@
 
+\section{Our Implementation}
+\hh{this section needs a better name, but we need it, because here is where your work starts. it should not be part of the introduction.}
   \subsection{Data Structure}
@@
 
-\section{Usage Examples}
+\section{Examples of Networks}
 In this section, we demonstrate the current capabilities of \texttt{geom\_net} in a series of diverse examples.

### 2015-10-06 R/stat-net.r

@@ -26,6 +26,12 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
     data$to_id[is.na(data$to_id)] <- "..NA.."
     data$to_id <- factor(data$to_id, levels = c(levels, "..NA.."))
 
+    # check that all vertices are included in the data set
+    only_to <- setdiff(levels(data$to_id), levels(data$from_id))
+    only_to <- setdiff(only_to, "..NA..")
+    if (length(only_to) > 0)
+      warning(sprintf("There are %d nodes without node information: %s\n\nDid you use all=T in merge?\n\n", length(only_to), paste(only_to, collapse=", ")))
+
     if (fiteach) return(data)
 
     self$compute_network(data, layout=params$layout, layout.par=params$layout.par)

### 2015-10-06 inst/paper.Rnw
intro proof-read

@@
 For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.
-Additionally, \citet{networkfailures} examine node failure in interdependent networks like power grids.
+\citet{networkfailures} examine node failure in interdependent networks like power grids.
 Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C.
@@
 The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network.
-The world wide web is an example of a directed network because one webpage may link to another, and not necessarily the other way around.
-Undirected edges, however, are simply connections between vertices.
-Coauthorship networks that encode information about academic publications are examples of undirected networks because if two people author a paper together, that creates a connection between them that is bidirectional.
+The world wide web is an example of a directed network because one webpage may link to another, but not necessarily the other way around.
+Undirected edges are simply connections between vertices.
+In co-authorship networks nodes are authors connected by an edge, if they author an academic publication together.
+Co-authorship networks are examples of undirected networks because if two people author a paper together, it creates a connection between them that is bidirectional.
 

### 2015-10-06 inst/paper.Rnw
edits through intro

@@
 MMnet <- merge(madmen$edges, madmen$vertices, by.x="Name1", by.y="label", all=TRUE)
+levels(MMnet$Gender) <- c("female", "male")
 MM2net <- merge(mm.directed$edges, mm.directed$vertices, by.x="Name1", by.y="label", all=TRUE)
+levels(MM2net$Gender) <- c("female", "male")
+
 bloodnet <- merge(blood$edges, blood$vertices, by.x="from", by.y="label", all=TRUE)
@@
 To demonstrate the functionality of our geometry for plotting networks, we have chosen an example of a social network from the popular television show Mad Men.
-This network was compiled in \citet{madmen}.
-In this example network, there are 52 vertices and 87 edges.
-Each vertex is a character on the show, and there is an edge between every two characters who have had a romantic relationship.
+This network, compiled by \citet{madmen}, is made up of 52 vertices and 87 edges.
+Each vertex represents a character on the show, and there is an edge between every two characters who have had a romantic relationship.
 
-This network is shown in figure~\ref{fig.cap:madmen_ex}.
+Figure~\ref{fig.cap:madmen_ex} shows this network.
+In the plot, we can see one central character who has many more relationships than any other character.
+This vertex represents the main character of the show, Don Draper, who is quite the ``ladies' man."  This example shows just how ubiquitous networks are.
+%% if we have to say that something is fun, it no longer is ...
 
@@
 
-In the plot, we can see one central character who has many more relationships than any other character.
-This vertex represents the main character of the show, Don Draper, who is quite the ``ladies' man."  This fun example shows just how ubiquitous networks are.
+% Network analysis is important in many fields
+\hh{include at least one citation for each one of the examples:}
 \par There are many kinds of networks, and networks are extensively studied across many disciplines.
@@
 As different as these and the many other disciplines that study networks are, they all need the ability to quickly and effectively visualize networks.
+\hh{quickly and effectively are a bit vague as measures for why we need a geom implementation.
+We don't want to do any studies measuring time to result or define effectiveness.
+It might be better to give some concrete examples of packages doing network layouts and discuss the problematic that there is no standard way of bundling up the results, which makes working with them hard, and even harder to modify output to include additional information for visualizations.
+}
+
+% Visualization of networks helps with analysis
+Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
+The vertices in a network and the edges between them compose the structure of a network, and being able to discover patterns among them visually is a key part of network analysis.
+Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.
+
+% Many network packages exist, but are not easy to use
+Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by \citet{igraph}, \texttt{sna} by \citet{sna}, and \texttt{network} by \citet{network.jss, network} but we have found these packages to have unintuitive or burdensome \hh{you're shooting sharp -- it might be better to describe that our approach is more customizable and more intuitive to use} methods for customizing the colors, sizes, etc of the vertices and edges of the network.
+For instance, the \texttt{igraph} package allows for coloring vertices by groups but the user must assign the colors to each vertex individually as opposed to assigning color by a grouping factor variable.
+
+\hh{just moving your words around}
+
+% ggplot2 as an implementation framework
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
@@
 
-\par The three necessary elements of any network visualization are the vertices, the edges, and the layout.
-But once those three items are visualized, we usually find our visualization to be lacking.
-We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.
-Many \texttt{R} packages already exist for network analysis and visualization such as \texttt{igraph} by \citet{igraph}, \texttt{sna} by \citet{sna}, and \texttt{network} by \citet{network.jss, network} but we have found these packages to have unintuitive or burdensome methods for customizing the colors, sizes, etc of the vertices and edges of the network.
-For instance, the \texttt{igraph} package allows for coloring vertices by groups but the user must assign the colors to each vertex individually as opposed to assigning color by a grouping factor variable.
-\\
+\hh{There are two main approaches to making use of the ggplot2 framework: (i) implement network visualizations using ggplot2, i.e.
+providing a wrapper for the user to visualize a network with ggplot2 elements, and (ii) implement networks as an internal layer of ggplot2.
+We will discuss both of these approaches in this paper.
+}
 
+\hh{XXX Provide roadmap for the rest of the paper. }
+
+\par The three necessary elements of any network visualization are the vertices, the edges, and the layout.   \\
+%But once those three items are visualized, we usually find our visualization to be lacking.
+\hh{pull it over to a positive statement}  We may want to color the vertices (or edges) by some sort of grouping variable, or we may want to make vertices of degree ten twice the size of vertices of degree five.
+
+\hh{XXXX not sure where this next paragraph should go - but we need a more detailed discussion of ggnet anyways.}
 \par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
-Coloring the vertices or edges in a graph is a quick and easy way to visualize grouping and can help with pattern or cluster detection.
-The vertices in a network and the edges between them compose the structure of a network, and being able to discover patterns among them visually is a key part of network analysis.
-Viewing multiple layouts of the same network can also help reveal patterns or clusters that would not be discovered when only viewing one layout or analyzing only an adjacency matrix.
  The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.
 Our work builds off of the \texttt{ggnet} function and presents an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.
+ \hh{XXXX}
 
-\section{Our Implementation}
+\section{Two approaches of network visualizations}
+
 \hh{this section needs a better name, but we need it, because here is where your work starts. it should not be part of the introduction.}

### 2015-10-06 inst/paper.Rnw
parameter added to stop drawing lines early

@@
 @
+
 <<readingdata, eval=TRUE, echo=FALSE>>=

### 2015-10-06 R/geom-net.r

@@ -16,6 +16,7 @@
 #' @param directed logical value. Should an arrow be drawn from 'from' to 'to' node?
 #' @param arrow what kind of arrow should be drawn? See specification of function \code{arrow} in grid package
 #' @param arrowsize numeric value (non-negative). How big should the arrow be drawn? Multiplicative of a pre-specified unit.
+#' @param epullback numeric value between 0 and 1 specifying how much (as a proportion of the line length) earlier the line segment should be stopped drawing before reaching the target node. This parameters is only regarded in directed networks.
 #'
 #' @export
 #' @examples
@@ -26,6 +27,7 @@
 #' p + geom_net()
 #' p + geom_net(aes(colour=rho))
 #' p + geom_net(aes(colour=rho), label=TRUE)
+#' p + geom_net(aes(colour=rho), label=TRUE, directed=TRUE, epullback=0.02)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 0, linewidth=.75, directed = TRUE)
@@ -96,13 +98,13 @@
 #'   theme(legend.position="bottom")
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
-                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, directed = FALSE, arrowsize=1,
+                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, epullback=0, directed = FALSE, arrowsize=1,
                       labelcolour=NULL, ...) {
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
     params = list(layout=layout, layout.par=layout.par, fiteach=fiteach, label=label,
-                  ecolour = ecolour, ealpha=ealpha, directed=directed,
+                  ecolour = ecolour, ealpha=ealpha, epullback=epullback, directed=directed,
                   arrowsize=arrowsize, labelcolour=labelcolour, ...)
   )
 }
[…]

### 2015-10-06 R/geom-net.r
fiddling with the examples a bit

@@ -27,7 +27,7 @@
 #' p + geom_net()
 #' p + geom_net(aes(colour=rho))
 #' p + geom_net(aes(colour=rho), label=TRUE)
-#' p + geom_net(aes(colour=rho), label=TRUE, directed=TRUE, epullback=0.02)
+#' p + geom_net(aes(colour=rho), label=TRUE, labelcolour="black", directed=TRUE, epullback=0.03)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 0, linewidth=.75, directed = TRUE)
@@ -117,7 +117,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
                              shape = 19, colour = "grey30",
                              size = 4, fill = NA, alpha = NA, stroke = 0.5,
-                             linewidth=1, angle=0, vjust=0),
+                             linewidth=1, angle=0, vjust=-0.5),
 
   draw_key = function(data, params)  {
     with(data, grobTree(

### 2015-10-07 R/geom-net.r
relative to absolute arrowgap

@@ -16,7 +16,7 @@
 #' @param directed logical value. Should an arrow be drawn from 'from' to 'to' node?
 #' @param arrow what kind of arrow should be drawn? See specification of function \code{arrow} in grid package
 #' @param arrowsize numeric value (non-negative). How big should the arrow be drawn? Multiplicative of a pre-specified unit.
-#' @param epullback numeric value between 0 and 1 specifying how much (as a proportion of the line length) earlier the line segment should be stopped drawing before reaching the target node. This parameters is only regarded in directed networks.
+#' @param arrowgap numeric value between 0 and 1 specifying how much (as a proportion of the line length) earlier the line segment should be stopped drawing before reaching the target node. This parameters is only regarded in directed networks.
 #'
 #' @export
 #' @examples
@@ -27,7 +27,7 @@
 #' p + geom_net()
 #' p + geom_net(aes(colour=rho))
 #' p + geom_net(aes(colour=rho), label=TRUE)
-#' p + geom_net(aes(colour=rho), label=TRUE, labelcolour="black", directed=TRUE, epullback=0.03)
+#' p + geom_net(aes(colour=rho), label=TRUE, vjust=-0.5, labelcolour="black", directed=TRUE, arrowgap=0.01) + theme_net()
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 0, linewidth=.75, directed = TRUE)
@@ -42,6 +42,8 @@
 #' p + geom_net(label=TRUE)
 #' p + geom_net(aes(colour=Gender), size=6, linewidth=1, label=TRUE, fontsize=3, labelcolour="black")
 #' p + geom_net(aes(colour=Gender), size=6, linewidth=1, label=TRUE, labelcolour="black") +
+#'     scale_colour_manual(values=c("#FF69B4", "#0099ff")) + xlim(c(-.05,1.05))
+#' p + geom_net(aes(colour=Gender), size=6, linewidth=1, directed=TRUE, label=TRUE, arrowgap=0.01, labelcolour="black") +
 #'     scale_colour_manual(values=c("#FF69B4", "#0099ff")) + xlim(c(-.05,1.05))
 #'
 #' p <- ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2))
@@ -98,13 +100,13 @@
[…]

### 2015-10-07 inst/paper.Rnw
added Francois
paper structure

@@
 
-\title{A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}}
-\author{Sam Tyner and Heike Hofmann}
+\title{Network Visualizations in \MakeLowercase{\texttt{ggplot2}}}
+\author{Sam Tyner and Fran\c{c}ois Briatte and  Heike Hofmann}
 \date{} % delete this line to display the current date
@@
 
-\par The three necessary elements of any network visualization are the vertices, the edges, and the layout.   \\
+\section{Two approaches of network visualizations}
+ The three necessary elements of any network visualization are the vertices, the edges, and the layout.
 %But once those three items are visualized, we usually find our visualization to be lacking.
@@
 
-\hh{XXXX not sure where this next paragraph should go - but we need a more detailed discussion of ggnet anyways.}
+\subsection{ggnet}
+\hh{XXX needs massive changes :)}
 \par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
@@
 
-\section{Two approaches of network visualizations}
-
-\hh{this section needs a better name, but we need it, because here is where your work starts. it should not be part of the introduction.}
-  \subsection{Data Structure}
+\subsection{geom\_net}
+  \subsubsection{Data Structure}
 In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
@@
 
-  \subsection{Vertex Aesthetics}
+  \subsubsection{Vertex Aesthetics}
 
@@
 \par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion.
-  \subsection{Edge Aesthetics}
+  \subsubsection{Edge Aesthetics}
   Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for segments.
@@
 
-\subsection{Other Arguments}
+\subsubsection{Other Arguments}
   Similar to the other geometries and functions in \texttt{ggplot2}, our geometry takes several arguments outside of the aesthetic mapping parameters.

### 2015-10-08 R/stat-net.r
allow layout to be NULL - then (x, y) have to be specified

@@ -38,12 +38,19 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
   },
 
 compute_network = function(data, layout="kamadakawai", layout.par=list()) {
-#  browser()
   edges <- unique(subset(data, to_id != "..NA..")[,c('from_id', 'to_id')])
 
   net <- network::as.network(edges, matrix.type = "edgelist") #from network package
   m <- network::as.matrix.network.adjacency(net)
 
+  if (is.null(layout)) {
+    vert.coord <- data[, c("x", "y", "from_id")]
+    vert.coord <- subset(vert.coord, from_id %in% row.names(m))
+    vert.coord <- unique(vert.coord)
+    vert.coord$x <- as.numeric(scale(vert.coord$x, center=min(vert.coord$x), scale=diff(range(vert.coord$x))))
+    vert.coord$y <- as.numeric(scale(vert.coord$y, center=min(vert.coord$y), scale=diff(range(vert.coord$y))))
+    names(vert.coord)[3] <- "label"
+  } else {
   #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
   require(sna)
   layoutFun <- paste('gplot.layout.',layout,sep='')
@@ -53,15 +60,17 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
   vert.coord$X1 <- scale(vert.coord$X1, center=min(vert.coord$X1), scale=diff(range(vert.coord$X1))) # center nodes
   vert.coord$X2 <- scale(vert.coord$X2, center=min(vert.coord$X2), scale=diff(range(vert.coord$X2)))
   names(vert.coord) <- c("x", "y", "label")
-
+}
   edgelist <- network::as.matrix.network.edgelist(net) #network pkg
+
   edge.coord <- data.frame(vert.coord[edgelist[,1],], vert.coord[edgelist[,2],], row.names=NULL)
   names(edge.coord) <- c('x','y', "from", 'xend','yend', "to")
 
-  fromto <- subset(data, to_id != "..NA..")
+  relVars <- setdiff(names(data), c("x", "y"))
+  fromto <- subset(data, to_id != "..NA..")[,relVars]
   edges <- merge(edge.coord, fromto, by.x=c("from", "to"), by.y=c("from_id", "to_id"), all=TRUE)
 
-  fromonly <- subset(data, to_id == "..NA..")
[…]

### 2015-10-08 inst/paper.Rnw
added bike example

@@
 library(ggnet) # install from github
-data(blood, football, theme_elements, lesmis,madmen, mm_directed, email)
+data(blood, football, theme_elements, lesmis,madmen, mm_directed, email,
+     bikes)
 
@@
 
+tripnet <- merge(bikes$trips, bikes$stations, by.x="Start.station", by.y="name", all=TRUE)
 @
@@
 <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
+library(dplyr)
+TEnet <- TEnet %>% group_by(parent) %>% mutate(degree=sqrt(10*n()+1))
+
 ggplot(data = TEnet, aes(from_id = parent, to_id = child)) +
-  geom_net(directed = TRUE, label = TRUE, vjust=-.5) +
+  geom_net(aes(fontsize=degree), directed = TRUE, label = TRUE, vjust=-.5,
+           size=3) +
   theme_net() + xlim(c(-0.05, 1.05))
@@
 \caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
+\end{figure}
+\afterpage{\clearpage}
+
+\subsection{Bikesharing in D.C.}
+The data shows the second quartal trips in 2015 taken with bikes from the bike share company https://secure.capitalbikeshare.com/.
+While this bikesharing company is located in the heart of Washington D.C.\ they do have a set of bike stations just outside of Washington in  Rockville, MD and north of it.
+Each station is shown as a dot, lines between stations indicate that at least five trips were taken between these two stations; the wider the line, the more trips have been taken between stations.
+In order to reflect distance between stations, we use as an additional restriction that the fastest trip was at most ten minutes long.
+
+<<bikes_ex, echo=TRUE, fig.cap="Bikesharing in Rockville", warning=FALSE, fig.keep='all'>>=
+tripnet$Metro = FALSE
+idx <- grep("Metro", tripnet$Start.station)
+tripnet$Metro[idx] <- TRUE
+
+ggplot(aes(from_id=Start.station, to_id=End.station), data=tripnet) +
+  geom_net(aes(linewidth=n/15, colour=Metro), label=TRUE, vjust=-0.5) +
+  theme_net() + xlim(c(-0.1, 1.1)) +
+  scale_colour_manual(values=c("grey40", "darkorange"))
+# compare to https://secure.capitalbikeshare.com/map/
+
+ggplot(aes(from_id=Start.station, to_id=End.station),  data=tripnet) +
+  geom_net(layout=NULL, label=TRUE, vjust=-0.5,
+           aes(x=long, y=lat, linewidth=n/15, colour=Metro)) +
+  theme_net() +
+    scale_colour_manual(values=c("grey40", "darkorange"))
+
+@
+
+\begin{figure}
+\centering
+\includegraphics[width=.49\textwidth]{figure/bikes_ex-1.pdf}
+\includegraphics[width=.49\textwidth]{figure/bikes_ex-2.pdf}
+\caption{\label{fig:bikes} Network of bike trips using a Kamada-Kawai layout (left) and a geographically true representation (right).
+Metro stations are shown in orange.
+In the Kamada-Kawai layout based on the trips taken, metro stations take a much more central position than in the geographically true representation.
+}
 \end{figure}

### 2015-10-08 R/geom-net.r

@@ -100,7 +100,7 @@
 #'   theme(legend.position="bottom")
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
-                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, arrowgap=0, directed = FALSE, arrowsize=1,
+                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, arrowgap=0.01, directed = FALSE, arrowsize=1,
                       labelcolour=NULL, ...) {
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
@@ -138,11 +138,11 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
     data
   },
 
-  draw_panel = function(data, panel_scales, coord,  ecolour="grey60", ealpha=NULL, arrowgap=0,
+  draw_panel = function(data, panel_scales, coord,  ecolour="grey60", ealpha=NULL, arrowgap=0.01,
                         directed=FALSE, arrowsize=1, label=FALSE, labelcolour=NULL) {
 
 #    browser()
-    data$self <- data$to == data$from
+    data$self <- as.character(data$to) == as.character(data$from)
     edges <- data.frame(
       x = data$x,
       xend = data$xend,
@@ -171,22 +171,6 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
 
     }
 
-#     vertices <- data.frame(
-#       x = unique(c(data$x, data$xend)),
-#       y = unique(c(data$y, data$yend)))
-#     vertices$colour = data$colour[unique(c(match(data$x, vertices$x),
-#                                            match(data$xend, vertices$x)))]
-#     vertices$shape = data$shape[unique(c(match(data$x, vertices$x),
-#                                           match(data$xend, vertices$x)))]
-#     vertices$size = data$size[unique(c(match(data$x, vertices$x),
-#                                          match(data$xend, vertices$x)))]
-#     vertices$stroke = data$stroke[unique(c(match(data$x, vertices$x),
-#                                            match(data$xend, vertices$x)))]
-#     vertices$fill = data$fill[unique(c(match(data$x, vertices$x),
-#                                          match(data$xend, vertices$x)))]
[…]

### 2015-10-08 R/stat-net.r
error message if layout is NULL and no (x,y) coordinates are given

@@ -44,6 +44,7 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
   m <- network::as.matrix.network.adjacency(net)
 
   if (is.null(layout)) {
+    if (is.null(data$x) || is.null(data$y)) stop("If no layout mechanism is specified, x and y coordinates have to be given\n\n")
     vert.coord <- data[, c("x", "y", "from_id")]
     vert.coord <- subset(vert.coord, from_id %in% row.names(m))
     vert.coord <- unique(vert.coord)

### 2015-10-09 R/geom-net.r
added vertices parameter

@@ -22,6 +22,9 @@
 #' @examples
 #' library(ggplot2)
 #' data(blood)
+#' p <- ggplot(data = blood$edges, aes(from_id = from, to_id = to))
+#' p + geom_net(vertices=blood$vertices, aes(colour=..type..))
+#'
 #' bloodnet <- merge(blood$edges, blood$vertices, by.x="from", by.y="label", all=TRUE)
 #' p <- ggplot(data = bloodnet, aes(from_id = from, to_id = to))
 #' p + geom_net()
@@ -101,13 +104,13 @@
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
                       layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, arrowgap=0.01, directed = FALSE, arrowsize=1,
-                      labelcolour=NULL, ...) {
+                      labelcolour=NULL, vertices=NULL, ...) {
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
     params = list(layout=layout, layout.par=layout.par, fiteach=fiteach, label=label,
                   ecolour = ecolour, ealpha=ealpha, arrowgap=arrowgap, directed=directed,
-                  arrowsize=arrowsize, labelcolour=labelcolour, ...)
+                  arrowsize=arrowsize, labelcolour=labelcolour, vertices=vertices,  ...)
   )
 }
 

### 2015-10-09 R/stat-net.r

@@ -10,6 +10,7 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
   non_missing_aes = "weight",
 
   setup_params = function(data, params) {
+#    browser()
 #    print(str(params))
 
     params
@@ -19,6 +20,10 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
     fiteach=params$fiteach
     if (!is.factor(data$from_id)) data$from_id <- factor(data$from_id)
     if (!is.factor(data$to_id)) data$to_id <- factor(data$to_id)
+
+    if (!is.null(params$vertices)) {
+      data <- merge(data, params$vertices, by.x="from_id", by.y="label", all=TRUE)
+    }
 
     # we want to keep all of the values that are NA in the second edge - give them a special value, so we can pull them out later
     levels <- levels(data$to_id)
@@ -84,7 +89,7 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
   edges
 },
   compute_panel = function(self, data, scales, params, na.rm = FALSE,
-                           layout="kamadakawai", layout.par=list(), fiteach=FALSE) {
+                           layout="kamadakawai", layout.par=list(), fiteach=FALSE, vertices=NULL) {
 
     if (fiteach) return(self$compute_network(data, layout=layout, layout.par=layout.par))
 
@@ -100,13 +105,13 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
 #' @export
 stat_net <- function(mapping = NULL, data = NULL, geom = "point",
                      position = "identity", show.legend = NA,
-                     inherit.aes = TRUE, layout="kamadakawai", layout.par=list(), fiteach=FALSE,
+                     inherit.aes = TRUE, layout="kamadakawai", layout.par=list(), fiteach=FALSE, vertices=NULL,
                      na.rm=FALSE, ...) {
   layer(
     stat = StatNet, data = data, mapping = mapping, geom = geom, position = position,
     show.legend = show.legend, inherit.aes = inherit.aes,
     params = list(layout=layout, layout.par=layout.par, fiteach=fiteach,
-                  na.rm=na.rm, ...
+                  na.rm=na.rm, vertices=vertices, ...
     )
   )
 }

### 2015-10-09 inst/paper.Rnw
start of geom_net section

@@
 \newcommand{\sctq}[1]{{\color{red} #1}} %things in red are questions for you or concerns that i have.
+
+%just until we switch to the R journal style file
+\newcommand{\code}[1]{{\tt{ #1}}}
 
@@
 \subsubsection{Data Structure}
-In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
+\hh{Network analysis is usally working with two sources of information: one data set consisting of a description of the `players', represented as the nodes or vertices in the network, and another data set detailing the relationship between the players, the edge dataset.
+In order for this geometry to work, these two data sets need to be combined into a single dataset.
+For this, we are using the convention that all of the node information is merged into the edge data set using the `from' as the reference column.
+Generally, there will be some vertices that are sinks in the network, ie.\ they only show up in the `to' column.
+We can easily accommodate for these nodes by adding artificial edges in the data set, that have missing information for the `to' column.
+Fortunately, R provides functions that allow for an easy way of producing the required result: both \code{merge} and \code{join} can be used.
+In \code{merge} the paramter all needs to be set to TRUE, in \code{join}, the parameter setting has to be \code{type=`full'}, and \code{all=TRUE}.
+
+}
+\hh{The formal requirement of  \code{stat\_net} are two columns, called \code{from\_id} and \code{to\_id}.
+During this routine, columns \code{x, y} and \code{xend, yend} are calculated and used as a required input for \code{geom\_net}.}
+% In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information.
 The vertex data frame should contain all the relevant vertex information.
@@
 Other values of interest, such as grouping variables or degree of each vertex should be stored as columns in this dataframe, with an observation for each vertex.
-  \par The edge data frame should contain all the relevant edge information.
+%   \par The edge data frame should contain all the relevant edge information.
 The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network.
@@
 The formal names for these columns are \texttt{from\_id} and \texttt{to\_id}, respectively.
+
 Other variables may also be included for each edge, such as the edge weight or grouping variable.
-As before, the variables of interest are columns in the data frame and the rows are each edge in the network.
+%As before, the variables of interest are columns in the data frame and the rows are each edge in the network.
 
@@
 \includegraphics[width=.49\textwidth]{figure/bikes_ex-2.pdf}
-\caption{\label{fig:bikes} Network of bike trips using a Kamada-Kawai layout (left) and a geographically true representation (right).
+\caption{\label{fig:bikes} Network of bike trips using a Kamada-Kawai layout (left) and a geographically true representation (right, \hh{XXX I would like to include a ggmap here, we need to wait until that package is fixed}).
 Metro stations are shown in orange.

### 2015-10-10 inst/paper.Rnw
side by side examples

@@
 \usepackage{pdflscape}
+\usepackage[font=small,skip=5pt]{caption}
+\usepackage{subcaption}
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 library(ggnet) # install from github
+source("https://raw.githubusercontent.com/briatte/ggnet/master/ggnet2.R")
 data(blood, football, theme_elements, lesmis,madmen, mm_directed, email,
@@
 
+\subsubsection{Parameters and Aesthetics}
+\hh{Parameters currently implemented:
+\begin{itemize}
+\item layout: layout, layout.par, fiteach
+\item nodes/vertices: colour, size, alpha, shape
+\item edges: ecolour, ealpha, linewidth, stroke
+\item arrow: arrowsize, arrowgap (only if \code{directed = TRUE})
+\item labels: label (as parameter and variable), labelcolour
+\end{itemize}
+}
+
 \subsubsection{Vertex Aesthetics}
@@
 \par Evidently, we created all of these aesthetics to fit in exactly to the \texttt{ggplot2} grammar of graphics. We also created the edge aesthetics in a similar fashion.
+
   \subsubsection{Edge Aesthetics}
+
   Again, our edge aesthetics mimic the familiar \texttt{ggplot2} aesthetics, this time for segments.
@@
 \section{Examples of Networks}
-In this section, we demonstrate the current capabilities of \texttt{geom\_net} in a series of diverse examples.
+In this section, we demonstrate the current capabilities of \texttt{geom\_net} and \code{ggnet} in a series of side by side examples.
   \subsection{Blood Donation}
@@
 ggplot(data = bloodnet, aes(from_id = from, to_id = to)) +
-  geom_net(colour="red", layout="circle", label=TRUE, size=3,
-           directed=TRUE, vjust=-0.5, arrowsize=1.5) + theme_net()
+  geom_net(colour="darkred", layout="circle", label=TRUE, size=12,
+           directed=TRUE, vjust=0.5, labelcolour="grey80",
+           arrowsize=1.5, linewidth=0.75, arrowgap=0.04) + theme_net()
+@
+
+<<blood_ex_ggnet2, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
+ggnet2(network(blood$edges[, 1:2]), mode = "circle", size = 12,
+       label = TRUE, arrow.size = 7.5, arrow.gap=0.04, vjust = 0.5,
+       node.color="darkred", label.color="grey80")
 @
 This network is shown in figure \ref{fig.cap:blood_ex}.
-Here, we have used the aesthetics \texttt{vcolour} and \texttt{vsize} set to identity values to change the size and color of all vertices.
-We have also used the \texttt{layout} and \texttt{vlabel} arguments to change the default layout to a circle layout and to print the blood types, respectively.
+Here, we have used the aesthetics \texttt{colour} and \texttt{size} set to identity values to change the size and color of all vertices.
+We have also used the \texttt{layout} and \texttt{label} arguments to change the default layout to a circle layout and to print the blood types, respectively.
 The circle layout places blood types of the same ABO type next to each other and spreads the vertices out far enough to distinguish between the various ``in" and ``out" types.
-You can tell clearly from this plot that the O- type is the universal donor: it has out degree of seven and in degree of zero.
-Additionally, you can see that the AB+ type is the universal recepient, with in degree of seven and out degree of zero.
+You can tell clearly from this plot that the O- type is the universal donor: it has an out-degree of seven and an in-degree of zero.
+Additionally, we can see that the AB+ type is the universal recepient, with an in-degree of seven and an out-degree of zero.
 Anyone looking at this plot can quickly determine which type(s) of blood they can receive and which type(s) can receive their blood.
 \begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{figure/blood_ex-1.pdf}
+\begin{subfigure}[t]{0.49\textwidth}
+\caption{geom\_net}
+\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:blood_ex} Network of blood donation possibilities in humans by ABO and RhD blood types.}
@@
 The network contains two business weeks of emails across the entire company.
-In order to better visualize the structure of the communication network between employees, all emails that were sent out to all employees are removed in the subsequent examples.
-%
-\begin{figure}
-\centering
+In order to better visualize the structure of the communication network between employees, emails that were sent out to all employees are removed in the subsequent examples.
+
 <<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE, out.width='\\textwidth'>>=
@@
   geom_net(aes(colour= CurrentEmploymentType),
-           linewidth=0.5, alpha=0.1, size=4) +
-  scale_colour_brewer(palette="Set1") + theme_net() +
+           linewidth=0.5, ealpha=0.5, size=4, directed=FALSE) +
+  scale_colour_brewer("Employment Type", palette="Set1") + theme_net() +
   theme(legend.position="bottom")
 @
+
+<<email_ex_ggnet2, fig.cap = 'All emails within a company over a two week period.',echo=TRUE, out.width='\\textwidth'>>=
+em.cet = as.character(email$nodes$CurrentEmploymentType)
+names(em.cet) = email$nodes$label
+
+# full version
+em.net = subset(email$edges, nrecipients < 54)[, c("From", "to") ]
+em.net = network(em.net)
+em.net %v% "curr_empl_type" = em.cet[ network.vertex.names(em.net) ]
+
+ggnet2(em.net, size = 4, color = "curr_empl_type", palette = "Set1",
+       arrow.size = 0, edge.alpha = 0.5, color.legend = "Employment Type") +
+  theme(legend.position="bottom")
+@
+%
+\begin{figure}
+\begin{subfigure}[t]{0.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/email_ex_ggnet2-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
@@
 %\begin{landscape}
-\begin{figure}
-\centering
+
 <<email_ex_facet, fig.height=5, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
@@
   geom_net(aes(colour= CurrentEmploymentType), fiteach=TRUE,
-           linewidth=0.5, alpha=0.2, size=2) +
+           linewidth=0.5, ealpha=0.5, size=2) +
   scale_colour_brewer(palette="Set1") + theme_net() +
-  theme(legend.position="bottom") + facet_wrap(~day, nrow = 2)
+  facet_wrap(~day, nrow = 2, labeller="label_both") +
+  theme(legend.position="bottom",
+        panel.border = element_rect(fill=NA, colour = "grey60"))
 @
+
+<<email_ex_facet_ggnet2, fig.height=4, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
+em.day = subset(email$edges, nrecipients < 54)[, c("From", "to", "day") ]
+em.day = lapply(unique(em.day$day), function(x) subset(em.day, day == x)[, 1:2 ])
+em.day = lapply(em.day, network, directed = TRUE)
+for (i in 1:length(em.day)) {
+  em.day[[ i ]] %v% "curr_empl_type" = em.cet[ network.vertex.names(em.day[[ i ]]) ]
+  em.day[[ i ]] %n% "day" = unique(email$edges$day)[ i ]
+}
+
+g = list(length(em.day))
+for (i in 1:length(em.day)) {
+  g[[ i ]] = ggnet2(em.day[[ i ]], size = 2, color = "curr_empl_type",
+                    palette = "Set1", arrow.size = 0, edge.alpha = 0.5,
+                    legend.position = "none") +
+    ggtitle(paste("Day", em.day[[ i ]] %n% "day")) +
+    theme(panel.border = element_rect(color = "grey50", fill = NA))
+}
+gridExtra::grid.arrange(grobs = g, nrow = 2)
+
+@
+
+\begin{figure}
+\begin{subfigure}[t]{\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/email_ex_facet-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/email_ex_facet_ggnet2-1.pdf}
+\end{subfigure}
+
 \caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}

### 2015-10-10 R/geom-net.r

@@ -103,7 +103,7 @@
 #'   theme(legend.position="bottom")
 
 geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, inherit.aes = TRUE,  alpha = 0.25,
-                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey60", ealpha=NULL, arrowgap=0.01, directed = FALSE, arrowsize=1,
+                      layout="kamadakawai", layout.par=list(), fiteach=FALSE,  label=FALSE, ecolour="grey40", ealpha=NULL, arrowgap=0.01, directed = FALSE, arrowsize=1,
                       labelcolour=NULL, vertices=NULL, ...) {
     layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,

### 2015-10-10 R/geom-net.r
hjust added to defaults

@@ -122,7 +122,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
                              shape = 19, colour = "grey30",
                              size = 4, fill = NA, alpha = NA, stroke = 0.5,
-                             linewidth=1, angle=0, vjust=0),
+                             linewidth=1, angle=0, vjust=0, hjust=0.5),
 
   draw_key = function(data, params)  {
     with(data, grobTree(
@@ -199,6 +199,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
         angle = data$angle,
         alpha = NA,
         vjust= data$vjust,
+        hjust=data$hjust,
         stringsAsFactors = FALSE
       )
       labels <- unique(labels)

### 2015-10-11 R/geom-net.r
alpha for points

@@ -197,9 +197,9 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
         shape = data$shape,
         size = data$fontsize,
         angle = data$angle,
-        alpha = NA,
-        vjust= data$vjust,
-        hjust=data$hjust,
+        alpha = data$alpha,
+        vjust = data$vjust,
+        hjust = data$hjust,
         stringsAsFactors = FALSE
       )
       labels <- unique(labels)

### 2015-10-11 inst/CC/writeup.tex
finished side by side examples - problems with ggnet2 in football example

--- 2015-04-02 CC/writeup.Rnw
@@
-\documentclass[11pt]{article}
+\documentclass[11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{asa}
@@
 %%% BEGIN DOCUMENT
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \end{abstract}
-<<chunk_setup, echo=FALSE>>=
-library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
-@
-<<readingdata, eval=TRUE, echo=FALSE>>=
-library(ggplot2) # needs to be the version with geom_net!
- theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
- data(blood, football, theme_elements, lesmis,madmen, mm_directed,email)
-names(blood$vertices)[1] <- 'label'
-#load("data/madmen.rda") # relative path should work for both of us
-#load("data/mm_directed.rda")
-#names(madmen$vertices)[1] <- 'label'
-names(theme_elements$vertices)[1] <- 'label'
-football$vertices$label <- as.factor(football$vertices$label)
-football$vertices$value <- as.factor(football$vertices$value)
-football$edges$in.conf <- 0
-for (i in 1:nrow(football$edges)){
-  idx1 <- which(football$vertices$label == football$edges$from[i])
-  idx2 <- which(football$vertices$label == football$edges$to[i])
-  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
-    football$edges$in.conf[i] <- 1
-  }
-}
-#load("data/email.rda")
-#email.net <- read.csv("data/email-data.csv")
-#email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))
-#library(plyr)
-#email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
-#employee <- read.csv("data/EmployeeRecords.csv")
-#employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
-email.net <- email$edges
-employee <- email$nodes
-library(RColorBrewer)
-cols <- brewer.pal(8, name="Set1")
-yeast <- read.table("data/protein.txt")
-yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
-@
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'email' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'email' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'protein' not found}}\end{kframe}
+\end{knitrout}
 
@@
 This network is shown in figure~\ref{fig.cap:madmen_ex}.
-<<madmen_ex, dependson='madmen-prep', fig.cap = 'Network of Relationships in Mad Men'>>=
-ggplot(data = madmen$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = madmen$vertices, vsize=3, vlabel= TRUE, ecolour="grey30",
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.25), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): could not find function "{}geom\_net"{}}}\end{kframe}
+\end{knitrout}
 
@@
 In the example below, loops are removed because loops exist on every vertex in this example, as blood between two people of matching ABO and RhD type can always be exchanged.
-<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.">>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(vertices = blood$vertices, vcolour = I('red'), layout = 'circle',
-           vlabel = TRUE, vsize = I(3), directed = TRUE) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= blood}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= from,} \hlkwc{to_id} \hlstd{= to))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= blood}\hlopt{$}\hlstd{vertices,} \hlkwc{vcolour} \hlstd{=} \hlkwd{I}\hlstd{(}\hlstr{'red'}\hlstd{),} \hlkwc{layout} \hlstd{=} \hlstr{'circle'}\hlstd{,}
+           \hlkwc{vlabel} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{vsize} \hlstd{=} \hlkwd{I}\hlstd{(}\hlnum{3}\hlstd{),} \hlkwc{directed} \hlstd{=} \hlnum{TRUE}\hlstd{)} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): could not find function "{}geom\_net"{}}}\end{kframe}
+\end{knitrout}
 This network is shown in figure \ref{fig.cap:blood_ex}.
@@
 In order to better visualize the structure of the communication network between employees, all emails that were sent out to all employees are removed in the subsequent examples.
-<<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee,ealpha = 0.1, vsize = I(4), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{=} \hlkwd{subset}\hlstd{(email.net, nrecipients} \hlopt{<} \hlnum{54}\hlstd{),} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= From,} \hlkwc{to_id} \hlstd{= to))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= employee,}\hlkwc{ealpha} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{vsize} \hlstd{=} \hlkwd{I}\hlstd{(}\hlnum{4}\hlstd{),} \hlkwc{directed} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+           \hlkwd{aes}\hlstd{(}\hlkwc{vcolour}\hlstd{=cols[}\hlkwd{as.numeric}\hlstd{(CurrentEmploymentType)]))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset(email.net, nrecipients < 54): object 'email.net' not found}}\end{kframe}
+\end{knitrout}
 \begin{figure}
@@
 To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
-<<email_ex_facet, fig.height=8, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE,eval=FALSE>>=
-ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
-  geom_net(vertices = employee, ealpha = 0.2, vsize = I(2), directed = TRUE,
-           aes(vcolour=cols[as.numeric(CurrentEmploymentType)])) +
-  expand_limits(x=c(0,1), y = c(0,1)) + facet_wrap(~day, nrow = 2) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{=} \hlkwd{subset}\hlstd{(email.net, nrecipients} \hlopt{<} \hlnum{54}\hlstd{),} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= From,} \hlkwc{to_id} \hlstd{= to))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= employee,} \hlkwc{ealpha} \hlstd{=} \hlnum{0.2}\hlstd{,} \hlkwc{vsize} \hlstd{=} \hlkwd{I}\hlstd{(}\hlnum{2}\hlstd{),} \hlkwc{directed} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+           \hlkwd{aes}\hlstd{(}\hlkwc{vcolour}\hlstd{=cols[}\hlkwd{as.numeric}\hlstd{(CurrentEmploymentType)]))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{day,} \hlkwc{nrow} \hlstd{=} \hlnum{2}\hlstd{)} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 \begin{landscape}
@@
 Finally, we note that the vertices with no edges were incorporated into the plot by adding their labels to the edges data frame in both the  `from\_id' and `to\_id' columns before passing the edges data frame to \texttt{ggplot}.
-<<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-ggplot(data = theme_elements$edges, aes(from_id = parent, to_id = child)) +
-  geom_net(vertices = theme_elements$vertices, directed = TRUE, vlabel = TRUE) +
-  expand_limits(x = c(0,1.2), y = c(0,1)) +
-  theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= theme_elements}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= parent,} \hlkwc{to_id} \hlstd{= child))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= theme_elements}\hlopt{$}\hlstd{vertices,} \hlkwc{directed} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{vlabel} \hlstd{=} \hlnum{TRUE}\hlstd{)} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1.2}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+}
+  \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): could not find function "{}geom\_net"{}}}\end{kframe}
+\end{knitrout}
 \begin{figure}
@@
 We changed the vertex size and edge color for all vertices and edges, included vertex labels, and colored the vertices according to the character's gender.
-<<print_ex, ref.label="madmen_ex", eval=FALSE, echo=TRUE,cache=FALSE>>=
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= madmen}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= Name1,} \hlkwc{to_id} \hlstd{= Name2))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= madmen}\hlopt{$}\hlstd{vertices,} \hlkwc{vsize}\hlstd{=}\hlnum{3}\hlstd{,} \hlkwc{vlabel}\hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{ecolour}\hlstd{=}\hlstr{"grey30"}\hlstd{,}
+           \hlkwd{aes}\hlstd{(}\hlkwc{vcolour}\hlstd{=}\hlkwd{c}\hlstd{(} \hlstr{"#FF69B4"}\hlstd{,} \hlstr{"#0099ff"}\hlstd{)[}\hlkwd{as.numeric}\hlstd{(Gender)]))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1.25}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
  There is another Mad Men network included in the \texttt{gcookbook} package \citep{madmen}.
@@
 If the advance is reciprocated, like between Sal Romano and the Bellhop, there are two directed edges between the two vertices.
-<<madmen2_ex,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
-ggplot(data = mm.directed$edges, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(vertices = mm.directed$vertices, directed = T, vlabel = T,
-           vsize = I(2.5), layout = 'fruchtermanreingold',
-           aes(vcolour=c( "#FF69B4", "#0099ff")[as.numeric(Gender)])) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= mm.directed}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= Name1,} \hlkwc{to_id} \hlstd{= Name2))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= mm.directed}\hlopt{$}\hlstd{vertices,} \hlkwc{directed} \hlstd{= T,} \hlkwc{vlabel} \hlstd{= T,}
+           \hlkwc{vsize} \hlstd{=} \hlkwd{I}\hlstd{(}\hlnum{2.5}\hlstd{),} \hlkwc{layout} \hlstd{=} \hlstr{'fruchtermanreingold'}\hlstd{,}
+           \hlkwd{aes}\hlstd{(}\hlkwc{vcolour}\hlstd{=}\hlkwd{c}\hlstd{(} \hlstr{"#FF69B4"}\hlstd{,} \hlstr{"#0099ff"}\hlstd{)[}\hlkwd{as.numeric}\hlstd{(Gender)]))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1.1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(data = mm.directed\$edges, aes(from\_id = Name1, to\_id = Name2)): object 'mm.directed' not found}}\end{kframe}
+\end{knitrout}
 \begin{figure}
@@
 There is an additional variable in the vertex data frame corresponding to the conference each team belongs to, and there is an additional variable in the edge data frame that is equal to one if the game occured between teams in the same conference or zero if the game occured between teams in different conferences.
-<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-colors <- brewer.pal(12, name = 'Paired')
-ggplot(data = football$edges, aes(from_id = from, to_id = to,
-                                  elinetype = in.conf + 1)) +
-  geom_net(vertices = football$vertices, ealpha = 0.3, vsize = 2, vlabel = TRUE,
-           aes(vcolour=colors[as.numeric(factor(value))])) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{colors} \hlkwb{<-} \hlkwd{brewer.pal}\hlstd{(}\hlnum{12}\hlstd{,} \hlkwc{name} \hlstd{=} \hlstr{'Paired'}\hlstd{)}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= football}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= from,} \hlkwc{to_id} \hlstd{= to,}
+                                  \hlkwc{elinetype} \hlstd{= in.conf} \hlopt{+} \hlnum{1}\hlstd{))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= football}\hlopt{$}\hlstd{vertices,} \hlkwc{ealpha} \hlstd{=} \hlnum{0.3}\hlstd{,} \hlkwc{vsize} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{vlabel} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+           \hlkwd{aes}\hlstd{(}\hlkwc{vcolour}\hlstd{=colors[}\hlkwd{as.numeric}\hlstd{(}\hlkwd{factor}\hlstd{(value))]))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): could not find function "{}geom\_net"{}}}\end{kframe}
+\end{knitrout}
 The network of football games is given in figure \ref{fig.cap:football_ex}.
@@
 There is an extremely wide edge connecting Valjean and Cosette, which implies that they are strongly connected; in fact, Cosette is Valjean's adopted daughter.
-<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-ggplot(data = lesmis$edges, aes(from_id = from, to_id = to,
-                                esize = value/mean(value))) +
-  geom_net(vertices = lesmis$vertices, ealpha = 7/10,
-           vcolour = I("red"), vsize = I(2.5),
-           layout = 'fruchtermanreingold', vlabel=TRUE) +
-  expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= lesmis}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= from,} \hlkwc{to_id} \hlstd{= to,}
+                                \hlkwc{esize} \hlstd{= value}\hlopt{/}\hlkwd{mean}\hlstd{(value)))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= lesmis}\hlopt{$}\hlstd{vertices,} \hlkwc{ealpha} \hlstd{=} \hlnum{7}\hlopt{/}\hlnum{10}\hlstd{,}
+           \hlkwc{vcolour} \hlstd{=} \hlkwd{I}\hlstd{(}\hlstr{"red"}\hlstd{),} \hlkwc{vsize} \hlstd{=} \hlkwd{I}\hlstd{(}\hlnum{2.5}\hlstd{),}
+           \hlkwc{layout} \hlstd{=} \hlstr{'fruchtermanreingold'}\hlstd{,} \hlkwc{vlabel}\hlstd{=}\hlnum{TRUE}\hlstd{)} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1.1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): could not find function "{}geom\_net"{}}}\end{kframe}
+\end{knitrout}
 \begin{figure}
@@
 
-<<yeast_code>>=
-sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$V1,sample.edges$V2))))
-@
-<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-ggplot(data = yeast.samp$edges, aes(from_id = V1, to_id = V2)) +
-  geom_net(vertices = yeast.samp$vertices, ealpha = .25, valpha = .5, 
-           vcolour = 'magenta', layout = 'random', 
-           layout.par = list(dist = 'uniang')) +
-  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'yeast' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'sample.edges' not found}}\end{kframe}
+\end{knitrout}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{ggplot}\hlstd{(}\hlkwc{data} \hlstd{= yeast.samp}\hlopt{$}\hlstd{edges,} \hlkwd{aes}\hlstd{(}\hlkwc{from_id} \hlstd{= V1,} \hlkwc{to_id} \hlstd{= V2))} \hlopt{+}
+  \hlkwd{geom_net}\hlstd{(}\hlkwc{vertices} \hlstd{= yeast.samp}\hlopt{$}\hlstd{vertices,} \hlkwc{ealpha} \hlstd{=} \hlnum{.25}\hlstd{,} \hlkwc{valpha} \hlstd{=} \hlnum{.5}\hlstd{,}
+           \hlkwc{vcolour} \hlstd{=} \hlstr{'magenta'}\hlstd{,} \hlkwc{layout} \hlstd{=} \hlstr{'random'}\hlstd{,}
+           \hlkwc{layout.par} \hlstd{=} \hlkwd{list}\hlstd{(}\hlkwc{dist} \hlstd{=} \hlstr{'uniang'}\hlstd{))} \hlopt{+}
+  \hlkwd{expand_limits}\hlstd{(}\hlkwc{x} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{),} \hlkwc{y} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{))} \hlopt{+} \hlstd{theme_net}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(data = yeast.samp\$edges, aes(from\_id = V1, to\_id = V2)): object 'yeast.samp' not found}}\end{kframe}
+\end{knitrout}
 
@@
    \subsection{\texttt{stat\_net} Usage}
-<<usage_stat_net, eval=FALSE>>=
-stat_net(mapping = NULL, data = NULL, vertices = NULL, geom = "net", position = "identity", ...)
-@
+
   The usage of the \texttt{stat\_net} arguments is as follows:
@@
    \subsection{\texttt{geom\_net} Usage}
-<<usage_geom_net, eval=FALSE>>=
-geom_net(mapping = NULL, data = NULL, vertices = NULL, vlabel = FALSE, directed = FALSE, stat = "net", ...)
-@
+
   The usage of the \texttt{geom\_net} arguments is as follows:

### 2015-10-11 inst/paper.Rnw
finished side by side examples - problems with ggnet2 in football example

@@
 TEnet <- merge(theme_elements$edges, theme_elements$vertices, by.x="parent", by.y="name", all=TRUE)
+library(dplyr)
+TEnet <- TEnet %>% group_by(parent) %>% mutate(degree=sqrt(10*n()+1))
 
-football$edges$in.conf <- 1
-for (i in 1:nrow(football$edges)){
-  idx1 <- which(football$vertices$label == football$edges$from[i])
-  idx2 <- which(football$vertices$label == football$edges$to[i])
-  if (football$vertices$value[idx1] == football$vertices$value[idx2]){
-    football$edges$in.conf[i] <- 0
-  }
-}
 ftnet <- merge(football$edges, football$vertices, by.x="from", by.y="label", all=TRUE)
 
+lesmisnet <- merge(lesmis$edges, lesmis$vertices, by.x="from", by.y="label", all=TRUE)
+
 tripnet <- merge(bikes$trips, bikes$stations, by.x="Start.station", by.y="name", all=TRUE)
+
+yeast <- protein
+
 @
@@
 ggplot(data = bloodnet, aes(from_id = from, to_id = to)) +
-  geom_net(colour="darkred", layout="circle", label=TRUE, size=12,
+  geom_net(colour="darkred", layout="circle", label=TRUE, size=15,
            directed=TRUE, vjust=0.5, labelcolour="grey80",
-           arrowsize=1.5, linewidth=0.75, arrowgap=0.04) + theme_net()
+           arrowsize=1.5, linewidth=0.75, arrowgap=0.05) + theme_net()
 @
@@
 <<blood_ex_ggnet2, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
-ggnet2(network(blood$edges[, 1:2]), mode = "circle", size = 12,
-       label = TRUE, arrow.size = 7.5, arrow.gap=0.04, vjust = 0.5,
+ggnet2(network(blood$edges[, 1:2]), mode = "circle", size = 15,
+       label = TRUE, arrow.size = 7.5, arrow.gap=0.05, vjust = 0.5,
        node.color="darkred", label.color="grey80")
@@
   geom_net(aes(colour= CurrentEmploymentType),
-           linewidth=0.5, ealpha=0.5, size=4, directed=FALSE) +
+           linewidth=0.5, ealpha=0.5, size=4, directed=TRUE) +
   scale_colour_brewer("Employment Type", palette="Set1") + theme_net() +
@@
 ggnet2(em.net, size = 4, color = "curr_empl_type", palette = "Set1",
-       arrow.size = 0, edge.alpha = 0.5, color.legend = "Employment Type") +
+       arrow.size = 1.5, edge.alpha = 0.5, color.legend = "Employment Type") +
   theme(legend.position="bottom")
@@
 There are six distinct clusters in this network which almost perfectly correspond to the six different types of employee in this company: administration, engineering, executive, facilities, information technology, and security.
-Unfortunately, there is currently no legend to associate color with employment type.
+%Unfortunately, there is currently no legend to associate color with employment type.
 This will be remedied in the final version of our geometry.
-Additionally, the edges between employees in the same cluster are much darker than edges between employees in different clusters.
+Additionally, the edges between employees in the same cluster are  darker than edges between employees in different clusters.
 This is due to the value of the \texttt{ealpha} aesthetic: more emails between two employees lead to darker edges.
-The value is set to 0.1 in this example, so that 10 or more emails between 2 employees results in a completely opaque edge.
+The value is set to 0.1 in this example, so that ten or more emails between two employees result in a completely opaque edge.
 This pattern of heavy communication between employees of the same type is fairly unsurprising.
-To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig.cap:email_ex_facet} shows the different email networks associated with each day of the week.
+To make this visualization more interesting and informative, we facet the network by day: each panel in \ref{fig:email_ex_facet} shows the different email networks associated with each day of the week.
 %\begin{landscape}
@@
   geom_net(aes(colour= CurrentEmploymentType), fiteach=TRUE,
-           linewidth=0.5, ealpha=0.5, size=2) +
+           linewidth=0.5, ealpha=0.1, size=2) +
   scale_colour_brewer(palette="Set1") + theme_net() +
@@
   g[[ i ]] = ggnet2(em.day[[ i ]], size = 2, color = "curr_empl_type",
-                    palette = "Set1", arrow.size = 0, edge.alpha = 0.5,
-                    legend.position = "none") +
+                    palette = "Set1", arrow.size = 0, arrow.gap=0.01,
+                    edge.alpha = 0.1, legend.position = "none") +
     ggtitle(paste("Day", em.day[[ i ]] %n% "day")) +
@@
 
-\caption{\label{fig.cap:email_ex_facet} The same email network as in figure \ref{fig.cap:email_ex} facetted by day of the week.}
+\caption{\label{fig:email_ex_facet} The same email network as in figure~\ref{fig.cap:email_ex} facetted by day of the week.}
 \end{figure}
 %\end{landscape}
-This plot is shown in figure \ref{fig.cap: email_ex_facet}.
+This plot is shown in Figure~\ref{fig:email_ex_facet}.
 With the facetting, we can see that there are several days where one or more departments do not communicate with any of the other departments.
@@
 There is an arrow from one theme option to another if the element represented by the \emph{to} vertex inherits its values from the \emph{from} vertex.
-For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which inherits its value from the \texttt{line} option.
+For example, the \texttt{axis.ticks.x} option inherits its value from the \texttt{axis.ticks} value, which in turn inherits its value from the \texttt{line} option.
 Thus, setting the \texttt{line} option to a value such as \texttt{element\_blank()} sets the entire inheritance tree to \texttt{element\_blank()}, and no lines appear anywhere on the plot background.
@@
 <<theme_ex, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
-library(dplyr)
-TEnet <- TEnet %>% group_by(parent) %>% mutate(degree=sqrt(10*n()+1))
-
 ggplot(data = TEnet, aes(from_id = parent, to_id = child)) +
   geom_net(aes(fontsize=degree), directed = TRUE, label = TRUE, vjust=-.5,
-           size=3) +
+           size=3, ecolour="grey70") +
   theme_net() + xlim(c(-0.05, 1.05))
 @
+<<theme_ex_ggnet2, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
+te.net = network(theme_elements$edges)
+te.net %v% "size" = sqrt(10 * (sna::degree(te.net) + 1))
+ggnet2(te.net, label = TRUE, color = "white",
+       label.size = "size", layout.exp = 0.15)
+@
+
 \begin{figure}
-\centering
+\begin{subfigure}[t]{.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/theme_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/theme_ex_ggnet2-1.pdf}
+\end{subfigure}
+
 \caption{\label{fig.cap:theme_ex} Inheritance structure of \texttt{ggplot2} theme elements. This is a recreation of the graph found at \protect\url{http://docs.ggplot2.org/current/theme.html}.}
@@
   geom_net(aes(colour=Gender), directed=TRUE, label=TRUE,
-           size=2.5, layout='fruchtermanreingold') +
+           ecolour="grey50", linewidth=0.5, size=2.5, vjust=-.5,
+           layout='fruchtermanreingold') +
   scale_colour_manual(values=c( "#FF69B4", "#0099ff")) +
@@
 @
+
+<<madmen2_ex_ggnet2,echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
+rownames(mm.directed$vertices) = mm.directed$vertices$label
+mm.directed$vertices$Gender = as.character(mm.directed$vertices$Gender)
+
+mm.dir = network(mm.directed$edges, directed = TRUE)
+mm.dir %v% "gender" = mm.directed$vertices[ network.vertex.names(mm.dir), "Gender" ]
+
+# gender color palette
+mm.col = c("f" = "#ff69b4", "m" = "#0099ff")
+
+ggnet2(mm.dir, mode = "fruchtermanreingold", size = 3,
+       color = mm.col[ mm.dir %v% "gender" ],
+       label = TRUE, label.color = mm.col[ mm.dir %v% "gender" ],
+       hjust = -0.1, legend.position = "bottom", layout.exp = 0.15,
+       arrow.size = 10)
+@
 \begin{figure}
-\centering
+\begin{subfigure}[t]{.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/madmen2_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/madmen2_ex_ggnet2-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
@@
 This network is shown in figure \ref{fig.cap:madmen2_ex}.
-This network is much more connected than the previous Mad Men example.
-It also allows us to see much more of the drama in the show.
-For instance, you can see that Roger made advances towards Betty, his business partner's wife, which is much more scandalous than what we can see in the first network.
+It is a lot more densely connected than the previous Mad Men example.
+This network allows us to see much more of the drama in the show.
+For instance, we can see that Roger made advances towards Betty, his business partner's wife, which is much more scandalous than what appears in the first network.
 
@@
 <<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
-sid <- which(ftnet$from %in% c("Navy", "NotreDame", "UtahState", "CentralFlorida", "Connecticut"))
+sid <- which(ftnet$from %in% c("Navy", "NotreDame", "UtahState",
+                               "CentralFlorida", "Connecticut"))
 ftnet$schools <- ""
@@
 ggplot(data=ftnet, aes(from_id=from, to_id=to)) +
-  geom_net(aes(colour=value, linewidth=(2-in.conf)/3,  label=schools),
-           size=4, ecolour="grey30", vjust=-0.5,
-           alpha=0.3, layout='fruchtermanreingold') +
-  scale_colour_brewer("Conference", palette="Paired") + theme_net() +
-  theme(legend.position="bottom") + xlim(c(-0.05, 1.05))
+  geom_net(aes(colour=value, linewidth=0.5,
+               linetype=factor(same.conf==1), label=schools),
+           size=4,  vjust=-0.5,
+           alpha=0.3, layout='fruchtermanreingold') + theme_net() +
+  theme(legend.position="bottom") + xlim(c(-0.05, 1.05)) +
+  scale_colour_brewer("Conference", palette="Paired")  +
+  scale_linetype_manual("Same Conference", values=c(2,1))
 @
+
+<<football_ex_ggnet2, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
+rownames(football$vertices) = football$vertices$label
+
+fb.net = network(football$edges, directed = FALSE)
+fb.net %v% "conf" = football$vertices[ network.vertex.names(fb.net), "value" ]
+
+ggnet2(fb.net, mode = "fruchtermanreingold",
+       color = "conf", color.legend = "Conference", palette = "Paired")
+#       edge.color = c("color", "grey75"))
+@
+
 The network of football games is given in figure \ref{fig.cap:football_ex}.
-Here, we have changed the \texttt{elinetype} aesthetic to correspond to games that occur between teams in the same conference or different conferences.
+Here, we have changed the \texttt{linetype} aesthetic to correspond to games that occur between teams in the same conference or different conferences.
 These lines are dotted and solid, respectively.
@@
 This coloring and changing of the linetypes make the structure of the game network easier to view.
-There is one conference consisting of Navy, Notre Dame, Utah State, Central Florida, and Connecticut, which is spread out all over the network, whereas most other conferences' teams are all very close to each other because they play within conference much more than they play out of conference.
+There is one conference consisting of Navy, Notre Dame, Utah State, Central Florida, and Connecticut, which is spread out, whereas most other conferences' teams are all very close to each other because they play within conference much more than they play out of conference.
 At the time, these five schools were all independents and did not have a home conference.
@@
 \begin{figure}
-\centering
+\begin{subfigure}[t]{.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/football_ex_ggnet2-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
@@
 It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.
-There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.
+There are 77 vertices representing each of the 77 characters in the book.
+An edge connects two vertices if those two characters appear in the same chapter of the book.
 There are 254 edges in this network.
@@
 This network is shown in figure \ref{fig.cap:lesmis_ex}.
-This network reveals many important character groups, as well as the main characters in the story.
-The main characters are the ones with high degree, like Valjean, Cosette, and Javert.
-There is an extremely wide edge connecting Valjean and Cosette, which implies that they are strongly connected; in fact, Cosette is Valjean's adopted daughter.
+This
+  ted daughter.
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-lesmisnet <- merge(lesmis$edges, lesmis$vertices, by.x="from", by.y="label", all=TRUE)
 
-ggplot(data = lesmisnet, aes(from_id = from, to_id = to, linewidth = value/5)) +
-  geom_net(alpha = 7/10, colour = "red", size = 2.5,
+ggplot(data = lesmisnet, aes(from_id = from, to_id = to, linewidth = degree/5 + 0.1 )) +
+  geom_net(colour = "grey30",  aes(size=degree),
+           ecolour="grey60", vjust=-0.5,
            layout = 'fruchtermanreingold', label=TRUE) +
-  theme_net()
+  theme_net() +
+  theme(legend.position="bottom")
 @
+
+<<lesmis_ex_ggnet2, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
+ggnet2(lesmis$edges[, 1:2 ], mode = "kamadakawai", color = "white", label = TRUE,
+       edge.size = lesmis$edges$degree / mean(lesmis$edges$degree),
+       layout.exp = 0.25)
+@
+
 \begin{figure}
-\centering
+\begin{subfigure}[t]{.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/lesmis_ex_ggnet2-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:lesmis_ex} Co-appearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
@@
   theme_net() + xlim(c(-0.1, 1.1)) +
-  scale_colour_manual(values=c("grey40", "darkorange"))
+  scale_colour_manual("Metro Station", values=c("grey40", "darkorange")) +
+  theme(legend.position="bottom")
 # compare to https://secure.capitalbikeshare.com/map/
@@
   theme_net() +
-    scale_colour_manual(values=c("grey40", "darkorange"))
+  scale_colour_manual("Metro Station", values=c("grey40", "darkorange")) +
+  theme(legend.position="bottom")
 
@@
 There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.
-These edges represent ``direct physical interactions" between any 2 proteins \citep[][p.
+These edges represent ``direct physical interactions" between any two proteins \citep[][p.
 42]{protein}.
-Because the network is very large, we randomly sampled 500 interactions and their associated proteins to plot it in figure \ref{fig.cap:yeast_ex}.
+These interactions and their associated proteins are plotted  in figure \ref{fig.cap:yeast_ex}.
 We also demonstrate the layout capabilities by changing the layout to random and setting the distribution to \texttt{"uniang"}, which is a  ``gaussian donut" layout.
@@
 
-<<yeast_code>>=
-yeast <- protein
-sample.edges <- yeast$edges[sample(nrow(yeast$edges), 500),]
-yeast.samp <- list(edges = sample.edges, vertices = data.frame(label = unique(c(sample.edges$from,sample.edges$to))))
-
-proteinnet <- merge(yeast.samp$vertices, yeast.samp$edges, by.x="label", by.y="from", all=TRUE)
-@
 <<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
-ggplot(data = proteinnet, aes(from_id = label, to_id = to)) +
-  geom_net(alpha=0.25, ealpha = .1, colour = 'magenta', layout = 'random',
-           layout.par = list(dist = 'uniang')) +
+ggplot(data = protein$edges, aes(from_id = from, to_id = to)) +
+  geom_net(alpha=0.25, ealpha = .05, size=2, colour = 'magenta',
+           ecolour="grey70", linewidth=0.5,
+           layout = 'random', layout.par = list(dist = 'uniang')) +
   theme_net()
@@
 
+<<yeast_ex_ggnet2, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=FALSE>>=
+ggnet2(network(yeast$edges[,1:2]), size = 2, color = "magenta",
+       mode = "random", layout.par = list(dist = "uniang"),
+       edge.alpha = 0.05)
+@
+
 \begin{figure}
-\centering
+\begin{subfigure}[t]{.49\textwidth}
+\caption{geom\_net}
 \includegraphics[width=\textwidth]{figure/yeast_ex-1.pdf}
-\caption{\label{fig.cap:yeast_ex} A subset of the complete protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/yeast_ex_ggnet2-1.pdf}
+\end{subfigure}
+
+\caption{\label{fig.cap:yeast_ex} Protein-protein interaction network in \emph{S. cerevisiae}. The layout is random.}
 \end{figure}

### 2015-10-11 inst/paper.Rnw
moving data statements into the code for individual examples

@@
 source("https://raw.githubusercontent.com/briatte/ggnet/master/ggnet2.R")
-data(blood, football, theme_elements, lesmis,madmen, mm_directed, email,
+data(football, theme_elements, lesmis,madmen, mm_directed, email,
      bikes)
@@
 levels(MM2net$Gender) <- c("female", "male")
-
-bloodnet <- merge(blood$edges, blood$vertices, by.x="from", by.y="label", all=TRUE)
-
-employee <- data.frame(expand.grid(
-              label=unique(email$nodes$label), day=unique(email$edges$day)))
-employee <- merge(employee, email$nodes, by="label")
-emailnet <- merge(subset(email$edges, nrecipients < 54), employee, by.x=c("From", "day"), by.y=c("label", "day"), all=TRUE)
 
@@
 In the example below, loops are removed because loops exist on every vertex in this example, as blood between two people of matching ABO and RhD type can always be exchanged.
+
 <<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
-ggplot(data = bloodnet, aes(from_id = from, to_id = to)) +
+library(ggnet)
+data(blood)
+
+# source code for geom_net
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
   geom_net(colour="darkred", layout="circle", label=TRUE, size=15,
@@
 <<blood_ex_ggnet2, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
+# source code for ggnet2
+library(GGally) # use developmental version on github for now
+
 ggnet2(network(blood$edges[, 1:2]), mode = "circle", size = 15,
@@
 <<email_ex, fig.cap = 'All emails within a company over a two week period.',echo=TRUE, out.width='\\textwidth'>>=
+emailnet <- merge(subset(email$edges, nrecipients < 54), email$nodes, by.x="From", by.y="label", all=TRUE)
+
 ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
@@
 <<email_ex_facet, fig.height=5, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
+employee <- data.frame(expand.grid(label=unique(email$nodes$label),
+                                   day=unique(email$edges$day)))
+employee <- merge(employee, email$nodes, by="label")
+emailnet <- merge(subset(email$edges, nrecipients < 54), employee, by.x=c("From", "day"), by.y=c("label", "day"), all=TRUE)
+
 ggplot(data = emailnet, aes(from_id = From, to_id = to)) +
   geom_net(aes(colour= CurrentEmploymentType), fiteach=TRUE,
-           linewidth=0.5, ealpha=0.1, size=2) +
+           linewidth=0.5, ealpha=0.5, size=2, directed=TRUE,
+           arrowsize=.5) +
   scale_colour_brewer(palette="Set1") + theme_net() +

### 2015-10-11 inst/paper.Rnw
if cruvature is specified, geom_curve rather than geom_segment is used for edges

@@
   geom_net(colour="darkred", layout="circle", label=TRUE, size=15,
-           directed=TRUE, vjust=0.5, labelcolour="grey80",
+           directed=TRUE, vjust=0.5, labelcolour="grey80", curvature=0,
            arrowsize=1.5, linewidth=0.75, arrowgap=0.05) + theme_net()

### 2015-10-11 R/geom-net.r

@@ -122,7 +122,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   default_aes = ggplot2::aes(width = 0.75, linetype = "solid", fontsize=5,
                              shape = 19, colour = "grey30",
                              size = 4, fill = NA, alpha = NA, stroke = 0.5,
-                             linewidth=1, angle=0, vjust=0, hjust=0.5),
+                             linewidth=1, angle=0, vjust=0, hjust=0.5, curvature = 0),
 
   draw_key = function(data, params)  {
     with(data, grobTree(
@@ -173,6 +173,11 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
                          yend = y + (1-arrowgap)*(yend-y))
 
     }
+    if (any(data$curvature != 0))
+      edges_draw <- GeomCurve$draw_panel(edges, panel_scales,
+                      coord, arrow=arrow, curvature=data$curvature[1], angle=90)
+    else edges_draw <- GeomSegment$draw_panel(edges, panel_scales, coord, arrow)
+
 
     vertices <- data.frame(
       x = data$x,
@@ -207,7 +212,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
     }
 
     ggplot2:::ggname("geom_net", grobTree(
-      GeomSegment$draw_panel(edges, panel_scales, coord, arrow),
+      edges_draw,
       GeomPoint$draw_panel(vertices, panel_scales, coord),
       label_grob
     ))

### 2015-10-12 inst/paper.Rnw
update on ggplot2 statistics ... they are pretty amazing!

@@
 We found the current tools to be lacking in this ability, so we chose to fill this gap by adding network plotting capabilities to the popular and widely used \texttt{R} package \texttt{ggplot2}.
-Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}, from January 1, 2015 to March 21, 2015, \texttt{ggplot2} was downloaded over 270,000 times, or approximately 3,454 downloads per day.
-It has also been downloaded in 189 countries at least once, and in 31 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times\footnote{\texttt{ggplot2} usage statistics taken from \url{http://cran-logs.rstudio.com/}.}.
+Just to give an idea of the popularity and the wide-spread use of \texttt{ggplot2}, from January 1, 2015 to October 4, 2015, \texttt{ggplot2} was downloaded over 1.2 Million times, or approximately 4,342 downloads per day.
+It has also been downloaded in 215 countries at least once, and in 61 of those countries, including China, Israel, and Colombia, it has been downloaded over 1,000 times\footnote{\texttt{ggplot2} usage statistics taken from \url{http://cran-logs.rstudio.com/}.}.
 This is the user base we are aiming at by making network visualizations a part of \texttt{ggplot2}.

### 2015-10-12 inst/paper.Rnw
Francois' changes

@@
 \end{abstract}
-<<chunk_setup, echo=FALSE>>=
+<<setup_knitr, echo=FALSE>>=
 library(knitr)
-opts_chunk$set(tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
+opts_chunk$set(
+  tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE,
+  message = FALSE, warning = FALSE,
+  fig.height = 8, fig.width = 8, fig.align = 'center', fig.show='hide'
+)
 @
 
-<<readingdata, eval=TRUE, echo=FALSE>>=
+<<load_packages, eval=TRUE, echo=FALSE>>=
+## ggplot2
 library(ggplot2) # needs to be version ≥ 1.0.1.9003
-library(ggnet) # install from github
+
+## ggnet
+if (!require(ggnet, quietly = TRUE)) {
+  devtools::install_github("sctyner/ggnet")
+}
+library(ggnet)
+
+## ggnetwork
+if (!require(ggnetwork, quietly = TRUE)) {
+  devtools::install_github("briatte/ggnetwork")
+}
+library(ggnetwork)
+
+## ggnet2
 source("https://raw.githubusercontent.com/briatte/ggnet/master/ggnet2.R")
-data(football, theme_elements, lesmis,madmen, mm_directed, email,
-     bikes)
+@
 
+<<read_data, eval=TRUE, echo=FALSE>>=
+data(football, theme_elements, lesmis, madmen, mm_directed, email, bikes)
 data(protein)
+
 library(RColorBrewer)
@@
 # set up networks for all of the data sets
+
 MMnet <- merge(madmen$edges, madmen$vertices, by.x="Name1", by.y="label", all=TRUE)
@@
 
+library(network)
+
+## Mad Men -> network class, undirected
+rownames(madmen$vertices) = madmen$vertices$label
+mm.net = network(madmen$edges[, 1:2], directed = FALSE)
+mm.net %v% "Gender" = as.character(madmen$vertices[ network.vertex.names(mm.net), "Gender"])
+mm.net %v% "Gender" = ifelse(mm.net %v% "Gender" == "f", "female", "male")
+
+## Mad Men -> network class, directed
+mm.dir = network(madmen$edges[, 1:2], directed = TRUE)
+mm.dir %v% "Gender" = as.character(madmen$vertices[ network.vertex.names(mm.dir), "Gender"])
+mm.dir %v% "Gender" = ifelse(mm.dir %v% "Gender" == "f", "female", "male")
+
+library(dplyr)
 TEnet <- merge(theme_elements$edges, theme_elements$vertices, by.x="parent", by.y="name", all=TRUE)
-library(dplyr)
 TEnet <- TEnet %>% group_by(parent) %>% mutate(degree=sqrt(10*n()+1))
@@
 yeast <- protein
-
 @
@@
 \centering
-<<madmen_ex, dependson='readingdata', fig.cap = 'Network of Relationships in Mad Men', out.width='\\textwidth', fig.width=9, fig.height=9>>=
+<<madmen_ex_ggnet, dependson='reading_data', fig.cap = 'Network of Relationships in Mad Men', out.width='\\textwidth', fig.width=9, fig.height=9>>=
 ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2)) +
@@
            ecolour="grey50") +
+  scale_colour_manual(values=c( "#FF69B4", "#0099ff")) +
+  xlim(c(-0.05, 1.05)) +
   theme_net() +
-  scale_colour_manual(values=c( "#FF69B4", "#0099ff")) +
-  xlim(c(-0.05, 1.05)) + theme(legend.position="bottom")
+  theme(legend.position = "bottom")
 @
-\includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
+
+<<madmen_ex_ggnetwork, eval=FALSE, echo=FALSE>>=
+ggplot(data = ggnetwork(mm.net), aes(x, y, xend = xend, yend = yend)) +
+  geom_edges(color = "grey50") +
+  geom_nodes(aes(colour = Gender), size = 4) + # label= TRUE, vjust = -0.6
+  geom_nodetext(aes(colour = Gender, label = vertex.names), size = 4, vjust = -0.6) +
+  scale_colour_manual(values = c( "#FF69B4", "#0099ff")) +
+  xlim(c(-0.05, 1.05)) +
+  theme_blank() +
+  theme(legend.position = "bottom")
+@
+\includegraphics[width=\textwidth]{figure/madmen_ex_ggnet-1.pdf}
 \caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romantic relationship.}
@@
 
-<<blood_ex, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
-library(ggnet)
-data(blood)
-
-# source code for geom_net
+<<blood_ex_ggnet, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
 ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(colour="darkred", layout="circle", label=TRUE, size=15,
-           directed=TRUE, vjust=0.5, labelcolour="grey80", curvature=0,
-           arrowsize=1.5, linewidth=0.75, arrowgap=0.05) + theme_net()
+  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
+           directed = TRUE, vjust = 0.5, labelcolour = "grey80", curvature = 0,
+           arrowsize = 1.5, linewidth = 0.75, arrowgap = 0.05) +
+  theme_net()
 @
@@
 <<blood_ex_ggnet2, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
-# source code for ggnet2
-library(GGally) # use developmental version on github for now
-
 ggnet2(network(blood$edges[, 1:2]), mode = "circle", size = 15,
-       label = TRUE, arrow.size = 7.5, arrow.gap=0.05, vjust = 0.5,
+       label = TRUE, arrow.size = 7.5, arrow.gap = 0.95, vjust = 0.5,
        node.color="darkred", label.color="grey80")
 @
+
+<<blood_ex_ggnetwork, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
+ggplot(ggnetwork(network(blood$edges[, 1:2]), layout = "circle"),
+       aes(x, y, xend = xend, yend = yend)) +
+  geom_edges(arrow = arrow(length = unit(7.5, "pt"), type = "closed")) +
+  geom_nodes(size = 15, color = "darkred") +
+  geom_nodetext(aes(label = vertex.names), color = "grey80") +
+  theme_blank()
+@
+
 This network is shown in figure \ref{fig.cap:blood_ex}.
@@
 Anyone looking at this plot can quickly determine which type(s) of blood they can receive and which type(s) can receive their blood.
+
 \begin{figure}
-\begin{subfigure}[t]{0.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}
 \caption{geom\_net}
-\includegraphics[width=\textwidth]{figure/blood_ex-1.pdf}
+\includegraphics[width=\textwidth]{figure/blood_ex_ggnet-1.pdf}
 \end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}
 \caption{ggnet2}
 \includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnetwork}
+\includegraphics[width=\textwidth]{figure/blood_ex_ggnetwork-1.pdf}
 \end{subfigure}
@@
 \subsection{Email Network}
+
   This email network comes from the 2014 VAST Challenge \citep{emailnet}.
@@
 ggnet2(em.net, size = 4, color = "curr_empl_type", palette = "Set1",
-       arrow.size = 1.5, edge.alpha = 0.5, color.legend = "Employment Type") +
+       arrow.size = 10, arrow.gap = 0.95, edge.alpha = 0.5, color.legend = "Employment Type") +
   theme(legend.position="bottom")
+@
+
+<<email_ex_ggnetwork, fig.cap = 'All emails within a company over a two week period.',echo=TRUE, out.width='\\textwidth'>>=
+ggplot(ggnetwork(em.net), aes(x, y, xend = xend, yend = yend)) +
+  geom_edges(alpha = 0.5, color = "grey50",
+             arrow = arrow(length = unit(10, "pt"), type = "closed")) +
+  geom_nodes(aes(color = curr_empl_type), size = 4) +
+  scale_color_brewer("Employment Type", palette = "Set1") +
+  theme_blank() +
+  theme(legend.position = "bottom")
 @
@@
 \includegraphics[width=\textwidth]{figure/email_ex_ggnet2-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.49\textwidth}
+\caption{ggnetwork}
+\includegraphics[width=\textwidth]{figure/email_ex_ggnetwork-1.pdf}
 \end{subfigure}
@@
 gridExtra::grid.arrange(grobs = g, nrow = 2)
+@
+
+<<email_ex_facet_ggnetwork, fig.height=4, fig.width=8, fig.cap = 'All emails within a company over a two week period facetted by day of the month.',echo=TRUE, out.width='\\textwidth'>>=
+em.day = subset(email$edges, nrecipients < 54)[, c("From", "to", "day") ]
+em.day = lapply(unique(em.day$day), function(x) subset(em.day, day == x)[, 1:2 ])
+em.day = lapply(em.day, network, directed = TRUE)
+for (i in 1:length(em.day)) {
+  em.day[[ i ]] %v% "curr_empl_type" = em.cet[ network.vertex.names(em.day[[ i ]]) ]
+  em.day[[ i ]] %n% "day" = unique(email$edges$day)[ i ]
+}
+
+g = list(length(em.day))
+for (i in 1:length(em.day)) {
+  g[[ i ]] = ggplot(ggnetwork(em.day[[ i ]]), aes(x, y, xend = xend, yend = yend)) +
+    geom_edges(color = "grey50",
+               arrow = arrow(length = unit(6, "pt"), type = "closed")) +
+    geom_nodes(aes(color = curr_empl_type)) +
+    scale_colour_brewer(palette = "Set1") +
+    theme_blank() +
+    theme(legend.position = "none") +
+    ggtitle(paste("Day", em.day[[ i ]] %n% "day"))
+}
+gridExtra::grid.arrange(grobs = g, nrow = 2)
 
@@
 \includegraphics[width=\textwidth]{figure/email_ex_facet_ggnet2-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{\textwidth}
+\caption{ggnetwork}
+\includegraphics[width=\textwidth]{figure/email_ex_facet_ggnetwork-1.pdf}
 \end{subfigure}
@@
 @
+<<theme_ex_ggnetwork, echo=TRUE, fig.cap= 'Inheritance structure for \\texttt{theme} elements in \\texttt{ggplot2}.'>>=
+ggplot(ggnetwork(te.net), aes(x, y, xend = xend, yend = yend)) +
+  geom_edges() +
+  geom_nodes(size = 12, color = "white") +
+  geom_nodetext(aes(size = size, label = vertex.names)) +
+  scale_size_continuous(range = c(3, 5)) +
+  guides(size = FALSE) +
+  theme_blank()
+@
 
@@
 \includegraphics[width=\textwidth]{figure/theme_ex_ggnet2-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/theme_ex_ggnetwork-1.pdf}
 \end{subfigure}
@@
 @
+
+<<madmen2_ex_ggnetwork, echo=TRUE, fig.cap= 'A directed network of relationships in Mad Men.'>>=
+ggplot(ggnetwork(mm.dir, layout = "fruchtermanreingold"),
+       aes(x, y, xend = xend, yend = yend)) +
+  geom_edges(color = "grey50", arrow = arrow(length = unit(10, "pt"), type = "closed")) +
+  geom_nodes(size = 3, color = mm.col[ mm.dir %v% "gender" ]) +
+  geom_nodetext(aes(label = vertex.names), color = mm.col[ mm.dir %v% "gender" ], hjust = -0.1) +
+  theme_blank() +
+  theme(legend.position = "bottom")
+@
 \begin{figure}
@@
 \end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnetwork}
+\includegraphics[width=\textwidth]{figure/madmen2_ex_ggnetwork-1.pdf}
+\end{subfigure}
+
 \caption{\label{fig.cap:madmen2_ex} A directed network of relationships in Mad Men.}
@@
 
-fb.net = network(football$edges, directed = FALSE)
+fb.net = network(football$edges[, 1:2], directed = FALSE)
+set.edge.attribute(fb.net, "same.conf", football$edges$same.conf)
 fb.net %v% "conf" = football$vertices[ network.vertex.names(fb.net), "value" ]
@@
 ggnet2(fb.net, mode = "fruchtermanreingold",
-       color = "conf", color.legend = "Conference", palette = "Paired")
-#       edge.color = c("color", "grey75"))
+       color = "conf", color.legend = "Conference", palette = "Paired",
+       edge.color = c("color", "grey75"))
+@
+
+<<football_ex_ggnetwork, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
+ggplot(ggnetwork(fb.net, layout = "fruchtermanreingold"),
+       aes(x, y, xend = xend, yend = yend)) +
+  geom_edges(aes(linetype = as.logical(same.conf)), color = "grey50") +
+  geom_nodes(aes(color = conf), size = 4) +
+  scale_color_brewer("Conference", palette = "Paired") +
+  guides(linetype = FALSE) +
+  theme_blank()
 @
@@
 \end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{ggnetwork}
+\includegraphics[width=\textwidth]{figure/football_ex_ggnetwork-1.pdf}
+\end{subfigure}
 \caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
@@
 <<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
-
 ggplot(data = lesmisnet, aes(from_id = from, to_id = to, linewidth = degree/5 + 0.1 )) +

### 2015-10-12 inst/paper.Rnw
blood type example cleaned up

@@
 providing a wrapper for the user to visualize a network with ggplot2 elements, and (ii) implement networks as an internal layer of ggplot2.
-We will discuss both of these approaches in this paper.
+For the second approach, two main implementations exist.
+We will discuss all of these  in this paper.
 }
@@
 
-\section{Two approaches of network visualizations}
+\section{Three approaches of network visualizations}
  The three necessary elements of any network visualization are the vertices, the edges, and the layout.
@@
 \subsection{ggnet}
-\hh{XXX needs massive changes :)}
+\hh{XXX needs massive changes. Francois, could you take care of that?}
 \par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these three necessary graph attributes.
@@
 
+\subsection{geom\_node and geom\_edge}
+
+\hh{Fraancois, could you include a discussion of the background of the ggnetwork pacakge here?}
+
 \section{Examples of Networks}
@@
 
-<<blood_ex_ggnet, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
-ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
-           directed = TRUE, vjust = 0.5, labelcolour = "grey80", curvature = 0,
-           arrowsize = 1.5, linewidth = 0.75, arrowgap = 0.05) +
-  theme_net()
-@
-
 <<blood_ex_ggnet2, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
@@
        node.color="darkred", label.color="grey80")
+@
+
+<<blood_ex_ggnet, echo=TRUE, fig.cap= "Directed network of blood type donations in humans.", fig.width=6, fig.height=6>>=
+ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
+  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
+           directed = TRUE, vjust = 0.5, labelcolour = "grey80",
+           arrowsize = 1.5, linewidth = 0.75, arrowgap = 0.05) +
+  theme_net()
 @
@@
 \begin{figure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{.32\textwidth}
+\caption{ggnet2}
+\includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{.32\textwidth}
 \caption{geom\_net}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
-\caption{ggnet2}
-\includegraphics[width=\textwidth]{figure/blood_ex_ggnet2-1.pdf}
-\end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{.32\textwidth}
 \caption{ggnetwork}

### 2015-10-13 inst/paper.Rnw
bike sharing example

@@
   tidy = FALSE, echo=FALSE, cache=FALSE, eval = TRUE,
-  message = FALSE, warning = FALSE,
+  message = FALSE, warning = FALSE, highlight = FALSE,
   fig.height = 8, fig.width = 8, fig.align = 'center', fig.show='hide'

### 2015-10-13 R/geom-net.r

@@ -144,7 +144,6 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   draw_panel = function(data, panel_scales, coord,  ecolour="grey60", ealpha=NULL, arrowgap=0.01,
                         directed=FALSE, arrowsize=1, label=FALSE, labelcolour=NULL) {
 
-#    browser()
     data$self <- as.character(data$to) == as.character(data$from)
     edges <- data.frame(
       x = data$x,

### 2015-10-13 R/geom-net.r
open arrows when curvature is not zero

@@ -164,7 +164,10 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
 
     arrow = NULL
     if (directed) {
-      arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="closed")
+      if (any(data$curvature != 0))
+        arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="open")
+      else
+        arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="closed")
 
       arrowgap <- with(edges, arrowgap/sqrt((xend-x)^2+(yend-y)^2))
       edges <- transform(edges,

### 2015-10-13 R/geom-net.r
default arrow size set in point

@@ -29,8 +29,8 @@
 #' p <- ggplot(data = bloodnet, aes(from_id = from, to_id = to))
 #' p + geom_net()
 #' p + geom_net(aes(colour=rho))
-#' p + geom_net(aes(colour=rho), label=TRUE)
-#' p + geom_net(aes(colour=rho), label=TRUE, vjust=-0.5, labelcolour="black", directed=TRUE, arrowgap=0.01) + theme_net()
+#' p + geom_net(aes(colour=rho), label=TRUE, vjust = -0.5)
+#' p + geom_net(aes(colour=rho), label=TRUE, vjust=-0.5, labelcolour="black", directed=TRUE, curvature=0.2) + theme_net()
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 0, linewidth=.75, directed = TRUE)
@@ -165,9 +165,9 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
     arrow = NULL
     if (directed) {
       if (any(data$curvature != 0))
-        arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="open")
+        arrow = arrow(length = unit(arrowsize*10,"points"), type="open")
       else
-        arrow = arrow(length = unit(arrowsize*0.3,"cm"), type="closed")
+        arrow = arrow(length = unit(arrowsize*10,"points"), type="closed")
 
       arrowgap <- with(edges, arrowgap/sqrt((xend-x)^2+(yend-y)^2))
       edges <- transform(edges,

### 2015-10-15 R/geom-net.r
draw_key function has now a third parameter, size

@@ -124,7 +124,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
                              size = 4, fill = NA, alpha = NA, stroke = 0.5,
                              linewidth=1, angle=0, vjust=0, hjust=0.5, curvature = 0),
 
-  draw_key = function(data, params)  {
+  draw_key = function(data, params, size)  {
     with(data, grobTree(
       grid::pointsGrob(0.5, 0.5, pch = data$shape,
                        gp = grid::gpar(col = alpha(data$colour, data$alpha),

### 2015-10-15 inst/paper.Rnw
default arrow size set in point
bug fix

@@
 We can easily accommodate for these nodes by adding artificial edges in the data set, that have missing information for the `to' column.
-Fortunately, R provides functions that allow for an easy way of producing the required result: both \code{merge} and \code{join} can be used.
+Fortunately, R provides function that allow for an easy way of producing the required result: both \code{merge} and \code{join} can be used.
 In \code{merge} the paramter all needs to be set to TRUE, in \code{join}, the parameter setting has to be \code{type=`full'}, and \code{all=TRUE}.
@@
   geom_net(aes(fontsize=degree), directed = TRUE, label = TRUE, vjust=-.5,
-           size=3, ecolour="grey70") +
+           size=3, ecolour="grey70", curvature=0.25) +
   theme_net() + xlim(c(-0.05, 1.05))

### 2015-10-15 R/geom-net.r

@@ -137,10 +137,16 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
   },
 
   setup_data = function(data, params, mapping) {
+
+    data$from <- as.character(data$from)
+    data$to <- as.character(data$to)
     selfie <- (data$from == data$to) & (params$selfies == TRUE)
   # maximum radius is at the moment hard coded to 0.05
     data$ymax = max(with(data, pmax(y, yend) + 2*0.05*selfie))
     data$xmax = with(data, pmax(x, xend) + 2*0.05*selfie)
+
+    data$from <- factor(data$from)
+    data$to <- factor(data$to)
 
     data
   },

### 2016-10-19 inst/doc/comparing-graph-drawing-speed.Rmd
add another vignette for paper code examples

---
title: "Speed comparisons of graph drawing packages"
author: "Samantha Tyner, François Briatte, and Heike Hofmann"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Speed comparisons of graph drawing packages}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction

This vignette provides access to the online supplement to the paper ``Network Visualization with `ggplot2` ``.

The code to access data for all of the examples and to reproduce all of the figures is available at [tyner-briatte-hofmann.R](https://raw.githubusercontent.com/sctyner/ggnet-paper/master/tyner-briatte-hofmann.R).

A summary of the speed comparisons is shown in the paper. A more detailed outline together with the code to reproduce findings is given in the next section.

# Speed Comparisons

This vignette compares the speeds of our package to the other `ggplot2`-based network visualization package `ggnetwork`, the `GGally` function `ggnet2()`, and the visualization capabilities of the `igraph` and `network` packages.

## Speed Comparison 1

In this first example, we visualize the protein-protein interaction network in the yeast species *S. cerevisiae* using several different methods.

First, load all of the necessary packages and data sets.

```{r speed1_loadpkg, echo = TRUE, eval = FALSE, message=FALSE, warning = FALSE}
## igraph (current: v1.0.1)
if (!require(igraph, quietly = TRUE)) {
  install.package("igraph")
}
library(igraph)

## network (current: v1.13.0)
if (!require(network, quietly = TRUE)) {
  install.package("network")
}
library(network)

## ggnet2
if (!require(GGally, quietly = TRUE)) {
  install.packages("GGally")
}
library(GGally)

## geom_net
library(geomnet) # also currently requires dplyr

## ggnetwork
if (!require(ggnetwork, quietly = TRUE)) {
  install.packages("ggnetwork")
}
library(ggnetwork)

## pre-loading
library(sna)

```

```{r, echo = TRUE, message = FALSE}
## ggplot2
library(ggplot2)

## other packages
library(dplyr)
library(tidyr)

## load data 
data(protein, package = "geomnet")
```

Next, generate the network visualization of this network 100 times using the random layout option in each package and store the time to plot each run.

```{r speed1_simcode, eval=FALSE, echo=TRUE}
d = data.frame()

if (!file.exists("runtimes-protein-100.csv")) {

  for (i in 1:100) {

    cat("Iteration", sprintf("%3.0f", i), "/ 100\n")

    n = as.matrix(protein$edges[, 1:2])
    n = igraph::graph_from_edgelist(n, directed = FALSE)

    t1 = system.time({
      plot(n, vertex.label = NA, layout = layout_randomly)
    })[1]

    n = network(protein$edges[, 1:2], directed = FALSE)

    t2 = system.time({
     plot(n, coord = gplot.layout.random(n, NULL))
    })[1]

    t3 = system.time({
      print(ggnet2(n, mode = "random"))
    })[1]

    t4 = system.time({
      print(ggplot(data = protein$edges, aes(from_id = from, to_id = to)) +
              geom_net(layout = "random"))
    })[1]

    t5 = system.time({
      print(ggplot(ggnetwork(n, layout = "random"),
                   aes(x, y, xend = xend, yend = yend)) +
              geom_edges() +
              geom_nodes())
    })[1]

    d = rbind(d, data.frame(
      iteration = i,
      igraph = t1,
      network = t2,
      ggnet2 = t3,
      geomnet = t4,
      ggnetwork = t5,
      row.names = NULL
    ))

  }

  write.csv(d, file = "runtimes-protein-100.csv", row.names = FALSE)

}
```

Finally, compare the 100 runtimes for each method side-by-side. In the resulting plot below, we use our own data, which can be found [here](https://raw.githubusercontent.com/sctyner/ggnet-paper/master/runtimes-protein/runtimes-protein-100.csv).

```{r speed1_cheatdata, echo=FALSE}
load("../inst/runtimes-protein.rda")
```
```{r speed1_viewdata, echo = TRUE, eval = FALSE}
link_to_data <- "runtimes-protein-100.csv"
g = read.csv(link_to_data, stringsAsFactors = FALSE) %>%
      gather(`Visualization approach`, time, -iteration)
```
```{r speed1_plot, echo=TRUE}
ggplot(g, aes(x = `Visualization approach`, y = time,
              fill = `Visualization approach`)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set2") +
  coord_flip() +
  labs(y = "\nPlotting time (in seconds)", x = "Implementation\n") +
  guides(fill = FALSE) 
```

# Speed Comparison 2

The figure summarizes the results of these benchmarks using a convenience sample of machines accessible to the authors, including authors' hardware and additional results from friends' and colleagues' machines. The data containing runtimes of for all of these machines can be found [here](https://github.com/sctyner/ggnet-paper/tree/master/runtimes).

This time, we visualize graphs of various sizes that we generated randomly with edge probability 0.2 using the `sna` package. The sizes we generated are overall fairly small, from 25 to 250 nodes, going up by 25 each time. Each network visualization method was test 100 times in total for each size of network.

```{r speed_2code, eval = FALSE}
# note: plot.network might be slower because it actually prints the plots, so to
# compare the methods, we need to evaluate the ggplot plots by printing them too

# create 10 files per size 
for (k in 0:9) {

  # the 10 different sizes   
  for (i in seq(250, 25, -25)) {

    f = paste0("runtimes-", sprintf("%04.0f", i), "-", k, ".csv")

    if (!file.exists(f)) {

      d = data.frame()

      # in each file store the runtimes of plotting 10 random graphs 
      # do this for each of the 5 methods 
      for (j in 1:10) {

        r = sna::rgraph(i, tprob = 0.2)

        cat("Network size", i,
            "iteration", sprintf("%3.0f", 10 * k + j), "/ 100\n")

        n = igraph::graph_from_adjacency_matrix(r, mode = "undirected")

        t1 = system.time({
          plot(n, vertex.label = NA)
        })[1]

        n = network::network(r, directed = FALSE)

        t2 = system.time({
          plot.network(n)
        })[1]

        t3 = system.time({
          print(ggnet2(n))
        })[1]

        e = data.frame(sna::as.edgelist.sna(n))

        t4 = system.time({
          print(ggplot(data = e) +
                  geom_net(aes(from_id = X1, to_id = X2)))
        })[1]

        t5 = system.time({
          print(ggplot(ggnetwork(n),
                       aes(x, y, xend = xend, yend = yend)) +
                  geom_edges() +
                  geom_nodes())
        })[1]

        d = rbind(d, data.frame(
          network_size = i,
          iteration = 10 * k + j,
          igraph = t1,
          network = t2,
          ggnet2 = t3,
          geomnet = t4,
          ggnetwork = t5,
          row.names = NULL
        ))

        dev.off()

      }

      write.csv(d, f, row.names = FALSE)

    }

  }
```

Again, for this plot we use our data that is available on Github. Network sizes are plotted horizontally, execution times of 100 runs under each visualization approach are plotted on the $y$-axis. Each panel shows a different machine as indicated by the facet label. Note that each panel is scaled separately to account for differences in the overall speed of these machines.

```{r speed2_code, echo = TRUE, eval=FALSE}  
# get all files containing runtime information and combine them into a single data frame.
g = list.files(pattern = "runtimes-(.*)csv$") %>%
        lapply(read.csv, stringsAsFactors = FALSE) %>%
        bind_rows %>%
        gather(`Visualization approach`, time, -network_size, -iteration) %>%
        group_by(network_size, `Visualization approach`) %>%
        summarise(mean_time = mean(time),
                  q05 = quantile(time, 0.05),
                  q95 = quantile(time, 0.95))
```

```{r all-runtimes, eval=TRUE, echo=FALSE, fig.width = 8, fig.height = 8}
# link_to_full_comparison <- "https://raw.githubusercontent.com/sctyner/ggnet-paper/master/data/compare-all.csv"
# g <- readr::read_csv(link_to_full_comparison)
load("../inst/compare-all.rda")
qplot(data = g, network_size, time, colour = `Visualization approach`, alpha = I(0.1)) +
  facet_wrap(facets = ~setup, scales = "free_y") +
  geom_smooth(fill="white") +
  scale_colour_brewer("Implementation", palette = "Set1") +
  xlab("Network size (edge probability p = 0.2)") +
  ylab("Time (in seconds)") +
  theme_bw() +
  theme(legend.position = "bottom")
```

What these plots indicate is that we have surprisingly large variability in relative run times across different machines.
However, the results support some  general findings.

The  plotting routine in the network package  is by far the slowest across all machines, while the  plotting with `igraph` is generally among the fastest. Our three approaches generally feature in between these two, with `ggnet2` being as fast or faster than `igraph` plotting, followed by `ggnetwork` and `geomnet`, which is generally the slowest among the three. These differences become more pronounced as the size of the network increases.

### 2016-10-19 inst/doc/comparing-graph-drawing-speed.R

@@ -1,0 +1,222 @@
+## ----setup, echo=FALSE---------------------------------------------------
+
+knitr::opts_chunk$set(fig.width = 6,
+                      fig.height = 4,
+                      fig.align='center',
+                      dev = "png")
+
+
+## ----speed1_loadpkg, echo = TRUE, eval = FALSE, message=FALSE, warning = FALSE----
+#  ## igraph (current: v1.0.1)
+#  if (!require(igraph, quietly = TRUE)) {
+#    install.package("igraph")
+#  }
+#  library(igraph)
+#  
+#  ## network (current: v1.13.0)
+#  if (!require(network, quietly = TRUE)) {
+#    install.package("network")
+#  }
+#  library(network)
+#  
+#  ## ggnet2
+#  if (!require(GGally, quietly = TRUE)) {
+#    install.packages("GGally")
+#  }
+#  library(GGally)
+#  
+#  ## geom_net
+#  library(geomnet) # also currently requires dplyr
+#  
+#  ## ggnetwork
+#  if (!require(ggnetwork, quietly = TRUE)) {
+#    install.packages("ggnetwork")
+#  }
+#  library(ggnetwork)
+#  
+#  ## pre-loading
+#  library(sna)
+#  
+
+## ---- echo = TRUE, message = FALSE---------------------------------------
+## ggplot2
+library(ggplot2)
+
+## other packages
+library(dplyr)
+library(tidyr)
+
+## load data 
+data(protein, package = "geomnet")
+
+## ----speed1_simcode, eval=FALSE, echo=TRUE-------------------------------
+#  d = data.frame()
+#  
+#  if (!file.exists("runtimes-protein-100.csv")) {
+#  
+#    for (i in 1:100) {
+#  
+#      cat("Iteration", sprintf("%3.0f", i), "/ 100\n")
+#  
+#      n = as.matrix(protein$edges[, 1:2])
+#      n = igraph::graph_from_edgelist(n, directed = FALSE)
+#  
+#      t1 = system.time({
+#        plot(n, vertex.label = NA, layout = layout_randomly)
+#      })[1]
+#  
+#      n = network(protein$edges[, 1:2], directed = FALSE)
+#  
+#      t2 = system.time({
+#       plot(n, coord = gplot.layout.random(n, NULL))
+#      })[1]
+#  
+#      t3 = system.time({
+#        print(ggnet2(n, mode = "random"))
+#      })[1]
+#  
[…]

### 2016-10-19 inst/doc/examples-from-paper.R

@@ -1,0 +1,551 @@
+## ----load_packages, echo=TRUE--------------------------------------------
+library(dplyr)
+library(tidyr)
+library(ggplot2) # needs to be version ≥ 2.1.0
+library(scales)
+
+## ggnet2
+# if (!require(ggnet, quietly = TRUE)) {
+#   devtools::install_github("briatte/ggnet")
+# }
+# library(ggnet)
+if (!require(GGally, quietly = TRUE)) {
+  devtools::install_github("ggobi/ggally")
+}
+
+## geom_net
+if (!require(geomnet, quietly = TRUE)) {
+  devtools::install_github("sctyner/geomnet")
+}
+
+## ggnetwork
+if (!require(ggnetwork, quietly = TRUE) ||
+    packageVersion("ggnetwork") < "0.2.1") {
+  devtools::install_github("briatte/ggnetwork")
+}
+
+## pre-load
+library(network)
+library(sna)
+library(GGally)
+library(geomnet)
+library(ggnetwork)
+
+## ----madmen_geom_net, fig.width=7, fig.height=7.5------------------------
+# make data accessible 
+data(madmen, package = "geomnet")
+
+# code for geom_net
+# data step: merge edges and nodes by the "from" column
+MMnet <- merge(madmen$edges, madmen$vertices,
+               by.x = "Name1", by.y = "label", all = TRUE)
+# create plot
+set.seed(10052016)
+ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2)) +
+  geom_net(aes(colour = Gender), layout = "kamadakawai", 
+           size = 2, label = TRUE, vjust = -0.6, ecolour = "grey60",
+           directed =FALSE, fontsize = 3, ealpha = 0.5) +
+  scale_colour_manual(values = c("#FF69B4", "#0099ff")) +
+  xlim(c(-0.05, 1.05)) +
+  theme_net() +
+  theme(legend.position = "bottom")
+
+## ----madmen_ggnet2, eval=FALSE, echo=TRUE, fig.cap="The code required to generate the Mad Men example using the `ggnet2` function in the `GGally` package."----
+#   library(GGally)
+#   library(network)
+#   # make the data available
+#   data(madmen, package = 'geomnet')
+#   # data step for both ggnet2 and ggnetwork
+#   # create undirected network
+#   mm.net <- network(madmen$edges[, 1:2], directed = FALSE)
+#   # create node attribute (gender)
[…]

### 2016-11-04 R/geom-net.r
change label argument to "labelon"

@@ -132,13 +132,13 @@ geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "ide
                       layout="kamadakawai", layout.par=list(), directed = FALSE, fiteach=FALSE,  selfloops = FALSE,
                       alpha = 0.25,
                       ecolour=NULL, ealpha=NULL, arrow=NULL, arrowgap=0.01, arrowsize=1,
-                      label=FALSE, labelcolour=NULL, labelgeom = 'text', repel = FALSE,
+                      labelon=FALSE, labelcolour=NULL, labelgeom = 'text', repel = FALSE,
                        vertices=NULL, ...) {
 
     ggplot2::layer(
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
-    params = list(na.rm = na.rm, layout=layout, layout.par=layout.par, fiteach=fiteach, label=label, labelgeom=labelgeom,
+    params = list(na.rm = na.rm, layout=layout, layout.par=layout.par, fiteach=fiteach, labelon=labelon, labelgeom=labelgeom,
                   ecolour = ecolour, ealpha=ealpha, arrow=arrow, arrowgap=arrowgap, directed=directed, repel = repel,
                   arrowsize=arrowsize,
                   labelcolour=labelcolour, vertices=vertices, selfloops = selfloops,
@@ -204,7 +204,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
 
   draw_panel = function(data, panel_scales, coord,  ecolour=NULL, ealpha=NULL, arrow=NULL, arrowgap=0.01,
                         directed=FALSE, arrowsize=1, repel = FALSE,
-                        label=FALSE, labelgeom='text', labelcolour=NULL, selfloops = FALSE) {
+                        labelon=FALSE, labelgeom='text', labelcolour=NULL, selfloops = FALSE) {
 
  #
     data$self <- as.character(data$to) == as.character(data$from)
@@ -293,7 +293,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
     }
 
     label_grob <- NULL
-    if (label | !is.null(data$label)) {
+    if (labelon | !is.null(data$label)) {
       labels <- data.frame(
         x = data$x,
         y = data$y,

### 2016-11-04 R/geom-net.r
change layout argument to layout.alg for layout algorithm

@@ -10,7 +10,7 @@
 #' @param layout character value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?gplot.layout} in the package sna for other choices.
 #' @param layout.par list of parameters detailing algorithmic specs. Default parameters are used initially. See \code{?gplot.layout} in the package sna for other choices.
 #' @param fiteach logical value. Should the network be fit in each of the panels separately, or is there going to be one fit for all?
-#' @param label logical value. Include labels for (all) nodes. labelcolour specifies colour of labels, if they should not be the same as the nodes. labels are taken from the from_id variable, unless a label variable is given.
+#' @param labelon logical value. Include labels for (all) nodes. labelcolour specifies colour of labels, if they should not be the same as the nodes. labels are taken from the from_id variable, unless a label variable is given.
 #' @param labelcolour character of colour for the labels.
 #' @param labelgeom character. Which ggplot2 \code{geom} to use to draw the labels. Either "text" or "label".
 #' @param repel logical value. If \code{TRUE}, uses the ggrepel package geoms to draw the node labels instead of the ggplot2 geoms.
@@ -34,8 +34,8 @@
 #' p <- ggplot(data = bloodnet, aes(from_id = from, to_id = to))
 #' p + geom_net()
 #' p + geom_net(aes(colour=rho)) + theme_net()
-#' p + geom_net(aes(colour=rho), label=TRUE, vjust = -0.5)
-#' p + geom_net(aes(colour=rho), label=TRUE, vjust=-0.5, labelcolour="black",
+#' p + geom_net(aes(colour=rho), labelon=TRUE, vjust = -0.5)
+#' p + geom_net(aes(colour=rho), labelon=TRUE, vjust=-0.5, labelcolour="black",
 #'              directed=TRUE, curvature=0.2) + theme_net()
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6)
 #' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
@@ -58,11 +58,11 @@
 #' data(madmen)
[…]

### 2016-11-04 R/stat-net.r

@@ -55,7 +55,7 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
     data
   },
 
-compute_network = function(data, layout="kamadakawai", layout.par=list()) {
+compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
 # cat("compute_network\n")
 #  browser()
     require(dplyr)
@@ -78,7 +78,7 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
 #    m <- network::as.matrix.network.adjacency(net)
 #  }
 
-  if (is.null(layout)) {
+  if (is.null(layout.alg)) {
     if (is.null(data$x) || is.null(data$y)) stop("If no layout mechanism is specified, x and y coordinates have to be given\n\n")
     vert.coord <- data[, c("x", "y", "from_id")]
     vert.coord <- subset(vert.coord, from_id %in% attr(edgelist, "vnames"))
@@ -89,7 +89,7 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
   } else {
   #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
   require(sna)
-  layoutFun <- paste('gplot.layout.',layout,sep='')
+  layoutFun <- paste('gplot.layout.',layout.alg,sep='')
   vert.coord <- data.frame(do.call(layoutFun, list(edgelist, layout.par = layout.par)))
 
   vert.coord$label <- attr(edgelist, "vnames") #row.names(m)
@@ -119,12 +119,12 @@ compute_network = function(data, layout="kamadakawai", layout.par=list()) {
 },
 
 compute_panel = function(self, data, scales, na.rm = FALSE,
-                         layout="kamadakawai", layout.par=list(),
+                         layout.alg="kamadakawai", layout.par=list(),
                          fiteach=FALSE, vertices=NULL) {
 #  cat("compute_panel in stat_net\n")
 #  browser()
 #    if (fiteach)
-      data <- self$compute_network(data, layout=layout, layout.par=layout.par)
+      data <- self$compute_network(data, layout.alg =layout.alg, layout.par=layout.par)
 
     #    data <- plyr::ddply(data, "group", plyr::mutate, .samegroup = to %in% unique(from))
     if (any(data$group) != -1)
@@ -134,7 +134,7 @@ compute_panel = function(self, data, scales, na.rm = FALSE,
     data.frame(data)
   },
 
-compute_layer = function(self, data, params, panel, na.rm = FALSE, layout,
[…]

### 2016-11-04 R/geom-net.r
examples still throwing weird errors

@@ -37,16 +37,16 @@
 #' p + geom_net(aes(colour=rho), labelon=TRUE, vjust = -0.5)
 #' p + geom_net(aes(colour=rho), labelon=TRUE, vjust=-0.5, labelcolour="black",
 #'              directed=TRUE, curvature=0.2) + theme_net()
-#' p + geom_net(colour = "orange", layout = 'circle', size = 6)
-#' p + geom_net(colour = "orange", layout = 'circle', size = 6, linewidth=.75)
-#' p + geom_net(colour = "orange", layout = 'circle', size = 0, linewidth=.75,
+#' p + geom_net(colour = "orange", layout.alg = 'circle', size = 6)
+#' p + geom_net(colour = "orange", layout.alg = 'circle', size = 6, linewidth=.75)
+#' p + geom_net(colour = "orange", layout.alg = 'circle', size = 0, linewidth=.75,
 #'              directed = TRUE)
 #' p + geom_net(aes(size=Predominance, colour=rho, shape=rho, linetype=group_to),
-#'              linewidth=0.75, label =TRUE, labelcolour="black") +
+#'              linewidth=0.75, labelon =TRUE, labelcolour="black") +
 #'     facet_wrap(~Ethnicity) +
 #'     scale_colour_brewer(palette="Set2")
 #' gg <- ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-#'   geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
+#'   geom_net(colour = "darkred", layout.alg = "circle", labelon=TRUE, size = 15,
 #'          directed = TRUE, vjust = 0.5, labelcolour = "grey80",
 #'          arrowsize = 1.5, linewidth = 0.5, arrowgap = 0.05,
 #'          selfloops = TRUE, ecolour = "grey40") +
@@ -68,7 +68,7 @@
 #'
 #' p <- ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2))
 #' # alternative labelling: specify label variable.
-#' p + geom_net(aes(colour=Gender, label=Gender), size=6, linewidth=1, fontsize=3,
+#' p + geom_net(aes(colour=Gender, labelon=Gender), size=6, linewidth=1, fontsize=3,
 #'              labelcolour="black")
 #'
 #' ## visualizing ggplot2 theme elements

### 2016-11-12 R/reshapeNet.R
add "reshapeNet" function; update documentation

@@ -1,0 +1,110 @@
+#' Function for converting various network structures into the correct format for use with geomnet
+#' 
+#' @param net. A network object. Can be a class "network", "igraph", "matrix" (for adjaceny matrices), or "data.frame" object. If net is a "data.frame", must provide the ndata argument and the first column of net must be the "from" node column.
+#' @param ndata. Data frame containing network node list and other node information. Only applicable for class "data.frame." First column name should be node ids. 
+#' 
+#' @examples 
+#' 
+#' # class network (sna, network, statnet packages)
+#' 
+#' library(network)
+#' data(emon)
+#' reshapeNet(net = emon$Cheyenne)
+#' 
+#' # class igraph (igraph, igraphdata packages)
+#' library(igraph)
+#' library(igraphdata)
+#' data("USairports", package = "igraphdata")
+#' head(reshapeNet(net = USairports))
+#' 
+#' # class matrix (for adjacency matrices)
+#' 
+#' adjmat <- network::as.matrix.network.adjacency(emon$MtSi)
+#' str(adjmat)
+#' reshapeNet(net = adjmat)
+#' 
+#' # eclass data.frame and ndata 
+#' data(blood)
+#' reshapeNet(net = blood$edges, ndata = blood$vertices)
+#' 
+#' @export
+reshapeNet <- function(net, ...){
+  UseMethod("reshapeNet", net)
+}
+#' @export
+reshapeNet.network <- function(net, ...){
+  require(network)
+  node.attr <- network::list.vertex.attributes(net)
+  edge.attr <- network::list.edge.attributes(net)
+  N <- network::network.size(net) 
+  P <- length(node.attr)
+  node.data <- data.frame(matrix("", nrow = N, ncol = P+1), stringsAsFactors = F)
+  names(node.data) <- c("ID", node.attr)
+  node.data$ID <- 1:N
+  for (i in 1:P){
+    node.data[,(i+1)] <- network::get.vertex.attribute(net, node.attr[i])
+  }
+  NE <- nrow(network::as.edgelist(net))  
+  P2 <- length(edge.attr)
+  edge.data <- data.frame(network::as.edgelist(net), matrix("", nrow = NE, ncol = P2), stringsAsFactors = F)
+  names(edge.data) <- c("from", "to", edge.attr)
+  
+  for (i in 1:P2){
[…]

### 2016-11-12 R/geom-net.r
layout -> layout.alg in Rmd

@@ -158,7 +158,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
                              linewidth=1, angle=0, vjust=0, hjust=0.5, curvature = 0),
 
   draw_key = function(data, params, size)  {
-#browser()
+# browser()
     draw_arrow <-  NULL
     if (params$directed) {
       if (any(data$curvature != 0)){
@@ -168,9 +168,9 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
         draw_arrow <- arrow(length = unit(params$arrowsize*5,"points"), type="closed")
       }
    }
-      
+
     with(data, grobTree(
-      grid::pointsGrob(x = c(.15, .85), y = c(.85, .15), 
+      grid::pointsGrob(x = c(.15, .85), y = c(.85, .15),
                        pch = data$shape, size = unit(data$size, "points"),
                        gp = grid::gpar(col = alpha(data$colour, data$alpha),
                                        fill = alpha(data$fill, data$alpha),
@@ -178,14 +178,14 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
                                        lwd = data$stroke * .stroke/2)
                        ),
      grid::segmentsGrob(x0 = .15, y0 = .85 ,
-                        x1 = ifelse(is.null(draw_arrow), .85, .82), 
+                        x1 = ifelse(is.null(draw_arrow), .85, .82),
                         y1 = ifelse(is.null(draw_arrow), .15, .18),
                      gp = grid::gpar(
                        col = alpha(data$colour, data$alpha),
                        fill = alpha(data$colour, data$alpha),
-                       lwd = data$linewidth, 
+                       lwd = data$linewidth,
                        lty = data$linetype,
-                       lineend="butt"), 
+                       lineend="butt"),
                      arrow = draw_arrow
                      )
 
@@ -221,7 +221,7 @@ GeomNet <- ggplot2::ggproto("GeomNet", ggplot2::Geom,
       y = data$y,
       yend = data$yend,
       weight = data$weight,
-      colour = ecolour %||% ifelse(data$.samegroup, data$colour, "grey60"),
+      colour = ecolour %||% ifelse(data$.samegroup, data$colour, "grey40"),
       size = data$linewidth %||% (data$size / 4),
       nodesize = data$size,
       alpha = ealpha %||% data$alpha,

### 2016-11-14 R/fortify-net.R
change reshapeNet to a fortify method

@@ -1,7 +1,7 @@
 #' Function for converting various network structures into the correct format for use with geomnet
 #' 
-#' @param net. A network object. Can be a class "network", "igraph", "matrix" (for adjaceny matrices), or "data.frame" object. If net is a "data.frame", must provide the ndata argument and the first column of net must be the "from" node column.
-#' @param ndata. Data frame containing network node list and other node information. Only applicable for class "data.frame." First column name should be node ids. 
+#' @param model. A network object. Can be a class "network", "igraph", "matrix" (for adjaceny matrices), or "data.frame" object. If net is a "data.frame", must provide the data argument and the first column of net must be the "from" node column.
+#' @param data. Data frame containing network node list and other node information. Only applicable for class "data.frame." First column name should be node ids. 
 #' 
 #' @examples 
 #' 
@@ -9,30 +9,27 @@
 #' 
 #' library(network)
 #' data(emon)
-#' reshapeNet(net = emon$Cheyenne)
+#' fortify(emon$Cheyenne)
 #' 
 #' # class igraph (igraph, igraphdata packages)
 #' library(igraph)
 #' library(igraphdata)
 #' data("USairports", package = "igraphdata")
-#' head(reshapeNet(net = USairports))
+#' head(fortify(USairports))
 #' 
 #' # class matrix (for adjacency matrices)
 #' 
 #' adjmat <- network::as.matrix.network.adjacency(emon$MtSi)
 #' str(adjmat)
-#' reshapeNet(net = adjmat)
+#' fortify(adjmat)
 #' 
 #' # eclass data.frame and ndata 
 #' data(blood)
-#' reshapeNet(net = blood$edges, ndata = blood$vertices)
+#' fortify(blood$edges, blood$vertices)
 #' 
 #' @export
-reshapeNet <- function(net, ...){
-  UseMethod("reshapeNet", net)
-}
-#' @export
-reshapeNet.network <- function(net, ...){
+fortify.network <- function(model, data = NULL, ...){
+  net <- model
   require(network)
   node.attr <- network::list.vertex.attributes(net)
   edge.attr <- network::list.edge.attributes(net)
@@ -57,7 +54,9 @@ reshapeNet.network <- function(net, ...){
   
   return(dat)
 }
-reshapeNet.igraph <- function(net, ...){
+#' @export
+fortify.igraph <- function(model, data = NULL, ...){
[…]

### 2016-11-14 R/stat-net.r
singletons are plotted

@@ -26,6 +26,7 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
 
   setup_data = function(self, data, params) {
 #cat("setup_data stat_net\n")
+#    browser()
     fiteach=params$fiteach
     if (!is.factor(data$from_id)) data$from_id <- factor(data$from_id)
     if (!is.factor(data$to_id)) data$to_id <- factor(data$to_id)
@@ -37,8 +38,9 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
     # we want to keep all of the values that are NA in the second edge - give them a special value, so we can pull them out later
     levels <- levels(data$to_id)
     data$to_id <- as.character(data$to_id)
-    data$to_id[is.na(data$to_id)] <- "..NA.."
-    data$to_id <- factor(data$to_id, levels = c(levels, "..NA.."))
+    data$to_id[is.na(data$to_id)] <- data$from_id[is.na(data$to_id)]
+#    data$to_id[is.na(data$to_id)] <- "..NA.."
+#    data$to_id <- factor(data$to_id, levels = c(levels, "..NA.."))
 
     # check that all vertices are included in the data set
     only_to <- setdiff(levels(data$to_id), levels(data$from_id))

### 2016-11-14 R/geom-net.r
na.rm = FALSE

@@ -128,7 +128,7 @@
 #'   theme(legend.position="bottom")
 #'   }
 
-geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, na.rm = TRUE, inherit.aes = TRUE,
+geom_net <- function (mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, na.rm = FALSE, inherit.aes = TRUE,
                       layout.alg="kamadakawai", layout.par=list(), directed = FALSE, fiteach=FALSE,  selfloops = FALSE,
                       alpha = 0.25,
                       ecolour=NULL, ealpha=NULL, arrow=NULL, arrowgap=0.01, arrowsize=1,

### 2016-11-21 visual_test/net.r
visual test from fri

@@ -14,17 +14,17 @@ save_vtest("Blood example basic net")
 p + geom_net(aes(colour=rho)) + theme_net()
 save_vtest("Blood example nodes colored by rho factor")
 
-p + geom_net(aes(colour=rho), label=TRUE, vjust = 0.5, hjust = 0.5, size = 10,
+p + geom_net(aes(colour=rho), labelon=TRUE, vjust = 0.5, hjust = 0.5, size = 10,
              labelcolour = "grey10")
 save_vtest("Blood example nodes colored by rho factor, labelled")
 
-p + geom_net(aes(colour=rho), label=TRUE, vjust = 0.5, hjust = 0.5, size = 10,
+p + geom_net(aes(colour=rho), labelon=TRUE, vjust = 0.5, hjust = 0.5, size = 10,
              labelcolour = "grey10", arrowgap =0.05,
              directed=TRUE, curvature=0.2) + theme_net()
 save_vtest("Blood example nodes colored by rho factor, labelled, curved edges")
 
 p + geom_net(aes(size=Predominance, colour=rho, shape=rho, linetype=group_to),
-             linewidth=0.75, label =TRUE, labelcolour="black", size = 10,
+             linewidth=0.75, labelon =TRUE, labelcolour="black", size = 10,
              vjust = 0.5, hjust = 0.5) +
     facet_wrap(~Ethnicity) +
     scale_colour_brewer(palette="Set2") +
@@ -33,10 +33,10 @@ save_vtest("Blood example, facetted")
 
 
 gg <- ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
-  geom_net(colour = "darkred", layout = "circle", label = TRUE, size = 15,
+  geom_net(colour = "darkred", layout.alg = "circle", labelon = TRUE, size = 15,
          directed = TRUE, vjust = 0.5, labelcolour = "grey80",
          arrowsize = 1.5, linewidth = 0.5, arrowgap = 0.05,
-         selfies = TRUE, ecolour = "grey40") +
+         selfloops = TRUE, ecolour = "grey40") +
   theme_net()
 gg
 save_vtest("Blood example, with self references")
@@ -47,14 +47,14 @@ save_vtest("Blood example, with self references")
 data(madmen)
 MMnet <- merge(madmen$edges, madmen$vertices, by.x="Name1", by.y="label", all=TRUE)
[…]

### 2016-11-21 R/geom-net.r
update examples. remove vignettes from inst/doc

@@ -7,7 +7,7 @@
 #' @param ealpha numeric value of alpha blending of edges.
 #' @param na.rm If \code{FALSE} (the default), removes missing values with
 #'    a warning. If \code{TRUE} silently removes missing values.
-#' @param layout character value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?gplot.layout} in the package sna for other choices.
+#' @param layout.alg character value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?gplot.layout} in the package sna for other choices.
 #' @param layout.par list of parameters detailing algorithmic specs. Default parameters are used initially. See \code{?gplot.layout} in the package sna for other choices.
 #' @param fiteach logical value. Should the network be fit in each of the panels separately, or is there going to be one fit for all?
 #' @param labelon logical value. Include labels for (all) nodes. labelcolour specifies colour of labels, if they should not be the same as the nodes. labels are taken from the from_id variable, unless a label variable is given.
@@ -25,7 +25,7 @@
 #' @export
 #' @examples
 #' \dontrun{
-#' library(ggplot2)
+#' library(geomnet)
 #' data(blood)
 #' p <- ggplot(data = blood$edges, aes(from_id = from, to_id = to))
 #' p + geom_net(vertices=blood$vertices, aes(colour=..type..)) + theme_net()
@@ -68,7 +68,7 @@
 #'
 #' p <- ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2))
 #' # alternative labelling: specify label variable.
-#' p + geom_net(aes(colour=Gender, labelon=Gender), size=6, linewidth=1, fontsize=3,
+#' p + geom_net(aes(colour=Gender, label=Gender), size=6, linewidth=1, fontsize=3,
 #'              labelcolour="black")
 #'
 #' ## visualizing ggplot2 theme elements
@@ -114,12 +114,12 @@
 #' data(lesmis)
 #' lesmisnet <- merge(lesmis$edges, lesmis$vertices, by.x="from", by.y="label", all=TRUE)
 #' p <- ggplot(data=lesmisnet, aes(from_id=from, to_id=to))
-#' p + geom_net(layout="fruchtermanreingold")
+#' p + geom_net(layout.alg="fruchtermanreingold")
[…]

### 2016-11-21 R/stat-net.r

@@ -62,7 +62,7 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
 compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
 # cat("compute_network\n")
 #browser()
-    require(dplyr)
+    requireNamespace("dplyr")
   edges <- subset(data, to_id != "..NA..")[,c('from_id', 'to_id')]
   edges <- edges %>% group_by(from_id, to_id) %>% summarise(wt = n())
 
@@ -92,7 +92,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
     names(vert.coord)[3] <- "label"
   } else {
   #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
-  require(sna)
+  requireNamespace("sna")
   layoutFun <- paste('gplot.layout.',layout.alg,sep='')
   vert.coord <- data.frame(do.call(layoutFun, list(edgelist, layout.par = layout.par)))
 

### 2016-11-21 inst/doc/comparing-graph-drawing-speed.R

@@ -1,222 +1,0 @@
-## ----setup, echo=FALSE---------------------------------------------------
-
-knitr::opts_chunk$set(fig.width = 6,
-                      fig.height = 4,
-                      fig.align='center',
-                      dev = "png")
-
-
-## ----speed1_loadpkg, echo = TRUE, eval = FALSE, message=FALSE, warning = FALSE----
-#  ## igraph (current: v1.0.1)
-#  if (!require(igraph, quietly = TRUE)) {
-#    install.package("igraph")
-#  }
-#  library(igraph)
-#  
-#  ## network (current: v1.13.0)
-#  if (!require(network, quietly = TRUE)) {
-#    install.package("network")
-#  }
-#  library(network)
-#  
-#  ## ggnet2
-#  if (!require(GGally, quietly = TRUE)) {
-#    install.packages("GGally")
-#  }
-#  library(GGally)
-#  
-#  ## geom_net
-#  library(geomnet) # also currently requires dplyr
-#  
-#  ## ggnetwork
-#  if (!require(ggnetwork, quietly = TRUE)) {
-#    install.packages("ggnetwork")
-#  }
-#  library(ggnetwork)
-#  
-#  ## pre-loading
-#  library(sna)
-#  
-
-## ---- echo = TRUE, message = FALSE---------------------------------------
-## ggplot2
-library(ggplot2)
-
-## other packages
-library(dplyr)
-library(tidyr)
-
-## load data 
-data(protein, package = "geomnet")
-
-## ----speed1_simcode, eval=FALSE, echo=TRUE-------------------------------
-#  d = data.frame()
-#  
-#  if (!file.exists("runtimes-protein-100.csv")) {
-#  
-#    for (i in 1:100) {
-#  
-#      cat("Iteration", sprintf("%3.0f", i), "/ 100\n")
-#  
-#      n = as.matrix(protein$edges[, 1:2])
-#      n = igraph::graph_from_edgelist(n, directed = FALSE)
-#  
-#      t1 = system.time({
-#        plot(n, vertex.label = NA, layout = layout_randomly)
-#      })[1]
-#  
-#      n = network(protein$edges[, 1:2], directed = FALSE)
-#  
-#      t2 = system.time({
-#       plot(n, coord = gplot.layout.random(n, NULL))
-#      })[1]
-#  
-#      t3 = system.time({
-#        print(ggnet2(n, mode = "random"))
-#      })[1]
-#  
[…]

### 2016-11-21 inst/doc/examples-from-paper.R

@@ -1,551 +1,0 @@
-## ----load_packages, echo=TRUE--------------------------------------------
-library(dplyr)
-library(tidyr)
-library(ggplot2) # needs to be version ≥ 2.1.0
-library(scales)
-
-## ggnet2
-# if (!require(ggnet, quietly = TRUE)) {
-#   devtools::install_github("briatte/ggnet")
-# }
-# library(ggnet)
-if (!require(GGally, quietly = TRUE)) {
-  devtools::install_github("ggobi/ggally")
-}
-
-## geom_net
-if (!require(geomnet, quietly = TRUE)) {
-  devtools::install_github("sctyner/geomnet")
-}
-
-## ggnetwork
-if (!require(ggnetwork, quietly = TRUE) ||
-    packageVersion("ggnetwork") < "0.2.1") {
-  devtools::install_github("briatte/ggnetwork")
-}
-
-## pre-load
-library(network)
-library(sna)
-library(GGally)
-library(geomnet)
-library(ggnetwork)
-
-## ----madmen_geom_net, fig.width=7, fig.height=7.5------------------------
-# make data accessible 
-data(madmen, package = "geomnet")
-
-# code for geom_net
-# data step: merge edges and nodes by the "from" column
-MMnet <- merge(madmen$edges, madmen$vertices,
-               by.x = "Name1", by.y = "label", all = TRUE)
-# create plot
-set.seed(10052016)
-ggplot(data = MMnet, aes(from_id = Name1, to_id = Name2)) +
-  geom_net(aes(colour = Gender), layout = "kamadakawai", 
-           size = 2, label = TRUE, vjust = -0.6, ecolour = "grey60",
-           directed =FALSE, fontsize = 3, ealpha = 0.5) +
-  scale_colour_manual(values = c("#FF69B4", "#0099ff")) +
-  xlim(c(-0.05, 1.05)) +
-  theme_net() +
-  theme(legend.position = "bottom")
-
-## ----madmen_ggnet2, eval=FALSE, echo=TRUE, fig.cap="The code required to generate the Mad Men example using the `ggnet2` function in the `GGally` package."----
-#   library(GGally)
-#   library(network)
-#   # make the data available
-#   data(madmen, package = 'geomnet')
-#   # data step for both ggnet2 and ggnetwork
-#   # create undirected network
-#   mm.net <- network(madmen$edges[, 1:2], directed = FALSE)
-#   # create node attribute (gender)
[…]

### 2016-11-28 R/fortify-net.R
examples tag missing, visible binding for global variables

@@ -1,12 +1,14 @@
 #' Function for converting a network object into the correct format for use with geomnet
-#' @param model. object of class "network"
-#' @param data NULL
+#'
+#' @param model object of class "network"
+#' @param data NULL - not used in this function
 #' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#' @param ...  not used in this function
 #' @examples
 #' # class network (sna, network, statnet packages)
 #'
 #' library(network)
-#' data(emon)
+#' data(emon, package = "network")
 #' fortify(emon$Cheyenne)
 #'
 #' @import network
@@ -46,13 +48,14 @@ fortify.network <- function(model, data = NULL, group = NULL, ...){
   return(dat)
 }
 #' Function for converting an igraph object into the correct format for use with geomnet
-#' @param model. A an "igraph" object
-#' @param data. NULL
-#' @param group. character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#'
+#' @param model A an "igraph" object
+#' @param data NULL - not used in this function
+#' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#' @param ...  not used in this function
 #' @examples
 #' # class igraph (igraph, igraphdata packages)
 #' library(igraph)
-#' library(igraphdata)
 #' data("USairports", package = "igraphdata")
 #' head(fortify(USairports))
 #'
@@ -76,9 +79,11 @@ fortify.igraph <- function(model, data = NULL, group = NULL, ...){
   return(dat)
 }
 #' Function for converting a network edge list in data frame form into the correct format for use with geomnet
-#' @param model. A network edgelist of class "data.frame" object. The first column should contain the "from" node column.
-#' @param data. Data frame containing network node list and other node information. First column should contain node ids.
-#' @param group. character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#'
+#' @param model A network edgelist of class "data.frame" object. The first column should contain the "from" node column.
+#' @param data Data frame containing network node list and other node information. First column should contain node ids.
[…]

### 2016-11-28 R/network-tools.r

@@ -7,6 +7,8 @@
 #' @param data data set
 #' @param from_id name of the variable of edge tails
 #' @param to_id name of the variable of edge heads
+#' @importFrom network as.network.matrix
+#' @importFrom network as.matrix.network
 #' @export
 #' @examples
 #' data(blood)

### 2016-11-28 R/data.r
typo in example

@@ -15,7 +15,7 @@
 #'   \item n: Number of trips between the two stations
 #'   \item minlength: Duration of shortest trip between the two stations (in seconds). Only those stations are included, if the shortest trip between them lasted not more than 15 minutes.
 #' }
-#' 
+#'
 #' \item the vertices data set consists of five variables with information on 21 stations:
 #' \itemize{
 #'   \item id: Station ID number
@@ -121,7 +121,7 @@
 #'
 #' # label independent schools
 #' ftnet$schools <- ifelse(ftnet$value == "Independents", ftnet$from, "")
-#' 
+#'
 #' library(geomnet)
 #' library(dplyr)
 #' # create data plot
@@ -139,7 +139,7 @@
 #'   ) +
 #'   theme_net() +
 #'   theme(legend.position = "bottom") +
-#'   scale_colour_brewer("Conference", palette = "Paired") 
+#'   scale_colour_brewer("Conference", palette = "Paired")
 "football"
 
 #' Coappearance network of characters in Les Miserables (undirected)
@@ -171,12 +171,12 @@
 #' # create plot
 #' library(geomnet)
 #' library(dplyr)
-#' 
+#'
 #' ggplot(data = lesmisnet, aes(from_id = from, to_id = to,
 #'                              linewidth = degree / 5 + 0.1 )) +
 #'   geom_net(aes(size = degree, alpha = degree),
 #'            colour = "grey30", ecolour = "grey60",
-#'            layout.alg = "fruchtermanreingold", label = TRUE, vjust = -0.75) +
+#'            layout.alg = "fruchtermanreingold", labelon = TRUE, vjust = -0.75) +
 #'   scale_alpha(range = c(0.3, 1)) +
 #'   theme_net()
 "lesmis"
@@ -207,10 +207,10 @@
 #'
 #' A dataset containing information to draw a map of Rockville, MD, and vicinity using \code{ggmap}.
 #' This information was pulled from Google Maps using the \pkg{ggmap} package.
-#' 
+#'
 #' @references D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2.
 #' The R Journal, 5(1), 144-161. \url{http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf}
-#' @format An object of class "ggmap" and "raster" containing a map of the Rockville, MD, area. 
[…]

### 2016-11-28 R/fortify-net.R

@@ -123,7 +123,7 @@ fortify.data.frame <- function(model, data, group = NULL, ...){
 #' adjmat <- network::as.matrix.network.adjacency(emon$MtSi)
 #' str(adjmat)
 #' fortify(adjmat)
-fortify.matrix <- function(model, data = NULL, ...){
+fortify.matrix <- function(model, data = NULL, group = NULL, ...){
   net <- model
   if (dim(net)[1] != dim(net)[2]){
     stop("Error: Please supply a square adjacency matrix.")
@@ -139,6 +139,7 @@ fortify.matrix <- function(model, data = NULL, ...){
   to <- NULL
   from <- NULL
   value <- NULL
+  if (!is.null(group)) cat("group argument not implemented yet\n")
 
   net <- tidyr::gather(net, to, value, -from)
 

### 2016-11-28 R/stat-net.r

@@ -5,6 +5,9 @@
 #' @rdname geom_net
 #' @importFrom sna as.edgelist.sna
 #' @importFrom network as.matrix.network.edgelist
+#' @importFrom dplyr group_by
+#' @importFrom dplyr mutate
+#' @importFrom dplyr summarise
 #' @export
 StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
   required_aes = c("from_id", "to_id"),
@@ -64,7 +67,8 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
 #browser()
 
   edges <- subset(data, to_id != "..NA..")[,c('from_id', 'to_id')]
-  edges <- edges %>% group_by(from_id, to_id) %>% summarise(wt = n())
+  edges <- dplyr::group_by(edges, from_id, to_id)
+  edges <- dplyr::summarise(edges, wt = n())
 
     # there should not be any missing values at this point, but just make sure
   if (any(is.na(edges$from_id))) message(sprintf("%d missing values excluded\n", sum(is.na(edges$from_id))))
@@ -121,7 +125,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
 
     edges <- rbind(edges, fromonly[, names(edges)])
   }
-  edges <- edges %>% group_by(from, to) %>% mutate(weight = n())
+  edges <- mutate(group_by(edges, from, to), weight = n())
   unique(edges)
 },
 
@@ -134,8 +138,9 @@ compute_panel = function(self, data, scales, na.rm = FALSE,
       data <- self$compute_network(data, layout.alg =layout.alg, layout.par=layout.par)
 
     #    data <- plyr::ddply(data, "group", plyr::mutate, .samegroup = to %in% unique(from))
-    if (any(data$group) != -1)
-      data <- data %>% group_by(group) %>% mutate(.samegroup = to %in% unique(from))
+    if (any(data$group) != -1) {
+      data <- mutate(group_by(data, group), .samegroup = to %in% unique(from))
+    }
 
    # #browser()
     data.frame(data)

### 2016-11-28 R/stat-net.r
comment out unnecessary call

@@ -98,7 +98,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
   #print("it would be nice at this point to check, whether layout is one of the supported functions, and if not,
 # browser()
   layoutFun <- getFromNamespace(paste('gplot.layout.',layout.alg,sep=''), asNamespace("sna"))
-  requireNamespace("sna")
+#  requireNamespace("sna")
 #  layoutFun <- paste('gplot.layout.',layout.alg,sep='')
 #  vert.coord <- data.frame(do.call(layoutFun, list(edgelist, layout.par = layout.par)))
   vert.coord <- data.frame(layoutFun(edgelist, layout.par = layout.par))

### 2016-11-28 R/fortify-net.R
group argument gone from fortify.matrix

@@ -113,7 +113,6 @@ fortify.data.frame <- function(model, data, group = NULL, ...){
 #'
 #' @param model An adjacency matrix of class "matrix".
 #' @param data NULL - not used in this function
-#' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
 #' @param ...  not used in this function
 #' @export
 #' @importFrom tidyr gather
@@ -123,7 +122,7 @@ fortify.data.frame <- function(model, data, group = NULL, ...){
 #' adjmat <- network::as.matrix.network.adjacency(emon$MtSi)
 #' str(adjmat)
 #' fortify(adjmat)
-fortify.matrix <- function(model, data = NULL, group = NULL, ...){
+fortify.matrix <- function(model, data = NULL,  ...){
   net <- model
   if (dim(net)[1] != dim(net)[2]){
     stop("Error: Please supply a square adjacency matrix.")
@@ -139,7 +138,6 @@ fortify.matrix <- function(model, data = NULL, group = NULL, ...){
   to <- NULL
   from <- NULL
   value <- NULL
-  if (!is.null(group)) cat("group argument not implemented yet\n")
 
   net <- tidyr::gather(net, to, value, -from)
 

### 2016-11-29 R/fortify-net.R
create new classes for use with fortify methods.

@@ -1,6 +1,6 @@
 #' Function for converting a network object into the correct format for use with geomnet
 #'
-#' @param model object of class "network"
+#' @param model object of class \code{"network"}
 #' @param data NULL - not used in this function
 #' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
 #' @param ...  not used in this function
@@ -49,7 +49,7 @@ fortify.network <- function(model, data = NULL, group = NULL, ...){
 }
 #' Function for converting an igraph object into the correct format for use with geomnet
 #'
-#' @param model A an "igraph" object
+#' @param model A network object of class \code{"igraph"}.
 #' @param data NULL - not used in this function
 #' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
 #' @param ...  not used in this function
@@ -80,21 +80,24 @@ fortify.igraph <- function(model, data = NULL, group = NULL, ...){
 }
 #' Function for converting a network edge list in data frame form into the correct format for use with geomnet
 #'
-#' @param model A network edgelist of class "data.frame" object. The first column should contain the "from" node column.
+#' @param model A network edgelist of class \code{"edgedf"}. See \code{\link{as.edgedf}}. Can contain edge variables as well. 
 #' @param data Data frame containing network node list and other node information. First column should contain node ids.
 #' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
 #' @param ...  not used in this function
 #' @examples
-#' # class data.frame and ndata
+#'
 #' data(blood)
-#' fortify(blood$edges, blood$vertices)
+#' fortify(as.edgedf(blood$edges), blood$vertices)
 #' @export
-fortify.data.frame <- function(model, data, group = NULL, ...){
+fortify.edgedf <- function(model, data, group = NULL, ...){
   edge.data <- model
   if (is.null(data)){
     stop("Error: Must provide the node data to the data argument.")
   }
   node.data <- data
+  message(paste("Joining edge and node information by", 
+                names(edge.data)[1], "and", 
+                names(node.data)[1], "respectively."))
   if(!is.null(group)){
     nodes <- unique(node.data[,1])
     groups <- unique(edge.data[,group])
@@ -111,15 +114,15 @@ fortify.data.frame <- function(model, data, group = NULL, ...){
 }
[…]

### 2016-11-29 R/utils.R

@@ -1,0 +1,35 @@
+#' Create and assign a new class for use of \code{geomnet}'s \code{\link{fortify.adjmat}} function.
+#' @param dat. A square adjacency matrix.
+#' @export
+as.adjmat <- function(dat){
+  if (class(dat) != "matrix"){
+    stop("as.adjmat requires a matrix object")
+  }
+  if (nrow(dat) != ncol(dat)){
+    stop("dat must be a square matrix")
+  }
+  if (sum(rownames(dat)!=colnames(dat)) > 0){
+    stop("row and column names of dat must match")
+  }
+  class(dat) <- c("adjmat", "matrix")
+  return(dat)
+}
+#' Create and assign a new class for use of \code{geomnet}'s \code{\link{fortify.edgedf}} function.
+#' @param dat. A network edgelist of class \code{"data.frame"} for use with \code{\link{fortify.edgedf}}.
+#' @export
+as.edgedf <- function(dat){
+  if (class(dat) != "data.frame"){
+    stop("as.edgedf requires a data.frame object")
+  }
+  if (ncol(dat) < 2){
+    stop("dat should have at least 2 columns (from node and to node)")
+  }
+  from_id <- names(dat)[1]
+  to_id <- names(dat)[2]
+  message(paste("Using", from_id, "as the from node column and",
+              to_id, "as the to node column. 
+If this is not correct, rewrite dat so that the first 2 columns are from and to node, respectively."))
+  names(dat)[1:2] <- c("from_id", "to_id")
+  class(dat) <- c("edgedf", "data.frame")
+  return(dat)
+}

### 2016-11-29 R/fortify-net.R
start tidying documentation

@@ -125,7 +125,7 @@ fortify.edgedf <- function(model, data, group = NULL, ...){
 #' adjmat <- as.adjmat(network::as.matrix.network.adjacency(emon$MtSi))
 #' str(adjmat)
 #' fortify(adjmat)
-fortify.matrix <- function(model, data = NULL,  ...){
+fortify.adjmat <- function(model, data = NULL,  ...){
   net <- model
   if (dim(net)[1] != dim(net)[2]){
     stop("Error: Please supply a square adjacency matrix.")

### 2016-11-29 R/geom-net.r

@@ -1,26 +1,23 @@
-#' Geom for network visualization within the ggplot2 framework
+#' Networks
 #'
-
-#' @inheritParams ggplot2::stat_identity
-#' @param stat character string of the network stat corresponding to geom_net.
-#' @param alpha numeric value of alpha blending of vertices.
-#' @param ealpha numeric value of alpha blending of edges.
-#' @param na.rm If \code{FALSE} (the default), removes missing values with
-#'    a warning. If \code{TRUE} silently removes missing values.
-#' @param layout.alg character value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?gplot.layout} in the package sna for other choices.
-#' @param layout.par list of parameters detailing algorithmic specs. Default parameters are used initially. See \code{?gplot.layout} in the package sna for other choices.
-#' @param fiteach logical value. Should the network be fit in each of the panels separately, or is there going to be one fit for all?
-#' @param labelon logical value. Include labels for (all) nodes. labelcolour specifies colour of labels, if they should not be the same as the nodes. labels are taken from the from_id variable, unless a label variable is given.
-#' @param labelcolour character of colour for the labels.
-#' @param labelgeom character. Which ggplot2 \code{geom} to use to draw the labels. Either "text" or "label".
-#' @param repel logical value. If \code{TRUE}, uses the ggrepel package geoms to draw the node labels instead of the ggplot2 geoms.
-#' @param ecolour colour for edges.
-#' @param directed logical value. Should an arrow be drawn from 'from' to 'to' node?
-#' @param selfloops logical value. Should self-references be shown (by drawing a circle adjacent to the corresponding node)? defaults to FALSE.
+#' The net geom is used visualize networks within the \pkg{ggplot2} framework. \code{geom_net} combines the many parts of a network visualization
+#' into a single layer in \pkg{ggplot2}. It combines various other geoms, including but not limited to, \code{\link[ggplot2]{geom_point}}, \code{\link[ggplot2]{geom_segment}}, and \code{\link[ggplot2]{geom_text}}.  
+#' @inheritParams ggplot2::geom_point
+#' @param layout.alg character. Value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{\link[sna]{gplot.layout}} documentation for more choices.
+#' @param layout.par list. Parameters detailing algorithmic specs. Default parameters from \pkg{sna} are used initially. See \code{\link[sna]{gplot.layout}} documentation for all options corresponding to all layouts.
[…]

### 2016-11-29 R/fortify-net.R
clean up geomnet examples

@@ -2,7 +2,7 @@
 #'
 #' @param model object of class \code{"network"}
 #' @param data NULL - not used in this function
-#' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#' @param group character. Used for facetting. If you wish to facet on a network variable provide the name of that variable here.
 #' @param ...  not used in this function
 #' @examples
 #' # class network (sna, network, statnet packages)
@@ -51,7 +51,7 @@ fortify.network <- function(model, data = NULL, group = NULL, ...){
 #'
 #' @param model A network object of class \code{"igraph"}.
 #' @param data NULL - not used in this function
-#' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#' @param group character. Used for facetting. If you wish to facet on a network variable provide the name of that variable here.
 #' @param ...  not used in this function
 #' @examples
 #' # class igraph (igraph, igraphdata packages)
@@ -82,12 +82,13 @@ fortify.igraph <- function(model, data = NULL, group = NULL, ...){
 #'
 #' @param model A network edgelist of class \code{"edgedf"}. See \code{\link{as.edgedf}}. Can contain edge variables as well. 
 #' @param data Data frame containing network node list and other node information. First column should contain node ids.
-#' @param group character. Used for facetting. If you wish to facet on an edge variable provide the name of that variable here.
+#' @param group character. Used for facetting. If you wish to facet on network variable provide the name of that variable here.
 #' @param ...  not used in this function
 #' @examples
 #'
 #' data(blood)
 #' fortify(as.edgedf(blood$edges), blood$vertices)
+#' fortify(as.edgedf(blood$edges), blood$vertices, group = "Ethnicity")
 #' @export
 fortify.edgedf <- function(model, data, group = NULL, ...){
   edge.data <- model
@@ -100,13 +101,21 @@ fortify.edgedf <- function(model, data, group = NULL, ...){
                 names(node.data)[1], "respectively."))
   if(!is.null(group)){
     nodes <- unique(node.data[,1])
-    groups <- unique(edge.data[,group])
+    if(group %in% names(edge.data)){
+      groups <- unique(edge.data[,group])
+    } else groups <- unique(node.data[,group])
     allnodes <- expand.grid(nodes, groups, stringsAsFactors = F)
[…]

### 2016-11-29 R/geom-net.r

@@ -1,10 +1,15 @@
 #' Networks
 #'
 #' The net geom is used visualize networks within the \pkg{ggplot2} framework. \code{geom_net} combines the many parts of a network visualization
-#' into a single layer in \pkg{ggplot2}. It combines various other geoms, including but not limited to, \code{\link[ggplot2]{geom_point}}, \code{\link[ggplot2]{geom_segment}}, and \code{\link[ggplot2]{geom_text}}.  
+#' into a single layer in \pkg{ggplot2}. It makes use of various other geoms, including but not limited to, \code{\link[ggplot2]{geom_point}}, \code{\link[ggplot2]{geom_segment}}, and \code{\link[ggplot2]{geom_text}}.  
+#'
+#' @section Aesthetics:
+#' \aesthetics{geom}{net}
+#' 
 #' @inheritParams ggplot2::geom_point
-#' @param layout.alg character. Value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{\link[sna]{gplot.layout}} documentation for more choices.
-#' @param layout.par list. Parameters detailing algorithmic specs. Default parameters from \pkg{sna} are used initially. See \code{\link[sna]{gplot.layout}} documentation for all options corresponding to all layouts.
+#' @param layout.alg character. Value specifying the layout algorithm to use. Defaults to "kamadakawai". See \code{?sna::}\code{\link[sna]{gplot.layout}} documentation for more choices.
+#' @param layout.par list. Parameters detailing algorithmic specs. Default parameters from \pkg{sna} are used initially. See  \code{?sna::}\code{\link[sna]{gplot.layout}} documentation for all options corresponding to all layouts.
+#' @param directed logical value. Should an arrow be drawn pointing to the \code{to_id} node? Default is \code{FALSE}.
 #' @param ealpha numeric. Value from 0-1 of alpha blending of edges.
 #' @param fiteach logical. Should the network be fit in each of the panels separately, or is there going to be one fit for all?
 #' @param labelon logical. Include labels for all nodes.  Labels are taken from the \code{from_id} variable, unless a \code{label} aesthetic is provided.
@@ -12,9 +17,8 @@
[…]

### 2016-11-30 R/stat-net.r
change protein data to characters

@@ -13,7 +13,7 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
   required_aes = c("from_id", "to_id"),
   non_missing_aes = "weight",
   setup_params = function(data, params) {
-  #  #browser()
+  #  browser()
 #    print(str(params))
 
     params

### 2016-12-05 R/data.r
use canonical urls

@@ -209,7 +209,7 @@
 #' This information was pulled from Google Maps using the \pkg{ggmap} package.
 #'
 #' @references D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2.
-#' The R Journal, 5(1), 144-161. \url{http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf}
+#' The R Journal, 5(1), 144-161. \url{https://journal.r-project.org/archive/2013-1/kahle-wickham.pdf}
 #' @format An object of class "ggmap" and "raster" containing a map of the Rockville, MD, area.
 #' @examples
 #' \dontrun{
@@ -227,7 +227,7 @@
 #' The variables are as follows:
 #'
 #' @references Winston Chang (2012). gcookbook: Data for "R Graphics Cookbook". R package
-#' version 1.0. \url{http://CRAN.R-project.org/package=gcookbook}
+#' version 1.0. \url{https://CRAN.R-project.org/package=gcookbook}
 #' @format A list of two data frames:
 #' \itemize{
 #' \item the edges data set consists of two variables of length 87:

### 2016-12-05 README.Rmd
add examples in readme file

---
title: "The `geomnet` package"
author: Sam Tyner, Heike Hofmann
date: "`r Sys.Date()`"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
[![CRAN Status](http://www.r-pkg.org/badges/version/geomnet)](https://cran.r-project.org/package=geomnet)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/geomnet)](http://www.r-pkg.org/pkg/geomnet)
[![Travis-CI Build Status](https://travis-ci.org/sctyner/geomnet.svg?branch=master)](https://travis-ci.org/sctyner/geomnet)

`geomnet` is a package built on top of the most recent major `ggplot2` release. It provides a `ggplot2` `geom` called `geom_net` to visualize graphs and networks. It also include the function `stat_net` to calculate network layouts with the `sna` package. Finally, the function `geom_circle` is included to draw circles using `ggplot2`.

You can install `geomnet` directly from CRAN
`install.packages('geomnet')`
or from Github
`devtools::install_github("sctyner/geomnet")`
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Examples

## `ggplot2` Theme Elements

This example shows the theme inheritance properties of the theme elements of `ggplot2`. Note: this example has not been updated since the release of `ggplot2 2.2.0` and as such the content may have changed.

```{r theme, fig.align='center', message=FALSE, warning=FALSE}
library(dplyr)
library(geomnet)
data(theme_elements)
# data step
TEnet <- fortify(as.edgedf(theme_elements$edges[,c("parent", "child")]), theme_elements$vertices)
# create a degree variable for use later
TEnet <- TEnet %>%
  group_by(from_id) %>%
  mutate(degree = sqrt(10 * n() + 1))
# plot
ggplot(data = TEnet,
       aes(from_id = from_id, to_id = to_id)) +
  geom_net(layout.alg = "fruchtermanreingold",
    aes(fontsize = degree), directed = TRUE,
    labelon = TRUE, size = 1, labelcolour = 'black',
    ecolour = "grey70", arrowsize = 0.5,
    linewidth = 0.5, repel = TRUE) +
  theme_net() +
  xlim(c(-0.05, 1.05))
```

## Blood Donation Diagram

In this example, we reimagine the [traditional blood donation diagram](http://www.redcrossblood.org/learn-about-blood/blood-types) as a directed network. Arrows point to the blood type that receives. This example provides two data frames to `geom_net()`.

```{r blood, fig.align='center'}
library(geomnet)
data(blood)
ggplot(data = blood$edges, aes(from_id = from, to_id = to)) +
  geom_net(colour = "darkred", layout.alg = "circle", labelon = TRUE, 
           size = 15, directed = TRUE, vjust = 0.5, labelcolour = "grey80",
           arrowsize = 1.5, linewidth = 0.5, arrowgap = 0.05,
           selfloops = TRUE, ecolour = "grey40") + 
  theme_net() 
```

## Harry Potter Peer Support Network

In this fun example from [this website](http://www.stats.ox.ac.uk/~snijders/siena/siena.html), there is a tie between two students if one provides emotional support to the other at some point in the book. It is a directed network, so in the visualization, the arrow points to the student receiving support.

```{r HP, echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
temp <- tempfile()
download.file("http://www.stats.ox.ac.uk/~snijders/siena/bossaert_meidert_harrypotter.zip",temp)
hp1 <- read.table(unz(temp, "hpbook1.txt"))
hp1 <- as.matrix(hp1)
diag(hp1) <- 0
hp1 <- data.frame(hp1)
hp1$id <- 1:64
hp1 <- hp1 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 1) %>% 
               dplyr::rename(from = id, to = var)
hp2 <- read.table(unz(temp, "hpbook2.txt"))
hp2 <- as.matrix(hp2)
diag(hp2) <- 0
hp2 <- data.frame(hp2)
hp2$id <- 1:64
hp2 <- hp2 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 2) %>% 
               dplyr::rename(from = id, to = var)
hp3 <- read.table(unz(temp, "hpbook3.txt"))
hp3 <- as.matrix(hp3)
diag(hp3) <- 0
hp3 <- data.frame(hp3)
hp3$id <- 1:64
hp3 <- hp3 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 3) %>% 
               dplyr::rename(from = id, to = var)
hp4 <- read.table(unz(temp, "hpbook4.txt"))
hp4 <- as.matrix(hp4)
diag(hp4) <- 0
hp4 <- data.frame(hp4)
hp4$id <- 1:64
hp4 <- hp4 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 4) %>% 
               dplyr::rename(from = id, to = var)
hp5 <- read.table(unz(temp, "hpbook5.txt"))
hp5 <- as.matrix(hp5)
diag(hp5) <- 0
hp5 <- data.frame(hp5)
hp5$id <- 1:64
hp5 <- hp5 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 5) %>% 
               dplyr::rename(from = id, to = var)
hp6 <- read.table(unz(temp, "hpbook6.txt"))
hp6 <- as.matrix(hp6)
diag(hp6) <- 0
hp6 <- data.frame(hp6)
hp6$id <- 1:64
hp6 <- hp6 %>% tidyr::gather(var, value, -id) %>%
               dplyr::filter(value == 1) %>% 
               dplyr::select(-value) %>% 
               dplyr::mutate(var = readr::parse_number(var), book = 6) %>% 
               dplyr::rename(from = id, to = var)
hp <- rbind(hp1, hp2, hp3, hp4, hp5, hp6)
hp.attributes <- read.table(unz(temp, "hpattributes.txt"), header=TRUE, stringsAsFactors = F)
hp.names <- read.table(unz(temp, "hpnames.txt"), header=TRUE, stringsAsFactors = F)
hp.names[,1] <- as.integer(hp.names[,1])
hp.1 <- merge(hp, hp.names, by.x = "from", by.y = "id")
names(hp.1)[4] <- "name1"
hp.2 <- merge(hp.1, hp.names, by.x = "to", by.y = "id")
names(hp.2)[5] <- "name2"
hp.edges <- hp.2[, c("name1", "name2", "book")] %>% dplyr::arrange(book, name1, name2)
hp.chars <- merge(hp.names, hp.attributes)
hp.chars$house <- factor(hp.chars$house, labels = c("Gryffindor", "Hufflepuff", "Ravenclaw", "Slytherin"))
hp.chars$gender <- factor(hp.chars$gender, labels = c("M", "F"))
unlink(temp)
hp.chars <- hp.chars[,-1]
```

```{r HPplot, fig.width=10, fig.align='center'}
library(geomnet)
head(hp.edges)
head(hp.chars)
hp.all <- fortify(as.edgedf(hp.edges), hp.chars, group = "book")
ggplot(data=hp.all, aes(from_id = from, to_id = to_id)) + 
  geom_net(fiteach=T, directed = T, size = 2, linewidth = .5, 
           ealpha = .5, labelon = T, fontsize = 2, repel = T, 
           labelcolour = "black", arrowsize = .5,
           aes(colour = house, group = house, shape = gender)) + 
  scale_colour_manual(values = c("#941B08","#F1F31C", 
                                 "#071A80", "#154C07")) + 
  facet_wrap(~book, labeller = "label_both") + 
  theme_net() + theme(panel.background = element_rect(colour = 'black'))
```

### 2016-12-06 README.Rmd
fix HP plot to ignore singletons

@@
 
-```{r HPplot, fig.width=10, fig.align='center'}
+```{r HPplot, fig.width=10, fig.height=8, fig.align='center'}
 library(geomnet)
@@
 hp.all <- fortify(as.edgedf(hp.edges), hp.chars, group = "book")
-ggplot(data=hp.all, aes(from_id = from, to_id = to_id)) + 
+# only plot the characters with any connections in a given book. 
+hp.sub <- NULL
+for (i in 1:6){
+  dat <- filter(hp.all, book == i)
+  singletons <- dat$from[is.na(dat$to_id)]
+  true_singletons <- singletons[!singletons %in% unique(dat$to_id)[-1]]
+  dat2 <- dat[!dat$from %in% true_singletons,]
+  hp.sub <- rbind(hp.sub, dat2)
+}
+ggplot(data=hp.sub, aes(from_id = from, to_id = to_id)) + 
   geom_net(fiteach=T, directed = T, size = 2, linewidth = .5, 

### 2016-12-06 R/stat-net.r

@@ -133,7 +133,7 @@ compute_panel = function(self, data, scales, na.rm = FALSE,
                          layout.alg="kamadakawai", layout.par=list(),
                          fiteach=FALSE, vertices=NULL) {
 #  cat("compute_panel in stat_net\n")
-#  #browser()
+#  browser()
 #    if (fiteach)
       data <- self$compute_network(data, layout.alg =layout.alg, layout.par=layout.par)
 

### 2016-12-06 R/geom2plotly.R
plotly hack to have midpoint edge labels on hover

@@ -15,15 +15,23 @@ to_basic.GeomNet <- function(data, prestats_data, layout, params, p, ...) {
   edge_data <- unique(data[, edge_names])
   edge_data <- edge_data[which(!edge_data$.selfie),]
   edge_data$size <- edge_data$linewidth
-  if (params$directed){
-    edge_data$hovertext <- paste("from", edge_data$from, "to", edge_data$to, sep = " ")
-  } else edge_data$hovertext <- paste("(", edge_data$from, ",", edge_data$to, ")", sep = "")
+  edge_data$hovertext <- NULL
+  edge_labels <- edge_data
   if (is.null(params$ealpha)){
     edge_data$alpha <- 1
   } else edge_data$alpha <- params$ealpha
   edge_data$colour <- ifelse(is.null(params$ecolour), "grey40", params$ecolour)
   edge_data <- getFromNamespace("to_basic.GeomSegment", asNamespace("plotly"))(edge_data)
   edge_data <- getFromNamespace("prefix_class", asNamespace("plotly"))(edge_data, "GeomNet")
-  data <- list(edge_data, node_data)
+  edge_labels$x <- (edge_labels$x + edge_labels$xend)/2
+  edge_labels$y <- (edge_labels$y + edge_labels$yend)/2
+  edge_labels$colour <- "white"
+  edge_labels$shape <- ""
+  edge_labels$size <- 0
+  if (params$directed){
+    edge_labels$hovertext <- paste("from", edge_labels$from, "to", edge_labels$to, sep = " ")
+  } else edge_labels$hovertext <- paste("(", edge_labels$from, ",", edge_labels$to, ")", sep = "")
+  edge_labels <- getFromNamespace("prefix_class", asNamespace("plotly"))(edge_labels, "GeomPoint")
+  data <- list(edge_labels, edge_data, node_data)
   data
 }

### 2016-12-06 README.Rmd
add "singletons" parameter

--- 2016-12-05 README.Rmd
@@
 
-```{r HPplot, fig.width=10, fig.align='center'}
+```{r HPplot, fig.width=10, fig.height=8, fig.align='center', warning = FALSE}
 library(geomnet)
@@
 hp.all <- fortify(as.edgedf(hp.edges), hp.chars, group = "book")
+# only plot the characters with any connections in a given book. 
 ggplot(data=hp.all, aes(from_id = from, to_id = to_id)) + 
-  geom_net(fiteach=T, directed = T, size = 2, linewidth = .5, 
-           ealpha = .5, labelon = T, fontsize = 2, repel = T, 
-           labelcolour = "black", arrowsize = .5,
+  geom_net(fiteach=T, directed = T, size = 3, linewidth = .5, 
+           ealpha = .5, labelon = T, fontsize = 3, repel = T, 
+           labelcolour = "black", arrowsize = .5, singletons = FALSE,
            aes(colour = house, group = house, shape = gender)) + 

### 2016-12-06 R/geom-net.r

@@ -29,6 +29,7 @@
 #' @param alpha numeric. Value from 0-1 of alpha blending of nodes.
 #' @param ealpha numeric. Value from 0-1 of alpha blending of edges.
 #' @param fiteach logical. Should the network be fit in each of the panels separately, or is there going to be one fit for all?
+#' @param singletons logical. Should singletons (nodes with no incoming or outgoing ties) be plotted? Default is \code{TRUE}.
 #' @param labelon logical. Include labels for all nodes.  Labels are taken from the \code{from_id} variable, unless a \code{label} aesthetic is provided.
 #' @param labelcolour character. Colour for the labels. If this argument is not specified, labels have the same colour as the nodes.
 #' @param labelgeom character. Which \pkg{ggplot2} \code{geom} should be used to draw the labels? Either \code{"text"} or \code{"label"}. Default is \code{"text"}
@@ -143,7 +144,7 @@
 
 geom_net <- function (
   mapping = NULL, data = NULL, stat = "net", position = "identity", show.legend = NA, na.rm = FALSE, inherit.aes = TRUE,
-  layout.alg="kamadakawai", layout.par=list(), directed = FALSE, fiteach=FALSE,  selfloops = FALSE,
+  layout.alg="kamadakawai", layout.par=list(), directed = FALSE, fiteach=FALSE,  selfloops = FALSE, singletons = TRUE,
   alpha = 0.25, ecolour=NULL, ealpha=NULL, arrow=NULL, arrowgap=0.01, arrowsize=1,
   labelon=FALSE, labelcolour=NULL, labelgeom = 'text', repel = FALSE,
   vertices=NULL, ...) {
@@ -152,8 +153,8 @@ geom_net <- function (
     geom = GeomNet, mapping = mapping,  data = data, stat = stat,
     position = position, show.legend = show.legend, inherit.aes = inherit.aes,
     params = list(na.rm = na.rm, layout.alg=layout.alg, layout.par=layout.par, fiteach=fiteach, labelon=labelon, labelgeom=labelgeom,
-                  ecolour = ecolour, ealpha=ealpha, arrow=arrow, arrowgap=arrowgap, directed=directed, repel = repel,
-                  arrowsize=arrowsize,
+                  ecolour = ecolour, ealpha=ealpha, arrow=arrow, arrowgap=arrowgap, directed=directed, repel=repel,
[…]

### 2016-12-06 R/stat-net.r

@@ -62,11 +62,16 @@ StatNet <- ggplot2::ggproto("StatNet", ggplot2::Stat,
     data
   },
 
-compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
+compute_network = function(data, layout.alg="kamadakawai", layout.par=list(), singletons = TRUE) {
 # cat("compute_network\n")
-#browser()
+# browser()
 
   edges <- subset(data, to_id != "..NA..")[,c('from_id', 'to_id')]
+  if (!singletons){
+    all_singletons <- edges$from_id[edges$from_id == edges$to_id]
+    true_singletons <- all_singletons[!all_singletons %in% edges$to_id[edges$from_id != edges$to_id]]
+    edges <- edges[!edges$from_id %in% true_singletons,]
+  }
   edges <- dplyr::group_by(edges, from_id, to_id)
   edges <- dplyr::summarise(edges, wt = n())
 
@@ -130,12 +135,12 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list()) {
 },
 
 compute_panel = function(self, data, scales, na.rm = FALSE,
-                         layout.alg="kamadakawai", layout.par=list(),
+                         layout.alg="kamadakawai", layout.par=list(), singletons=TRUE,
                          fiteach=FALSE, vertices=NULL) {
 #  cat("compute_panel in stat_net\n")
 #  browser()
 #    if (fiteach)
-      data <- self$compute_network(data, layout.alg =layout.alg, layout.par=layout.par)
+      data <- self$compute_network(data, layout.alg =layout.alg, layout.par=layout.par, singletons=singletons)
 
     #    data <- plyr::ddply(data, "group", plyr::mutate, .samegroup = to %in% unique(from))
     if (any(data$group) != -1) {
@@ -163,7 +168,7 @@ compute_layer = function(self, data, params, layout, na.rm = FALSE, layout.alg,
       self$compute_panel(data = data, scales = scales,
                          na.rm=params$na.rm, layout.alg=params$layout.alg,
                          layout.par=params$layout.par, fiteach=params$fiteach,
-                         vertices=params$vertices)
+                         vertices=params$vertices, singletons=params$singletons)
     })
   }
   else {
@@ -174,7 +179,7 @@ compute_layer = function(self, data, params, layout, na.rm = FALSE, layout.alg,
     self$compute_panel(data = data, scales = scales,
                        na.rm=params$na.rm, layout.alg=params$layout.alg,
[…]

### 2016-12-06 README.Rmd
add plotly code example

@@
 ```
+
+## Interactivity with `plotly`!
+
+Now including interactivity with `ggplotly()`!
+
+```{r football, eval = FALSE}
+library(geomnet)
+# devtools::install_github("ropensci/plotly")
+library(plotly)
+data("football")
+# data step: merge vertices and edges
+ftnet <- fortify(as.edgedf(football$edges), football$vertices)
+
+# create data plot
+gg <- ggplot(data = ftnet,
+       aes(from_id = from_id, to_id = to_id)) +
+  geom_net(layout.alg = 'fruchtermanreingold',
+    aes(colour = value, group = value,
+        linetype = factor(same.conf != 1)),
+    linewidth = 0.5,
+    size = 5, vjust = -0.75, alpha = 0.3) +
+  theme_net() +
+  theme(legend.position = "bottom") +
+  scale_colour_brewer("Conference", palette = "Paired")  +
+  guides(linetype = FALSE)
+ggplotly(gg)
+```

### 2017-04-05 R/geom2plotly.R
fix bug with plotly to.basic.

@@ -2,17 +2,22 @@
 #' @importFrom plotly to_basic
 #' @export
 to_basic.GeomNet <- function(data, prestats_data, layout, params, p, ...) {
+  #browser()
+  if (params$directed){
+    message("Note: the plotly package does not yet support arrows for segments.\nSee https://github.com/ropensci/plotly/issues/469 for updates")
+  }
   # get x,y and xend,yend coordinates, among other things
   dat2 <- ggplot2::ggplot_build(p)$data[[1]]
   data <- getFromNamespace("left_join", asNamespace("dplyr"))(data, dat2)
   node_names <- names(data)[!(names(data) %in% c("to","xend", "yend", ".selfie", "weight", ".samegroup", "ymax", "xmax", "width", "linetype", "fontsize", "arrowsize", "stroke"))]
-  edge_names <- c("from","to", "x", "y","xend", "yend", ".selfie", "weight", ".samegroup", "ymax", "xmax", "width", "linetype", "linewidth", "fontsize", "arrowsize", "stroke")
+  # edge_names <- c("from","to", "x", "y","xend", "yend", ".selfie", "weight", ".samegroup", "ymax", "xmax", "width", "linetype", "linewidth", "fontsize", "arrowsize", "stroke", "PANEL")
   node_data <- unique(data[, node_names])
   node_data <- getFromNamespace("prefix_class", asNamespace("plotly"))(node_data, "GeomPoint")
-  if (is.null(node_data$label)){
-    node_data$hovertext <- node_data$from
-  } 
-  edge_data <- unique(data[, edge_names])
+  htext <- apply(t(apply(do.call("rbind",strsplit(node_data$hovertext, "<br>")), 1, unique)), 1, paste, collapse = "<br>")
+  node_data$hovertext <- paste0("Node ID:", node_data$from,
+                                 "<br>", htext)
+  # edge_data <- unique(data[, edge_names])
+  edge_data <- data
   edge_data <- edge_data[which(!edge_data$.selfie),]
   edge_data$size <- edge_data$linewidth
   edge_data$hovertext <- NULL

### 2017-04-05 R/stat-net.r

@@ -130,6 +130,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list(), si
 
     edges <- rbind(edges, fromonly[, names(edges)])
   }
+
   edges <- mutate(group_by(edges, from, to), weight = n())
   unique(edges)
 },
@@ -147,7 +148,7 @@ compute_panel = function(self, data, scales, na.rm = FALSE,
       data <- mutate(group_by(data, group), .samegroup = to %in% unique(from))
     }
 
-   # #browser()
+#    browser()
     data.frame(data)
   },
 

### 2017-04-05 README.Rmd
plotly ex

--- 2016-12-06 README.Rmd
@@
 ```
+
+## Interactivity with `plotly`!
+
+Now including interactivity with `ggplotly()`!
+
+```{r football, eval = FALSE}
+library(geomnet)
+# devtools::install_github("ropensci/plotly")
+library(plotly)
+data(blood)
+bloodnet <- fortify(as.edgedf(blood$edges), blood$vertices)
+p <- ggplot(data = bloodnet, aes(from_id = from_id, to_id = to_id))
+
+# create data plot
+p2 <- p + geom_net(aes(size=Predominance, colour=type, shape=rho, linetype=group_to),
+             linewidth=0.75, labelon =TRUE, directed = TRUE, labelcolour="black") +
+  facet_wrap(~Ethnicity) +
+  scale_colour_brewer(palette="Set2") 
+ggplotly(p2)
+```

### 2017-04-05 README.Rmd
add a plotly example

--- 2016-12-06 README.Rmd
@@
 ```{r football, eval = FALSE}
+# blood donation example
 library(geomnet)
@@
 library(plotly)
+data(blood)
+bloodnet <- fortify(as.edgedf(blood$edges), blood$vertices)
+p <- ggplot(data = bloodnet, aes(from_id = from_id, to_id = to_id))
+
+# create data plot
+p2 <- p + geom_net(aes(size=Predominance, colour=type, shape=rho, linetype=group_to),
+             linewidth=0.75, labelon =TRUE, directed = TRUE, labelcolour="black") +
+  facet_wrap(~Ethnicity) +
+  scale_colour_brewer(palette="Set2") 
+ggplotly(p2) %>% hide_legend()
+
+# Classic College Football Example
 data("football")

### 2017-04-05 R/geom2plotly.R

@@ -13,9 +13,10 @@ to_basic.GeomNet <- function(data, prestats_data, layout, params, p, ...) {
   # edge_names <- c("from","to", "x", "y","xend", "yend", ".selfie", "weight", ".samegroup", "ymax", "xmax", "width", "linetype", "linewidth", "fontsize", "arrowsize", "stroke", "PANEL")
   node_data <- unique(data[, node_names])
   node_data <- getFromNamespace("prefix_class", asNamespace("plotly"))(node_data, "GeomPoint")
-  htext <- apply(t(apply(do.call("rbind",strsplit(node_data$hovertext, "<br>")), 1, unique)), 1, paste, collapse = "<br>")
+  
+  htext <- apply(t(apply(do.call("rbind",strsplit(node_data$hovertext, "<br />")), 1, unique)), 1, paste, collapse = "<br />")
   node_data$hovertext <- paste0("Node ID:", node_data$from,
-                                 "<br>", htext)
+                                 "<br />", htext)
   # edge_data <- unique(data[, edge_names])
   edge_data <- data
   edge_data <- edge_data[which(!edge_data$.selfie),]

### 2018-10-09 README.Rmd
website fixed

@@
   collapse = TRUE,
-  comment = "#>",
-  fig.path = "README-"
+  comment = "#>"
 )

### 2020-10-11 README.Rmd
- Update to fix changes in dplyr - specifically use `dplyr::n()` - Use `inhereits(dat, "matrix")` instead of `class(x) =! "matrix"` - Adds spell check WORDLIST

@@
 `devtools::install_github("sctyner/geomnet")`
+
 ```{r, echo = FALSE}
@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  fig.path = "man/figures/README-"
 )

### 2020-10-11 R/stat-net.r

@@ -73,7 +73,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list(), si
     edges <- edges[!edges$from_id %in% true_singletons,]
   }
   edges <- dplyr::group_by(edges, from_id, to_id)
-  edges <- dplyr::summarise(edges, wt = n())
+  edges <- dplyr::summarise(edges, wt = dplyr::n())
 
     # there should not be any missing values at this point, but just make sure
   if (any(is.na(edges$from_id))) message(sprintf("%d missing values excluded\n", sum(is.na(edges$from_id))))
@@ -131,7 +131,7 @@ compute_network = function(data, layout.alg="kamadakawai", layout.par=list(), si
     edges <- rbind(edges, fromonly[, names(edges)])
   }
 
-  edges <- mutate(group_by(edges, from, to), weight = n())
+  edges <- mutate(group_by(edges, from, to), weight = dplyr::n())
   unique(edges)
 },
 

### 2020-10-11 R/utils.R

@@ -4,7 +4,7 @@
 #' @param dat A square adjacency matrix.
 #' @export
 as.adjmat <- function(dat){
-  if (class(dat) != "matrix"){
+  if (!inherits(dat, "matrix")){
     stop("as.adjmat requires a matrix object")
   }
   if (nrow(dat) != ncol(dat)){

### 2020-11-19 README.Rmd
fixing little issues as i prepare for release

@@
 [![CRAN Status](http://www.r-pkg.org/badges/version/geomnet)](https://cran.r-project.org/package=geomnet)
-[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/geomnet)](http://www.r-pkg.org/pkg/geomnet)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/geomnet)](https://www.r-pkg.org/pkg/geomnet)
 [![Travis-CI Build Status](https://travis-ci.org/sctyner/geomnet.svg?branch=master)](https://travis-ci.org/sctyner/geomnet)

### 2020-11-19 R/data.r

@@ -53,7 +53,7 @@
 #' Email network (directed)
 #'
 #' A list of two datasets, vertices and edges, containing data on employees and (fictitious)
-#' email headers from two weeks of internal GAStech company email between employees made available as part of the VAST challenge 2014 (K. Cook, G. Grinstein, M. Whiting, see \url{http://hcil2.cs.umd.edu/newvarepository/benchmarks.php}).
+#' email headers from two weeks of internal GAStech company email between employees made available as part of the VAST challenge 2014 (K. Cook, G. Grinstein, M. Whiting, see \url{http://www.cs.umd.edu/hcil/varepository/benchmarks.php}).
 #' The variables are as follows:
 #'
 #' @format A list of two data frames:
@@ -245,7 +245,7 @@
 
 #' A protein interaction network (undirected)
 #'
-#' A list of two datasets, vertices and edges, containing data on the complete protein-protein interaction network in the yeast species S. cerevisiae (\url{http://www3.nd.edu/~networks/resources.htm}).
+#' A list of two datasets, vertices and edges, containing data on the complete protein-protein interaction network in the yeast species S. cerevisiae (\url{https://snap.stanford.edu/data/S-cerevisiae.html}).
 #' The variables are as follows:
 #' @references H. Jeong, S. Mason, A.L. Barabasi and Z.N. Oltvai, Centrality and lethality of protein networks Nature 411, 41 (2001)
 #' @format A list of two data frames:
@@ -302,7 +302,7 @@
 
 #' \pkg{ggplot2} theme attribute inheritance network (directed)
 #'
-#' A list of two datasets, vertices and edges, containing data on the inheritance structure of theme elements in \pkg{ggplot2} (see \url{http://docs.ggplot2.org/current/theme.html})
+#' A list of two datasets, vertices and edges, containing data on the inheritance structure of theme elements in \pkg{ggplot2} (see \url{https://ggplot2.tidyverse.org/reference/theme.html})
 #' The variables are as follows:
 #'
 #' @format A list of two data frames:

### 2020-11-20 README.Rmd
upload new hex sticker!

--- 2020-10-11 README.Rmd
@@
 ---
-title: "The `geomnet` package"
-author: Sam Tyner, Heike Hofmann
-date: "`r Sys.Date()`"
 output: github_document
@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+
+# geomnet <img src="man/figures/logo.png" align="right" width="120" />
+
 [![CRAN Status](http://www.r-pkg.org/badges/version/geomnet)](https://cran.r-project.org/package=geomnet)
-[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/geomnet)](http://www.r-pkg.org/pkg/geomnet)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/geomnet)](https://www.r-pkg.org/pkg/geomnet)
 [![Travis-CI Build Status](https://travis-ci.org/sctyner/geomnet.svg?branch=master)](https://travis-ci.org/sctyner/geomnet)
